### 2011-10-06 doc/manuscript.tex
The manuscript is being rewritten.

% We approach the project of streptococcus recombination and gene transfer from
% the angle of a resource paper. 
\documentclass[english]{article}
\usepackage[T1]{fontenc}
\usepackage[latin9]{inputenc}
\usepackage[letterpaper]{geometry}
\geometry{verbose,tmargin=1in,bmargin=1in}
\setlength{\parskip}{\bigskipamount}
\setlength{\parindent}{0pt}
\usepackage{url}
\usepackage{graphicx}
\usepackage{setspace}
\usepackage[authoryear]{natbib}
\usepackage{longtable}
\doublespacing

\newcommand{\chck}{\textbf{[CHECK]}}
\newcommand{\addfig}{\textbf{[ADD FIGURE]}}
\newcommand{\tba}{\textbf{[TO BE ADDED]}}

\makeatletter

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% LyX specific LaTeX commands.
%% Because html converters don't know tabularnewline
\providecommand{\tabularnewline}{\\}
%% A simple dot to overcome graphicx limitations
\newcommand{\lyxdot}{.}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% Textclass specific LaTeX commands.
\newcommand{\lyxaddress}[1]{
\par {\raggedright #1
\vspace{1.4em}
\noindent\par}
}

\@ifundefined{showcaptionsetup}{}{%
 \PassOptionsToPackage{caption=false}{subfig}}
\usepackage{subfig}
\makeatother

\usepackage{babel}

\begin{document}
\renewcommand\refname{Literature Cited}

\raggedright

\subsubsection*{Title: }

Inferring homologous recombination and horizontal gene transfer 
using whole genomes of \textit{Streptococcus pyogenes} 
and\textit{ S.\ dysgalactiae} subspecies \textit{equisimilis}

\subsubsection*{Authors:}

Sang Chul Choi$^{1}$, Matthew D. Rasmussen$^{1}$, Melissa Jane Hubisz$^{1}$, Ilan 
Gronau$^{1}$,
Michael J. Stanhope$^{2}$, and Adam Siepel$^{1}$

\subsubsection*{Affiliations:}

\lyxaddress{$^{1}$Department of Biological Statistics and Computational Biology,
Cornell University, Ithaca, NY 14853, USA.}

\lyxaddress{$^{2}$Department of Population Medicine and Diagnostic Sciences,
College of Veterinary Medicine, Cornell University, Ithaca, NY 14853,
USA.}

\subsubsection*{Author for Correspondence:}

Adam Siepel, Department of Biological Statistics and Computational
Biology, Cornell University, Ithaca, NY 14853, USA, Phone: 607-254-1157,
Fax: 607-255-4698, Email: acs4@cornell.edu\clearpage{}

\subsubsection*{Abbreviations:}

group A streptococci (GAS), group C streptococci (GCS), group G streptococci
(GGS), \textit{Streptococcus pyogenes} (SPY), \textit{S.\ dysgalactiae}
subspecies \textit{dysgalactiae} (SDD), \textit{S.\ dysgalactiae} subspecies
\textit{equisimilis} (SDE)

\subsubsection*{Keywords:}

bacterial recombination, gene duplication and transfer, ClonalOrigin,
ClonalFrame, Mowgli

\subsubsection*{Version: 0.67 (Oct 7, 2011) \clearpage{}}

\section*{Abstract}
Homologous recombination and horizontal gene transfer are investigated with
complete genomes of closely related bacterial species of \textit{Streptococcus
pyogenes} (SPY) and \textit{S.\ dysgalactiae} subspecies \textit{equisimilis}
(SDE).  The former is a pervasive human-specific pathogen, and the latter an
opportunistic human pathogen.  We aims at which parts of the genomic
organization and human pathogenic virulence factors of SDE could have
originated in those of SPY, or vice versa.  Our genome-wide survey of gene
transfer indicated that gene transfer via homologous recombination occurred
relatively from SPY to SDE.  Although we did not find enrichment of virulence
genes in the homologous recombining gene transfer, we found evidence that
virulence genes transferred more often via non-recombining fashion without
replacing parts of destination genomes than did non-virulence genes.  A
genome-wide resource of recombining and non-recombining gene transfer for the
study of the genomes is provided so that research community in pathogenic
bacteria could take advantage of it.

\clearpage{}

\section{Introduction}

Bacteria infect humans with a variety of diseases: e.g., food poisoning
(\textit{Eschericha Coli}), strep throat (\textit{Streptococcus pyogenes}),
tuberculosis (\textit{Mycobacterium tuberculosis}), cholera (\textit{Vibrio
cholera}), anthrax (\textit{Bacillus anthracis}), pneumonia
(\textit{Chlamydophila pneumoniae}), and influenza (\textit{Haemophilus
influenzae}), to name a few.  \textit{Streptococcus pyogenes} (SPY) belonging to
group A streptococci (GAS) primarily infect humans in the throat and skin
\citep{Broyles2009}. It has high prevalence throughout the world causing
diseases ranging from mild illnesses such as pharyngitis to severe invasive
streptococcal diseases including necrotizing fasciitis and toxic shock syndrome
\citep{Cunningham2000a}.  Its close relative, \textit{Streptococcus
dysgalactiae} subspecies \textit{equisimilis} (SDE) belonging to group C or G
streptococci (GCS/GGS), used to be generally regarded as a veterinary pathogen
\citep{Vandamme1996}. However, it has been reported to cause serious and
life-threatening streptococcal diseases to humans traditionally associated with
GAS \citep[e.g.,][]{Brandt2009}.  Because of the overlap in disease profile and
ecological niche of tissue sites betwen SPY and SDE, these bacteria may as
well interact with each other by exchanging parts of their genomes.

Therefore, the research topic on gene exchange between SPY and SDE has been
performed in various ways. \citet{Sachse2002} showed gene transfer from SDE to
SPY for the downstream of the genomic region of streptococcal pyogenic
exotoxins. \citet{Davies2007} investigated phage 3396 from SDE as a vehicle of
gene transfer from SDE to SPY. \citet{Davies2009} demonstrated that an
integrative conjugative element of SDE allowed gene transfer from SDE to other
streptococci.  \citet{Ahmad2009} did not conclusively deduce the net
directionality of housekeeping gene transfer between SPY and SDE.  Notably,
\citet{Kalia2001} attempted to show that the net direction of gene transfer from
SPY to SDE was more dominant than the other direction.  Because these researches
on the controversial issue, net directionality of gene transfer between SPY and
SDE, have been limited to relatively small regions in the genomes primarily on
representative loci from the Multi Locus Sequence Typing (MLST) database, it
remains to be elusive to determine the genome-wide degrees to which genomic
segments flowed between the two species, and net directions of thereof.

As sequencing costs plummet \citep{Mardis2011}, bacterial genome sequencing
produces individual genomes of high quality even within a bacterial species
\citep{Tettelin2009a}.  Studies on horizontal gene transfer in prokaryotes have
used genes in very large sets of species \citep{Koonin2001} from a great deal of
genomes available.  Although this type of study gave us insights into deep
evolutionary histories of prokaryotes, with multiple individual genomes per
species we now can go beyond the rough broad study on gene transfer towards
fine-scaled history of gene transfer between very closely related prokaryotes.
For instance, \citet{Liu2009} studied gene transfer events in
\textit{Lactobacillus bulgaricus} and \textit{Streptococcus thermophilus} using
the compositional approach of \citet{Karlin2001}. \citet{Luo2011} investigated
gene transfer events in \textit{Eschericha Coli} using the embedded quartet
decomposition due to \citet{Zhaxybayeva2006}.  \citet{Hamady2006} combined
phylogenetic and compositional approaches to detecting horizontal gene transfer.
\citet{Caro-Quintero2011} studied a number of \textit{Shewanella baltica}
genomes to identify the genetic elements that enabled the species to adapt to
redox gradients.  Although several complete genomes from SPY have been
available, \citet{Suzuki2011} just recently added one more complete genome of
SDE to the set of a previously unique SDE genome \citep{Shimomura2011}. These
multiple SDE genomes prompted us to study genome-wide gene flows between SPY and
SDE that afflicted human beings around the globe.

Bacteria evolves new biological features through horizontally transferred genes
from other lineages and vertically transmitted mutations on genes within the
same lineage.  Mutations on a gene modifiy relatively small portions of the gene
through substitutions, insertions, and deletions of deoxyribonucleic acids.
Gene transfers can reshape the organization of a bacterial genome,
which could revamp biological functions of the species.  Recipient bacteria of
transferred genes can substitute transferred genes for homologous parts of their
genomes, or integrate transferred genes into their genomes without replacement.
These two processes are here in this manuscript discriminated by calling the
former \textit{recombining gene transfer}, and the latter \textit{horizontal
gene transfer} \citep{Ochman2001,Lawrence2009}.  Despite the two distinct
evolutionary processes in bacteria, analyses of complete genome sequences in
efforts of better understanding bacterial evolution have paid less attention to
a combined approach to illuminating disparity of the two gene transfer
processes. We attempted here to use and compare two approaches for
distinguishing recombining gene transfer process from horizontal, or
non-recombining, gene transfer, and tackled their roles in association with
virulence genes.

Homologous recombining gene transfer could be studied more readily using the
core genomes of individual bacteria while horizontal gene transfer might be more
easily detected using the pan genomes. Recently, a method of inferring
homologous recombination with whole bacterial genomes was developed
\citep{Didelot2010}. The model of homologous recombination allowed them to study
flows of genetic segment between bacteria individuals of \textit{Bacillus
cereus} species in the core genomes.  Other evolutionary processes acting on
bacterial genomes include genome rearragement, translocation, inversion,
duplication, gain, loss, etc.  These processes did not readily lend themselves
to statistically more sound model-based approaches such as maximum likelihood.
Instead, a parsimony-based approach to inferring gene gain, loss, and transfer
events was developed by \citep{Doyon2011}.  We noted that the first approach
could be used to infer transfer of genomic segments due to homologous
recombination, and the second could do similarly attributed to horizontal gene
transfer.

In short, the complete genomes of the two spcies of SPY and SDE presented us a
unique opportunity to study gene transfer over relatively recent time scales.
We took advantage of the model-based approach for inferring homologous
recombination, and the parsimony-based one for detecting gene duplication, loss,
and horizontal transfer. We compared both of the results in recombining gene
transfer events to show that the parimony-based method inferred recombining gene
transfer events in biological functional categories similar to those inferred
with the model-based approach. This was the first attempt as far as we know to
analyaze SPY and SDE in genome-wide scale for detecting gene transfer events not
only in the core genome but also in the pan genome, which makes the analysis
complete in terms of the amount of used genome data. Our analysis of core
genomes also showed for the first time that the net directionality of
recombining gene transfer from SPY to SDE was stronger than the reverse
direction.

\clearpage{}

\section{Results}

\subsection{Recombination drived more substitutions than did point mutation}

Five genomes except \textit{S.\ equi} ssp.\textit{\ equi} strain 4047 were
chosen (Table S\ref{tab:genome}) to study recombining gene transfer using the
model-based approach.  The genome alignment of the five genomes showed two
distinct genome structures of SPY and SDE (Figure S\ref{fig:mauve}). The SDE and
SDD genomes were larger in length than those of SPY. Although it was unclear
whether SDD was different from either of SDE1 and SDE2 from the graphical view
of the genome alignment, the spcies tree inferred from the genome alignment
showed that SDD branch was attached to the clade branch of SDE1 and SDE2 (Figure
S\ref{fig:tree5}). The distribution of the alignment block lengths was skewed to
the right with mean of 4188 base pairs long (Figure S\ref{fig:blocksize}), which
supported that the model assumption of geometric block length distribution might
be plausible in \texttt{ClonalOrigin} method (see section \textbf{Materials and
Methods} and \citealt{Didelot2010}).  The relative strength ($r/m$) of
recombination compared to mutations indicated that recombination-driven
substitutions affected more sites than did mutation-driven substitutions (Table
S\ref{tab:clonalframe}). This could imply that in the \textit{Streptococcus}
species bacterial recombination process appeared to accelerate molecular
evolution of more sites than did point mutations process.

\subsubsection{Recombination hotspots were observed in genomes}

Along the genome of SDE1 population mutation rate did not appear to change much
(Figure S\ref{fig:scatter3}).  We found a few spots of genomic regions where the
mutation rates were larger than twice the global estimate of mutation rate.  The
recombinant tract length changed in a way that we could not note any particular
positions with notable tract length values compared with other positions along
the genome.  As recombinant tract lengths were smaller, recombination rates per
site seemed larger (Figure S\ref{fig:rhologdelta}).  The recombination rate
appeared to change more dynamically. Especially, a genomic region of SDE1 near
20 kilobase pairs from the origin of replication contained alignment blocks with
bell-shaped recombination rates.

\subsection{SPY and SDE were more closely related in about tenth of the genome}

There are 105 possible rooted topologies relating five taxa.  We expected the
dominant topology across sampled genealogies to be concordant with that of the
species tree.  Indeed, 66.7\% of the local gene tree topologies (see section
\textbf{Materials and Methods}) along the alignment blocks were the same as the
species tree, which allowed us to be more confident of the species tree inferred
using the alignment blocks (Table S\ref{tab:Gene-tree-topologies}).  Other
frequently sampled topologies were expected to be potentially informative about
common recombination events.  The second most frequent topology (9.3\%) was one
where SDD was the outgroup, and SPY and SDE were sister clades. Therefore, SPY
and SDE appeared be more closely related with each other than each to SDD's
genome in about tenth of the core genome. The other topologies had less than 4\%
of proportions.

\subsection{Unbalaned recombining gene flow from SPY to SDE in genome-wide scale}

The heat map of Figure \ref{fig:Heatmap-of-recombination} (see Tables
S\ref{tab:heatmap} and S\ref{tab:obsheatmap} for the heat map values and the
observed number of recombinant edges) shows the log base 2 ratios of the
\textit{a posteriori} number of recombinant edges for a pair of donor and
recipient species tree branches and their corresponding expected values given
the species tree (Figure S\ref{fig:tree5}) and the three population parameters
(Table S\ref{tab:three}).  Because recombinant edges from any species tree
branch to the root branch are impossible (see Figure S\ref{fig:clonalorigin}),
the column for the root as recipient was colored green, indicating invalid
recombinant edges.  Similarly, all of the cells for which a recipient branch was
older than the donor, and two branches did not share any contemporary time were
colored green. The diagonal cells were colored gray, indicating no significant
number of recombinant edges, because the observed and expected numbers of
recombinant edges were similar. As recombinant edges starting from and ending in
the same species tree branch would not change the species tree, which would not
change the likelihood much, the cells would be colored gray.  Because SDD is a
strictly veterinary pathogenic bacteria, we did not expect that SDD received
much gene transfer from either SPY1 or SPY2: the log base 2 ratio from SPY1 to
SDD was -2.20, and that from SPY2 to SDD was -1.97.  On the contrary, the
recombing gene transfer from SDE1 to SDD was 0.756, and that from SDE2 to SDD
was 0.698. This relatively higher recombining gene transfer from two SDE
individuals to SDD compared with that from two SPY individuals to SDD reflected
that the SDD was more closely related to SDE than to SPY in most of the genomes,
and the genomic organization between SDD and SDE1/SDE2 were much similar
(Figure S\ref{fig:mauve}).  However, the gene transfer from SDD to either SDE1
or SDE2 was less frequent than the case of reverse direction; SDD effectively
served as a recipient rather than a donor. It was not unclear why SDD received
more genes from SDE1 or SDE2 than did SDE1 or SDE2 from SDD.

Because it was expected that the recombining gene transfer within a clade would
be more frequent than that between clades, not surprisingly the recombining gene
transfer log base 2 ratios between SDE1 and SDE2 (1.17 and 1.20) were about
twice as large as that from SDE1 and SDE2 to SPY1 and SPY2 (0.67, 0.66, 0.60,
and 0.39).  The intensities of recombining gene transfer from external branches
of SDE1 and SDE2 to external branches of SPY1 and SPY2 was similar to those from
SDE1 or SDE2 to SDD (0.76 and 0.70).  After all, either SPY1 or SPY2, and SDD as
recipient appeared to be no different from the view point of SDE1 or SDE2 as
donor since the similar values of recombining gene transfer intensity from SDE1
or SDE2 to the other three external branches.  Surprisingly, the recombining
gene transfer log base 2 ratios between SPY1 and SPY2 (1.06 and 1.07) was
comparable with, but even less than those from SPY1 and SPY2 to SDE1 and SDE2
(1.25, 1.25, 1.17, and 1.29).  Effectively, the recombining gene transfer from a
clade SPY1 or SPY2 to a different clade SDE1 or SDE2 seemed to be no different
from that within the clade of SPY1 and SPY2.  The two external branches of SDE
received genes from either SPY1 or SPY2 via recombining gene transfer as if SDE1
and SDE2 were in the same clade as SPY1 and SPY2.  This was more surprising
because the bacterial recombination between SPY and SDE seemed to happen more
often than that between SDE and SDD even though the genome configurations of SDE
and SDD were more similar than each to SPY.  However, this result could be
concordant with the fact that SPY and SDE shared the same ecological niche
despite their different genomic organization.  The gene transfer log base 2
ratio from SPY1 and SPY2 to SDE branch was also larger than any type of
recombining gene transfer: 1.88 and 1.94, respectively.  Additionally we found
58 genes for which a recombinant edge of certain types were sampled with
probability at least 0.9 by focusing on specific types of recombinant edges
(Table S\ref{tab:genes-transfer}).  Many of the genes in the list were
transferred from SPY to SDE rather than the reverse direction overall.  All
things considered, this epitomized that the recombining gene transfer was biased
toward the net direction from SPY to SDE.

We tried in vain to quantify differential horizontal gene transfer between SPY
and SDE with the parsimony-based approach (Figure \ref{fig:hgt-heatmap}).  Among
the clade-crossing transfers, we did not see any significant enrichment for a
specific direction, as opposed to the result for recombining transfers from the
model-based analysis (Figure S\ref{fig:mowgli-recomb-heatmap}).  SDE1 and SDE2
shared the most events (26 and 34) compared to other sample pairs.  SDD appeared
to acquire many genes from the ancestral SDE branch.  Transfer events between
the SPY and SDE clades were inferred in smaller numbers (<6) compared to those
within each clade (23--34). We were puzzled by this finding because virulence
genes were found to be associated with horizontal gene transfer, and we could
not show net directionality of horizontal gene transfer between SPY and SDE.

\subsection{More frequent recombining gene transfer than horizontal gene transfer}

Overall, we inferred 304 duplications, 944 recombining gene transfers, and 294
horizontal gene transfers, comprising 23.7\% of total transfer events (Figure
\ref{fig:Gene-duplication-loss}).  Among the 2314 gene families the largest
family consisted of 59 genes, while nearly half of gene families (1066) had
exactly one gene per species, or six genes total (Figure S\ref{fig:famsizes}).
Within clades we found that duplications made up only a small fraction of the
events  (10.2\%), while horizontal transfers and losses were significantly more
frequent (41.7\% and 48.0\%, respectively). Duplications were most often
species-specific.  The species SDD and SDE were the most expanded, both in terms
of duplications and in-bound transfers.

\subsection{Virulence genes via recombining gene transfer}

About 10\% of the genes comprised of virulence facters, about half of which were
found in the core genome of the five genome alignment. Using the UCSC genome
browser tracks we located virulence genes with notable posterior recombination
probability.  These genes were by no means comprehensive list of virulence
factors that might have been transferred via homologous recombination.
Stressful host conditions for pathogeneic bacteria can include heat shock,
starvation, and iron depletion.  Iron uptake of bacteria was known to be
essential for survival of the bacteria under hostile host condition. We found
that parts of putative calcium transporter (MgtA) in SDE1 could have been
transferred from SPY1 (SpyM3\_0440).  An extracellular protein, putative
dipeptidase (SpyM3\_0465) that could be important for the survival of bacteria
to scavenge dipeptides from the host, was inferred to be transferred from SPY to
SDE: the whole gene was transferred, which was a case for the finding that
virulence genes tended to be transferred as a whole unit \citep{Chan2009}.
Another example of whole gene transfer was superoxide dismutase that was
reported to be associated virulence of \textit{Streptococcus agalactiae}
\citep{Poyart2001}: the whole gene including its 5' UTR, (SpyM3\_1071), was
affected by recombining gene transfer from SPY to SDE.  Some of the genes were
only affected in their parts.  The middle parts of an adhesins gene, putative
internalin A precursor (SpyM3\_1035), was found to be affected by recombining
gene transfer from SPY to SDE.  The middle part of a chaperonin GroEL, putative
heat shock protein (SpyM3\_1765), was inferred to transfer from SPY to SDE1:
increase of GroEL expression was known to allow for the bacteria to survive the
hostile host environment of acid shock.  Parts of the gene (SpyM3\_1749),
pyruvate formate lyase that was recently reported to have virulence
\citep{Yesilkaya2009}, seemed to be affected by recombining gene transfer.
Under starvation condition bacteria must use more effective energy metabolism: a
putative glucose kinase, (SpyM3\_1180), were inferred to be transferred from SPY
to SDE to adapt SDE's energy metabolism to survive such harsh condition.  We
also found that an MLST locus, called DNA mismatch repair protein (MutS),
appeared to be affected by recombining gene transfer.

\subsection{Virulence genes via horizontal gene transfer}

Horizontal gene transfers were found to be significantly associated with
virulence genes (P-value of $<1.33 \times 10^{-11}$) whereas recombining gene
transfers were not enriched for virulence genes (P-value of $>0.90$) using the
parsimony-based approach.  This was also consistent with the fact that we could
not find the association of recombining gene transfer with virulence genes in
the model-based approach.  This supported more evidence that non-recombining
gene transfer served as a more effective means by which SDE or SPY acquired
virulence genes, compared to recombining gene transfer.  
Families with high numbers of duplications were also significantly enriched for
virulence genes (P-value of $<7.13 \times 10^{-3}$) whereas families with high
losses showed no enrichment (P-value of $>0.54$).

Roughly, half of the virulence genes were in the pan genome.  We did not attempt
to comprehensively list virulence genes that were inferred to be transferred
horizontally. We rather sort virulence genes by the number of horizontal gene
transfer events to list the first several genes with high ranks.  Streptococcal
C5a peptidase (\textit{scpB}, SDEG\_0933) was inferred to have six horizontal
gene transfer.  Putative transposase were found to have a relatively large
number of horizontal gene transfer events (SpyM3\_1016, SpyM3\_1053,
SpyM3\_1054). Other genes with relatively large number of horizontal gene
transfer events included putative response regulator (SpyM3\_0174), putative
maltose operon transcriptional repressor (SpyM3\_0982), putative histidine
kinase (SpyM3\_0171), and putative cell wall hydrolase (SpyM3\_0731,
SpyM3\_0922, SpyM3\_1208).  Virulence genes encoding putative nucleases with a
secretion signal peptide were also found to have had experienced horizontal gene
transfer: genome-encoded streptodornase (SDEG\_0541), and prophage-associated
deoxyribonuclease (SDEG\_1103).  Protein F2 like fibronectin-binding protein
(SpyM3\_0104), an adhesin and invasin, was inferred to horizontally
transferred.

\subsection{Comparison between model-based and parsimony-based approaches}

Both of the model-based and parimony-based methods used discrepancy between a
reference species tree and gene trees to detect recombining gene transfer. In
essence the two methods worked in relatively similar manner. Firstly, functional
categories that were associated with recombining gene transfer were similar in
both methods (Tables S\ref{tab:functional-all} and
S\label{tab:go-events-recombining}).  Secondly, both of them did not show
association of virulence genes with recombining gene transfer. Lastly, we found
a positive correlation of 0.39 with P-value of $2.2\times10^{-16}$ between the
recombination intensity and the number of recombining gene transfer for the
genes (Figure S\ref{fig:cmpcomowgli}).  The correlation (0.03 with P-value of
0.18) disappeared in the comparison between recombination intensity and number
of horizontal gene transfer events (Figure S\ref{fig:cmpcomowglihgt}).

\subsection{UCSC genome browser for bacterial gene transfer study}

We created a web-resource for gene transfer available at
\url{http://strep-genome.bscb.cornell.edu}.  For each of the five individual
genomes as a reference genome several annotation tracks were displayed:
alignment blocks, recombination rate per site for each block, local gene tree
topologies, recombination intensity, locations of virulence genes, posterior
probability of recombining gene transfer.  
% To Matt: what could we show in the pan-genome? or from Mowgli's analysis?

\subsection{Units of recombining gene transfer}

The process of gene transfer may or may not move whole genes from one species to
another \citep[e.g.,][]{Chan2009}. We observed that units of recombining gene
transfer could happen with genes (Table S\ref{tab:genes-transfer}). In some
cases, parts of a gene were inferred to have been transferred, for instance,
transfer from SPY1 to SDE1 was inferred for 29\% of the aspartyl-tRNA synthetase
gene.  In other cases, the inferred transfer units covered a set of as much as
eight adjacent genes.  Of the genes listed in Table S\ref{tab:genes-transfer}
there were 12 groups of genes that recombinant tracts covered their gene
boudaries.  Most of the recombinant tracts often covered two genes. There were
transfer fragments that covered eight genes together: loci from SpyM3\_1518 to
SpyM3\_1525.

\subsection{Transporter, transferase, and aminoacyl tRNA synthetases}

Two classes of functional categories notably were associated with recombining
gene transfer.  Firstly, genes involved in sensing queues of environmental
changes would be important for bacteria to adapt to changing environment.  Genes
encoding membrane-bound proteins would be involved in the function.  Several of
functional categories that were associated with gene transfer might be related
with membrane-bound protein coding genes (Table \ref{tab:functional}).  Sugar
phosphotransferase classes were ranked top in the list: phosphrylation must be
important in signal transduction to recognize environmental changes.  The
transport functional categories of carbohydrate transport and sodium ion
transport also belonged to genes encoding membrane-bound proteins.  Also highly
ranked were fatty acid or lipid biosyntheses that must be involved in producing
membrane-bound molecules. These functional categories related with
membrane-bound proteins might allow for bacteria to adapt to changing
environment via recombining gene transfer.

Secondly, aminoacyl tRNA synthetases had been a focus of gene transfer
\citep{Woese2000}, and these aminoacyl tRNA synthetases belonged to the
functional category of translation (GO:0006412).  They were inferred to be 
associated with recombining gene transfer not horizontal gene transfer because
translation category was siginificant only in recombining gene transfer (see
Tables \ref{tab:functional}, \label{tab:go-events} and S\label{tab:go-events-recombining}).

In addition to the two kinds of functional categories, functional categories
concerning ribosomes were also associated with higher recombining gene transfer.
Because genes related with ribosomes were considered to be housekeeping genes,
gene transfer in ribosome genes would be rare, it was unexpected that ribosome
related genes were found to be associated with recombining gene transfer. This
was more significant when we considered all of the recombinant edges (Table
S\ref{tab:functional-all}) not just topology-changing recombinant edges (Table
\ref{tab:functional}) in computing recombination intensity.  Therefore, high
recombination intensity in ribosome related genes would likely to be attributed
to topology-not-changing recombinant edges (Table
S\ref{tab:functional-notopology}).  The parsimony-based approach also showed
enrichment of the recombining transfers in many of the same categories such as
translation, ribosome, and rRNA binding when using all types of recombinant
edges in computing recombination intensity (see Tables
S\ref{tab:go-events-recombining} and S\ref{tab:functional-all}).  Because the
parsimony-based method did not take species tree branch lengths into account, we
were likely to the agreement between the two different methods.

For horizontal gene transfer we found enrichment in very different categories
related to transposition, integration, and recombination (Table
\ref{tab:go-events}).  Gene duplication and loss events were enriched for
categories similar to those of horizontal gene transfer's enrichment (Tables
S\ref{tab:go-events-duplication} and S\ref{tab:go-events-loss}).
Gene duplication and horizontal transfers showed enrichment in the same
categories perhaps because both mechanisms increase gene copy number.  
Gene losses showed enrichment for transposition and integration, but less
significantly (P-value of $<10^{-6}$), and showed enrichment for several
additional categories such as transporter activity and transcription initiation.

Comparing two types of recombinant edges, one from SDE to SPY and another from
SPY to SDE transport related categories were more associated for the direction
from SPY to SDE
(Table S\ref{tab:functional-sde-spy} and S\ref{tab:functional-spy-sde}).

% More interpretations about Mawgli analysis of function categories?
% Why transposases?
% \texttt{IG: NEXT PARAGRAPH CONTAINS SOME DISCUSSION ON THESE RESULTS. WE SHOULD 
% GET BACK TO THIS ONCE WE DECIDE WHAT TO DO WITH THE RESULTS}

\section{Discussions}

We investigated closely related bacterial species to better understand gene
transfer processes between SPY and SDE. Two complementary methods were employed
to study recombining and horizontal gene transfers. By discriminating the two
processes in analyzing the genomes, we showed that the pathogenicity of SDE and
SPY was more attributed to horizontal gene transfer process than recombining
gene transfer.  Because recombination in bacteria seemed to occur at somewhat
random positions along the genome, the homologous parts transferred might be
unlikely to be enriched with virulence factors to human hosts.  Recombining gene
transfer might have rarely contributed to influx of virulence genes from SPY to
the opportunistic human pathogen of SDE as we showed a few cases of virulence
gene that might have been transferred via recombining gene transfer.

We also showed that the net amount of recombining gene flows from SPY to SDE was
relatively greater than that of the reverse direction in the core genome
sequences.  If there were unbalanced recombining gene transfer between the two
species, what were the barriers \citep{Thomas2005} to the gene transfer in the
bacteria? Although barriers to bacterial recombination would include geographic
isolation, genetic isolation could prevent recombination between coinhabiting
bacteria.  \citet{Kalia2001} discussed restriction-modification system (RMS) as
a possible mechanism of the unbalanced gene flow.  Restriction-modification
system is a biological self-defense mechanism of bacteria, which is diverse in
prokaryote world.  Biological roles of RMS might be to maintain and control
species identity \citep{Jeltsch2003}.  Recently, \citet{Budroni2011a} studied
that \textit{Neisseria meningitidis} phylogenetic clades were associated with
RMS, claiming that RMS modulated homologous recombination in the bacteria of
\textit{N.\ meningitidis}.  Most of \textit{S.\ pyogenes} genomes contain
restriction-modification type I system as well as type II system. The two
genomes of SDE contained only type II system. More tight self-defense system of
SPY may have made the species less capable of aquiring foreign genetic material
from SDE, and accordingly relatively unbalanced gene flow between SPY and SDE
may have been associated with different restriction-modification systems in the
two species \citep{Kalia2001}.

We used relatively small number of genomes to detect recombining gene transfer
that had only been possible using many samples for a few genes. We revisited the
problem of gene transfer in other previous studies.  We reexamined the seven
genes that \citet{Kalia2001} had studied for net directionality of gene transfer
between SPY and SDE.  Internal fragments of the six housekeeping genes, encoding
putative glucose kinase (\textit{gki}), glutamine transport protein
(\textit{gtr}), glutamate racemase (\textit{murI}), mismatch repair enzyme
(\textit{mutS}), transketolase (\textit{recP}), and xanthine
phosphoribosyltransferase (\textit{xpt}) were located in the core genome that we
investigated in this study.  Genes \textit{mutS}, \textit{recP}, and
\textit{xpt} had fragments of gene transfer with moderate posterior probability
of recombination. The other genes did not have noticeable recombination. Except
for \textit{mutS} where \citet{Kalia2001} did not find recombination, the result
was consistent with that of \citet{Kalia2001}.  We also revisited recombination
in gene \textit{parC} in SPY1 genome studied by \citet{Pinho2010}, where they
studied fluoroquinolone resistance in SDE and reported recombination in the
gene. We observed moderate recombination in the gene.

We used two heterogeneous methods: one model-based and another parsimony-based
for studying recombining and horizontal gene transfer.  The core genomes
consisted of sequence alignments with rare gaps. With this inherent filtering of
the input data preparation the model-based method did not accommodate gene
dupliation and loss in its core genome analysis. On the contrary, the
parsimony-based method used all of the genomes not just the core genomes.
Considering nucleotides as the level of gene duplication, loss, and transfer
towards a complete model for recombining and horizontal gene transfer could be
difficult because genome-scale application of those models based on DNA or
protein sequences might not be feasible.  Combining the two approaches for
studying the mode of gene transfer would be desirable to study more complicated
evolutionary processes including recombining gene transfer, gene duplication,
loss, and horizontal gene transfer.

\section{Materials and Methods}

\subsection{Complete bacterial genomes}

Two genomes of \textit{S.\ pyogenes} (SPY), and two of \textit{S.\ dysgalactiae}
ssp.\textit{\ equisimilis} (SDE) were downloaded from NCBI with the following
accessions: NC\_004070 \citep{Beres2002}, NC\_008024 \citep{Beres2006},
NC\_012891 \citep{Shimomura2011}, and CP002215 \citep{Suzuki2011}. We refer to
these samples as SPY1, SPY2, SDE1, and SDE2, respectively. The genome of
\textit{S.\ dysgalactiae} ssp.\textit{\ dysgalactiae} was provided by
\citet{Suzuki2011} with accession identifiers of SddyATCC27957, referred to here
as SDD. We also downloaded the genome (NC\_012471) of \textit{S.\ equi}
ssp.\textit{\ equi} strain 4047 referred to as SEE \citep{Holden2009}.

\subsection{Model-based method for recombining gene transfer}

We present a short summary of the statistical method developed by
\citet{Didelot2007} and \citet{Didelot2010} for studying bacterial
recombination. A detailed account was found in the two papers and in the user
guide available at \url{http://code.google.com/p/clonalorigin}.  We used
\texttt{progressiveMauve} v2.3.1 \citep{Darling2004,Darling2010} with default
options to attempt to align the six genomes, which led to too small proportion
of a core genome. We had to remove the SEE sample from the genome alignment
although we kept it in the parsimony-based approach.  We then used
\texttt{stripSubsetLCBs}, a program of the \texttt{ClonalOrigin} package, and
identified 276 blocks of sequence alignments with length threshold of 1500 base
pairs.  We excluded two of these blocks in a subsequent analysis because they
contained long consecutive gaps.  We applied \texttt{ClonalFrame} v1.1 available
at \url{http://www.xavierdidelot.xtreemhost.com} to the remaining 274 blocks,
and estimated the species tree relating the five genomes. \texttt{ClonalFrame}
estimates the average recombinant tract length and the ratio between
recombination and mutation rates, which were not used in the following analysis.
We used a single Markov chain Monte Carlo (MCMC) with 10000 burn-in generations
and 10000 sampling generations to obtain the tree estimate with branch lengths,
and additionally seven independent MCMC runs to validate adequate convergence of
the first run.

Two stages of MCMC with \texttt{ClonalOrigin}, subversion r19 that had been
downloaded from \url{http://clonalorigin.googlecode.com/svn/trunk}, were
employed independently for each of the 274 alignment blocks as described in
\citet{Didelot2010}.   In the first MCMC we obtained estimates of mutation rate,
recombination rate, and recombinant tract length.  Two independent chains were
executed with $10^6$ burn-in generations and $10^7$ sampling generations. We
sub-sampled the result every $10^5$-th generations.  A few alignment blocks were
not successfully finished even after a month of computation.  Because those were
not expected to affect the global estimates of population parameters
significantly, we used only finished blocks to estimate mutation rate,
recombination rate, and recombinant tract length.  The global estimators of the
three population parameters were medians weighted over block lengths, which
down-weighed less reliable estimates obtained from shorter alignment blocks.
One of the two independent chains was used to obtain the block-wise mean
estimates, and the other chain was used to validate convergence of the first
stage of MCMC.  Two independent MCMC runs resulted in similar estimates of the
three population parameters (Table S\ref{tab:three}).  The second stage of MCMC
with \texttt{ClonalOrigin} was executed with the three population parameters
fixed to the estimated values in the first stage.  Two independent MCMC chains
were also executed with a total generation being $1.1\times10^8$. The first
$10^7$ generations of each chain were discarded as burn-in. The remaining $10^8$
generations were sub-sampled every $10^5$-th generation.

\subsection{Local gene trees and recombination intensity}
Because the procedure of summarizing posterior sample of \textit{recombinant
tree} was not detailed in the user guide at the time when we used
\texttt{ClonalOrigin}, we elected to use Figure S\ref{fig:clonalorigin} to
illustrate a recombinant tree in order to clarify summarization of
\textit{recombinant history} posterior samples.  A recombination history is
represented by a \textit{recombinant tree}, which is a species tree augmented
with \textit{recombinant edges}.  Each recombinant edge is defined by the time
and branch of origin and destination, and the genomic interval that it affects.
The branch of origin is referred to as the \textit{donor} and the destination
branch is referred to as the \textit{recipient}. The type of a recombinant edge
is determined by the donor and recipient species tree branches.  A recombinant
tree can be viewed as a restricted version of an ancestral recombination graph
(ARG), and as such it induces a local genealogy for each site in the genomic
block it describes.  \citet{Didelot2010} presented two procedures of summarizing
recombinant trees by viewing trees from two different angles: by counting types
of recombinant edges regardless of genomic positions affected by the edges, and
by measuring recombinant edge changes to the species tree in site-wise manner.
By similarly following their approaches we quantified additional aspects of
recombinant trees; we counted gene tree topologies for each site, and also
quantified recombination intensity along the genome.

The recombination intensity allowed us to find genomic regions enriched for
recombination events in general, or regions enriched for recombination events of
certain types (e.g., between SPY and SDE).  We note that the recombination
intensity is not normalized to the interval $[0,1]$. Rather than giving
probabilistic meanings to the recombination intensity we interpreted it as an
average number of recombinant edge types per site.  The measure appeared to act
as an effective indicator of recombination intensity (see subsection of
\textbf{Simulation Study}).  Our simulation study indicated that edges of
certain types were better indicators of true recombination events than others.
For instance, recombinant edges which resulted in a recombinant tree with the
same topology as the species tree appeared to be more prone to spurious
sampling.  Such edges were ones where the donor was the same branch, sister
branch, or parent branch of the recipient branch.  Therefore, in some cases we
elected to remove such edges from consideration when measuring recombination
intensity.  The tree topologies changes would be correlated with recombination
intensity.  The tree topologies would show more effectively regions where
recombination occurs along the genome (see subsection of \textbf{UCSC Genome
Browser}).

\subsection{Parsimony-based method for horizontal gene transfer}

While \texttt{ClonalOrigin} could be used to understand recombining gene
transfers within the conserved core genome, we sought to understand the history
of the remainder of the genome where more complex evolutionary processes occur,
such as gene duplications, losses, and horizontal transfers.  In recent years,
several methods have been developed for inferring these evolutionary events in
gene families by reconciling gene trees with a species tree
\citep{David2011,Doyon2011,Tofigh2011}.  In this analysis, we complemented our
study of recombination by using \texttt{Mowgli} \citep{Doyon2011} to
characterize events that can change the copy number of a gene, such as
duplications, losses, and non-recombining horizontal gene transfers.  In
contrast to \texttt{ClonalOrigin}, \texttt{Mowgli} assumes a parsimony model for
inferring events and uses pre-determined gene clusters where genes are treated
as atomic units.  The advantage of using \texttt{Mowgli} was that it did not
assume the sequence data contains a single copy of each homologous segment, and
it attempted to explain different copy numbers by a series of gain and loss
events.  We here were interested in the approach of \texttt{Mowgli} to study
non-recombining horizontal gene transfers.  We obtained gene family clusters
defined in \citet{Suzuki2011} using the OrthoMCL method \citep{Li2003} over the
full set of six genomes including \textit{S.\ equi} subsp.\textit{\ equi} strain
4047 (SEE).  For each family cluster, the corresponding protein sequences were
aligned using the MUSCLE program \citep{Edgar2004a} and nucleotide alignments
were constructed by mapping nucleotide sequences onto each protein sequence
alignment. We then identified recombination within gene alignments using the
Single Breakpoint Recombination (SBP) method from the HyPhy package
\citep{KosakovskyPond2006}.  Using the Akaike Information Criterion (AIC), SBP
inferred that 47.5\% (1,094) of gene families contained at least one
topology-changing recombination. Each of these families was then split into two
gene fragments.  Gene trees were constructed from nucleotide alignments for each
gene fragment using the RAxML program \citep{Stamatakis2006}, and were rooted by
minimizing inferred duplications and losses.  Each rooted gene tree was then
reconciled using the \texttt{Mowgli} program \citep{Doyon2011}, which inferred
the most parsimonious gene duplication, loss, and transfer events for each
family.  In the case of multiple maximum parsimonious reconciliations, one was
chosen uniformly at random.  Unless specified otherwise, default setings were
used for all programs.  Since the method of \texttt{Mowgli} did not directly
model recombination events, recombination-based transfers were inferred in a
post-processing step, by checking whether the transfer was coupled with a loss
of a homologous gene in the recipient branch (indicating gene replacement rather
than gain).  A transfer event was classified as non-recombining if the
destination species of that transfer had an extant descendant that contained
copies of that genes which were descendants of the transferred gene as well as
copies which were non-descendants.  We noted that this parsimonious approach was
imperfect and would tend to undercount events, especially recombining gene
transfer events, which were modeled using two events: a transfer and a loss.
Nevertheless, this analysis allowed us to study non-recombining gene transfers,
which were not modeled by \texttt{ClonalOrigin}, and to validate some of the
\texttt{ClonalOrigin} results on recombining gene transfer.

\subsection{Functional categories association}

% FIXME
% \texttt{The recombination intensity scores were taken to be average site-wise
% intensity scores. SC: THIS NEEDS TO BE REANALYZED WITH RECOMBINATION INTENSITY
% WITH ONLY TREE-TOPOLOGY CHANGING RECOMBINANT EDGES.} 
We performed a series of analyses to check for functional gene categories
enriched for certain types of gene transfer events. Functional categories were
assigned in the same manner as \citet{Suzuki2011}. In short, blastP was used to
compare Streptococcus genes to bacterial proteins from the Uniref90 database.
Matches with E-values less than $1.0\times10^{-5}$ were assumed to have the same
Gene Ontology (GO) classification as the target gene, as determined from the
uniProt GOA database.  We also classified a category of  \textit{virulent genes}
using the virulence profile described in \citet{Suzuki2011}. Among 207 virulent
genes at least one of the genes in the gene family was identified by that
profile.  Each gene covered by our analysis was associated with a recombination
intensity score from the \texttt{ClonalOrigin} analysis and scores from the
\texttt{Mowgli} analysis corresponding to number of duplications, losses,
horizontal gene transfers, and recombining gene transfer events.  Mann-Whitney
tests were used to identify GO categories whose genes showed significantly
elevated values in each score.  A 5\% False Discovery Rate (FDR) correction
\citep{Benjamini1995} was used to determine an appropriate significance
threshold for our p-values.

\subsection{Simulation Study}

We conducted experiments on simulation data to check the accuracy of estimates
obtained by \texttt{ClonalOrigin} in a scenario similar to ours. We used a
simulation feature of \texttt{ClonalOrigin} to simulate the process of mutation
and recombination across the five species.  While populaiton mutation and
recombination parameters were accurately estimated, the recombinant tract length
was over-estimated  (Table S\ref{tab:sim-three-population-parameters}).  This
was not surprising as recombinant tract lengths were known to be difficult to
estimate from relatively short alignment blocks \citep{Didelot2010}. 
% When we
% performed simulations which much longer blocks (10000 bp), \texttt{ClonalOrigin}
% was able to estimate the average recombinant tract length accurately (data not
% shown).  
% FIXME: Show this result of 10,000 blocks.
We also evaluated the accuracy of the statistics estimated in the second stage
of \texttt{ClonalOrigin}. The number of recombinant edges for each pair of
recombinant tree branches was correlated with the true value in general while
there were slight over-estimation (Figure S\ref{fig:h1}).  However, this
comparison between inferrence from simulated data and that of real data set
suggested that the directional difference between SPY and SDE was due to neither
lack of power of \texttt{ClonalOrigin} nor hidden bias thereof.  Recombination
intensities that were measured on genes also showed rough correlation (Pearson's
correlation coefficient of 0.63) of the estimated and true values (Figure
S\ref{fig:ri1}).  Estimated values appeared to be driven somewhat by the prior
implied by the population parameter setting, explaining why lower values were
over-estimated and higher values were under-estimated.  Nonetheless, the
correlation indicated that the measure of recombination intensity used in our
functional gene category served as an adequate indicator of actual intensity of
recombination events along genes.

\clearpage{}

\section{Supplementary Information}

\subsection{Population mutation and recombination rates were accurately
estimated}

% FIXME: Newick tree in text in a file.
We describe the detailed procedure of simulations for checking if the three
population parameters were accurately recovered from the first stage of
\texttt{ClonalOrigin} MCMC.  Ten recombinant trees were sampled given the three
parameters and the species tree  set to the values inferred from the five
bacterial genomes (see section \textbf{Results}).  Each of the recombinant trees
was used to simulate the Jukes-Cantor model for DNA sequence evolution to create
274 blocks of alignments.  The block lengths configuration was the same as that
of the blocks used in the analysis of the five genomes.  The mean of the ten
estimates of the recombinant tract length was significantly higher than the true
value of (919 bp versus 744 bp) whereas the values of the mutation and
recombination rates were accurately estimated (Table
S\ref{tab:sim-three-population-parameters}).  Recombinant tract lengths were
known to be difficult to estimate from relatively short alignment blocks
\citep{Didelot2010}. Performing similar simulations with much longer blocks
(i.e., 10000 bp) we estimated the average recombinant tract length accurately
\texttt{(Table S data not shown)}.

\subsection{Estimating numbers of recombinant edges were accurate}

% FIXME 1: IG: RESULTS FROM THIS POINT ON REQUIRE SOME REFINEMENT. SC: Let's see.
% FIXME 2: ``TRUE" TO BE EXPECTED VALUE
% RATHER THAN AVERAGE OVER THE 10 REC.  TREES. SC: OKAY. LET'S FIND THE PRIOR
% EXPECTED NUMBER OF EVENTS. THIS MUST BE GIVEN SOMEWHERE}.  
% FIXME 3: CONSIDER SHOWING DETAILED COMPARISON OF ESTIMATED VALUES TO ONES 
% EXPECTED FROM THE ACTUAL REC. TREES. THESE DON'T LOOK AS NICE, BUT THEY MIGHT 
% BE ABLE TO DEMONSTRATE OUR ABILITY TO DETECT DEPARTURE FROM EXPECTATION. SC:
% THIS MAY BE ABOUT CHANGING AVERAGE TO PRIOR EXPECTED NUMBER OF EVENTS.
% FIXME 4: WE SHOULD PROBABLY CONSIDER ALTERNATIVE WAYS OF USING THESE RESULTS TO 
% ARGUE THAT OUR DIRECTIONAL SIGNAL IS REAL. SC: I MIGHT CONSIDER SIMULATIONS AND
% ARTIFICIALLY DOUBLE OR TRIPLE THE RECOMBINATION EVENTS OF ONE OF TYPES.
We checked if numbers of recombinant edge types were inferred correctly. Because
of the two-stage Bayesian method we first sampled ten recombinant trees given
the mutation rate, recombination rate, and tract length (see section
\textbf{Results}), and then created ten independent simulated data sets with the
Jukes-Cantor model for DNA sequence evolution using each of the recombinant
trees. This consisted of 100 replicates of data sets.  Figure S\ref{fig:h1}
shows the observed count of each recombinant edge type against the expected
count; the estimated numbers appeared to be slightly over-estimated, but in
correlation with the true numbers.

\subsection{Unbalanced gene flow model was adequate}

We checked if the model with unbalanced gene flow was adequate using posterior
predictive check. We used 100 recombinant trees from the posterior distribution
of the recombiant tree given the real data of the 274 alignment blocks in order
to simulate the ClonalOrigin model to create a new data set of 100 274-block
alignments.  Using the simulated data sets of 100 274-block alignments we
infer recombinant trees to draw the distribution of log base 2 ratio of
observed recombinant edge counts to the prior expected number of recombinant
edges. If the model with unbalanced gene flow was adequate, the log base 2
ratio from the real data set should center around the distribution. 
\texttt{What is the result of this simulation?}

\subsection{}

With the slight over-estimation of recombination events we were concerned with
any potential biases hidden in \texttt{ClonalOrigin} implementation. Because the
simulated data were generated with no directionality between SPY and SDE,  if we
estimate recombination events from the simulated data sets using
\texttt{ClonalOrigin} and compare those estimates from those from the real data
set, this could remove biases in \texttt{ClonalOrigin} implementation. We could
use inferred estimates of recombination from the simulated data under the prior
model of \texttt{ClonalOrigin} as the prior expected number of recombination.
Figure S\ref{fig:h2} shows a plot of numbers of recombination events from the
real data against those from the simulated data sets. Note that the numbers of
recombination events from simulated data sets appear to deviate from the line of
slope being 1 and intercept being 0.  The numbers of recombination events for
the recombinant edges from SPY to SDE increase by more than 2.5 fold whereas
those for the recombinant edges from SDE to SPY increase by about 1.7 fold.
This was not surprising because inference of number of recombinant edges was
quite accurate. We considered that subtraction of effects of number of
recombinant edges inferred from simulated data and subtraction of effects of the
prior expected number of recombinant edges were effectively equivalent.
However, this comparison between inferrence from simulated data and that of real
data set suggested that the directional difference between SPY and SDE was due
to neither lack of power of \texttt{ClonalOrigin} nor hidden bias thereof.

\subsection{Simulation for recombination intensity}

We tested if our measure of recombination intensity provided a reliable estimate
of the true intensity. For this purpose, we artificially annotaded our synthetic
blocks to genes according to the real annotations along the real bacterial
alignments. For each gene in each of the simulated 100 data sets, we plotted the
average estimated recombination intensity along that gene against the true
average intensity, i.e., the average number of different recombinant edge types
implied by the recombinant tree used in simulation (Figure S\ref{fig:ri1}).
Although the scatter plot did not strictly follow along the $X=Y$ line, there
was correlation of the estimated and true values (Pearson's correlation
coefficient of 0.63). Estimated values appeared to be driven somewhat by the
prior implied by the population parameter setting, explaining why lower values
were over-estimated and higher values were under-estimated. Nonetheless, the
correlation indicated that the measure of recombination intensity used in our
functional gene category served as an adequate indicator of actual intensity of
recombination events along genes.

\section*{Acknowledgments}

We are indebted to Xavier Didelot for his help in setting up \texttt{ClonalOrigin}
analysis.  We are thankful to Haruo Suzuki for his help
in using the data of SDD genome. This work was supported by the National
Institute of Allergy and Infectious Disease, US National Institutes
of Health, under grant number AI073368-01A2 awarded to A.S. and M.J.S.

\section*{Author Contributions}

Conceived and designed the experiments: SCC MDR MJS AS.
Performed the experiments: SCC MDR MJH.
Analyzed the data: SCC MDR MJH IG.
Contributed reagents/materials/analysis tools: MJS.
Wrote the paper: SCC MDR MJH IG.

\bibliographystyle{gberefs}
\bibliography{siepel-strep}

\clearpage{}

\section*{Tables}

\clearpage{}

\begin{table}
\caption{\label{tab:functional}Functional categories asscoiated with 
recombination intensity based on topology changing recombinant edge types.}
\begin{tabular}{cccc}
\hline 
P-Value & Count & GO term & Description\tabularnewline
\hline 
1.731e-06 & 38 & GO:0009401 & phosphoenolpyruvate-dependent sugar phosphotransferase system\tabularnewline
2.775e-06 & 33 & GO:0008982 & protein-N(PI)-phosphohistidine-sugar phosphotransferase activity\tabularnewline
4.581e-05 & 22 & GO:0006633 & fatty acid biosynthetic process\tabularnewline
0.0001248 & 20 & GO:0008610 & lipid biosynthetic process\tabularnewline
0.0002675 & 45 & GO:0008643 & carbohydrate transport\tabularnewline
0.000513 & 81 & GO:0006412 & translation\tabularnewline
0.0005617 & 13 & GO:0006814 & sodium ion transport\tabularnewline
0.001019 & 40 & GO:0030529 & ribonucleoprotein complex\tabularnewline
0.001203 & 12 & GO:0031402 & sodium ion binding\tabularnewline
0.00206 & 39 & GO:0003735 & structural constituent of ribosome\tabularnewline
0.002115 & 237 & GO:0008152 & metabolic process\tabularnewline
0.00215 & 26 & GO:0019843 & rRNA binding\tabularnewline
0.00287 & 371 & GO:0016740 & transferase activity\tabularnewline
0.003552 & 66 & GO:0005840 & ribosome\tabularnewline
\hline 
\end{tabular}
\end{table}
\clearpage{}

%
\begin{table}
\caption{\label{tab:go-events}Functional categories asscoiated with 
non-recombining horizontal gene transfer.}
\begin{tabular}{cccc}
\hline 
P-Value & Count & GO term & Description \\
\hline 
7e-61 &  67 & GO:0006313 & transposition, DNA-mediated\\
3e-58 &  59 & GO:0004803 & transposase activity\\
1e-30 &  52 & GO:0015074 & DNA integration\\
1e-28 &  23 & GO:0032196 & transposition\\
1e-13 & 527 & GO:0003677 & DNA binding\\
2e-12 &  86 & GO:0006310 & DNA recombination\\
1e-11 & 197 & GO:0003676 & nucleic acid binding\\
3e-09 &  36 & GO:0016987 & sigma factor activity\\
3e-09 &  36 & GO:0006352 & transcription initiation\\
3e-08 & 142 & GO:0043565 & sequence-specific DNA binding\\
\hline 
\end{tabular}
\end{table}
\clearpage{}

\section*{Figures\clearpage{}}

%=============================================================================
% Figure 1. Heat map of recombining gene transfer from ClonalOrigin
\begin{figure}
\includegraphics[scale=0.45]{figures/heatmap-recedge}
\caption{\label{fig:Heatmap-of-recombination}Heatmap of recombination events
using the five genomes.  Each cell of the heat map represents the logarithm base
2 of the number of recombination events inferred relative to its expectation
under the prior, for a donor/recipient pair of branches. The donor is given by
the y-axis; the recipient on the x-axis.  The green cells are meaningless
because the recipient species tree branch of the cells are older than the donor
species branch.}
\end{figure}
\clearpage{}
%=============================================================================

%=============================================================================
% FIGURE 2. - HGT Transfer heatmap
\begin{figure}
\begin{center}
% \includegraphics[width=5in]{figures/trans-hgt-heatmap}
\includegraphics[scale=0.45]{figures/heatmap-hgtedge}
\end{center}
\caption{\label{fig:hgt-heatmap}Distribution of horizontal gene transfers.
Each cell of the color matrix shows the number of inferred non-recombining
horizontal gene transfer events between the different branches of the species
tree.}
\end{figure}
\clearpage{}%
%=============================================================================

%=============================================================================
% Figure 3. Gene duplication, loss, and transfer
\begin{figure}
\includegraphics[width=6in]{figures/strep-events}
\caption{\label{fig:Gene-duplication-loss} Distribution of inferred gene
duplication, loss, and transfer events across the phylogeny.  Gene counts for
each extant and ancestral species are given at each node of the phylogeny.  For
each branch we denote the number of gains G*, duplications D*, losses L*,
non-recombining transfers T*, and recombinations R*.  The non-recombining
transfer and recombination counts are indicated on the recipient branch (donors
are not indicated).}
\end{figure}
\clearpage{}%
%=============================================================================

\clearpage{}\setcounter{figure}{0}
\setcounter{table}{0}
\renewcommand{\figurename}{Supplementary Figure}
\renewcommand{\tablename}{Supplementary Table}

\section*{Supplementary Tables}
\clearpage{}

%=============================================================================
% Table S1. List of six genomes.
\begin{table}
\caption{\label{tab:genome}The six complete bacterial genomes. SDD genome was
not yet available from NCBI.}
\noindent \begin{centering}
\begin{tabular}{cccc}
\hline 
Name & NCBI accession & Size (base pairs) & Reference\tabularnewline
\hline
SDE1 & CP002215 & 2159491 & Suzuki et al. (2011)\tabularnewline
SDE2 & NC\_012891 & 2106340 & Shimomura et al. (2011)\tabularnewline
SDD & N/A & 2141837 & Suzuki et al. (2011)\tabularnewline
SPY1 & NC\_004070 & 1900521 & Beres et al. (2002)\tabularnewline
SPY2 & NC\_008024 & 1937111 & Beres et al. (2006)\tabularnewline
SEE & NC\_012471 & 2253793 & Holden et al. (2009)\tabularnewline
\hline
\end{tabular}
\par\end{centering}
\end{table}
\clearpage{}
%=============================================================================

%=============================================================================
% Table S2. ClonalFrame Result.
\begin{table}
\caption{\label{tab:clonalframe}Parameter estimates from ClonalFrame. 
$\theta$ is population mutation rate,
$\delta$ is the average recombinant tract length,
$r/m$ is the relative strength of recombination compared to mutations,
TMRCA is the time to most recent common ancestor, and
$T$ is the total branch length of the inferred species tree.}
\noindent \centering{}\begin{tabular}{cc}
\hline
Parameter & Estimates (95\% credibility interval) \tabularnewline
\hline
$\theta$ & 92871.84 (N/A)\tabularnewline
$\nu$ & 0.134 (0.132, 0.135)\tabularnewline
$R$ & 9370 (8950, 9780)\tabularnewline
$\delta$ & 362 (347, 377)\tabularnewline
$r/m$ & 4.66 (4.45, 4.87)\tabularnewline
$\rho/\theta$ & 0.101 (0.096, 0.105)\tabularnewline
TMRCA & 0.331 (0.324, 0.337)\tabularnewline
$T$ & 1.02 (1.01, 1.04)\tabularnewline
\hline
\end{tabular}
\end{table}
\clearpage{}
%=============================================================================

%=============================================================================
% Table S3. Gene Tree topologies
\begin{table}
\caption{\label{tab:Gene-tree-topologies}Posterior probability of gene tree
topologies. See Table S\ref{tab:genome} for the reference of the bacteria
individual names.}
\noindent \begin{centering}
\begin{tabular}{cc}
\hline 
Gene tree topologies & Posterior probability\tabularnewline
\hline
(((SDE1,SDE2),SDD),(SPY1,SPY2)) & 66.7\%\tabularnewline
(((SDE1,SDE2),(SPY1,SPY2)),SDD) & 9.3\%\tabularnewline
((SDE1,SDE2),(SDD,(SPY1,SPY2))) & 3.5\%\tabularnewline
(((SDE1,(SDE2,SDD)),(SPY1,SPY2)) & 1.5\%\tabularnewline
(((SDE1,SDD),SDE2),(SPY1,SPY2)) & 1.3\%\tabularnewline
((SDE1,SDD),(SDE2,(SPY1,SPY2))) & 1.2\%\tabularnewline
((((SDE1,SDE2),SPY1),SPY2),SDD) & 1.0\%\tabularnewline
((SDE2,SDD),(SDE1,(SPY1,SPY2))) & 1.0\%\tabularnewline
\hline
\end{tabular}
\par\end{centering}
\end{table}
\clearpage{}
%=============================================================================

%=============================================================================
% Table S4. The values of the heat map.
\begin{table}
\caption{\label{tab:heatmap}Ratios of logarithm base 2 of the observed
recombinant edges to the expected number of recombinant edges.  Figure
\ref{fig:Heatmap-of-recombination} shows the values graphically using heat map.
NA's are for the branch pairs that are impossible. Rows are for donor, and
columns for recipient.}
\noindent \centering{}\begin{tabular}{cccccccccc}
\hline
& SDE1 & SDE & SDE2 & SD & SDD & ROOT & SPY1 & SPY & SPY2 \tabularnewline
\hline
SDE1& 0.176  &   NA  &1.170  &    NA & 0.7564  & NA & 0.67 &    NA & 0.663\tabularnewline
SDE & -0.118 & 0.081 &-0.202 &     NA&  0.3395 &  NA& -0.15&  0.045& -0.323\tabularnewline
SDE2& 1.201  &   NA  &0.092  &    NA & 0.6978  & NA & 0.60 &    NA & 0.387\tabularnewline
SD  & -0.629 & 0.022 &-0.661 &0.00670&  0.0078 &  NA& -0.94& -0.299& -1.030\tabularnewline
SDD & -0.317 & 0.371 &-0.197 &     NA&  0.0752 &  NA& -1.06& -0.525& -0.915\tabularnewline
ROOT& -1.745 &-0.698 &-1.786 &0.23039& -0.4854 &  NA& -1.37&  0.497& -1.306\tabularnewline
SPY1& 1.245  &1.883  &1.251  &    NA &-2.1997  & NA & 0.06 &    NA & 1.062\tabularnewline
SPY & 0.089  &0.200  &0.298 &0.00098 &-0.7903  & NA &-0.19 &-0.058 &-0.159\tabularnewline
SPY2& 1.174  &1.938  &1.287  &    NA &-1.9740  & NA & 1.07 &    NA & 0.063\tabularnewline
\hline
\end{tabular}
\end{table}
\clearpage{}
%=============================================================================

%=============================================================================
% Table S5. The values of the heat map of actual counts.
\begin{table}
\caption{\label{tab:obsheatmap}The observed number of recombinant edges.
Rows are for donor, and columns for recipient.}
\noindent \centering{}\begin{tabular}{cccccccccc}
\hline
& SDE1 & SDE & SDE2 & SD & SDD & ROOT & SPY1 & SPY & SPY2 \tabularnewline
\hline
SDE1&8.896 &  0.00 &17.712  & 0.00 & 12.872   & 0 &12.59  &  0.00  &12.53\tabularnewline
SDE &44.626& 136.49& 42.088 &  0.00& 227.061  &  0& 73.39 &  99.95 & 65.25\tabularnewline
SDE2&18.099&   0.00&  8.393 &  0.00&  12.360  &  0& 11.97 &   0.00 & 10.36\tabularnewline
SD  &9.671 &101.74 & 9.464  &31.26 &115.796   & 0 &13.68  & 98.38  &12.86\tabularnewline
SDD &45.539& 168.58& 49.513 &  0.00& 194.706  &  0& 43.09 &  67.14 & 47.65\tabularnewline
ROOT&43.553& 602.67& 42.337 &765.53& 802.813  &  0& 98.86 &2146.01 &103.60\tabularnewline
SPY1&40.800&  11.46& 40.973 &  0.00&   4.371  &  0& 21.15 &   0.00 & 42.35\tabularnewline
SPY &57.618& 262.08& 66.594 & 30.92& 163.518  &  0& 83.56 & 207.00 & 85.10\tabularnewline
SPY2&38.837&  11.90& 42.000 &  0.00&   5.111  &  0& 42.45 &   0.00 & 21.19\tabularnewline
\hline
\end{tabular}
\end{table}
\clearpage{}
%=============================================================================

%=============================================================================
% Table S6. Genes with evident recombining gene transfer.
\begin{table}
\caption{\label{tab:genes-transfer}Genes and their parts transferred via 
homologous
recombining gene transfer. }
{\footnotesize
\begin{tabular}{lllllll}
\hline 
Locus  & N$^a$  & S$^b$  & D$^c$ & F$^d$(\%)  & C$^e$(\%)  & gene 
product\tabularnewline
\hline 
SpyM3\_0074  &  & 4  & 0  & 20  & 100  & putative penicillin-binding protein 1b 
\tabularnewline
SpyM3\_0134  &  & 6  & 0  & 43, 0 & 79  & leucyl-tRNA synthetase \tabularnewline
SpyM3\_0156  &  & 0  & 4  & 77  & 100  & glucose-6-phosphate isomerase 
\tabularnewline
SpyM3\_0186  &  & 5  & 4  & 83  & 94  & putative glucose kinase \tabularnewline
SpyM3\_0192  & 1 & 6  & 5  & 14, 6 & 100  & ribulose-phosphate 3-epimerase 
\tabularnewline
SpyM3\_0193  & 1 & 6  & 5  & 99  & 100  & hypothetical protein\tabularnewline
SpyM3\_0197  &  & 6  & 5  & 7  & 100  & putative surface exclusion 
protein\tabularnewline
SpyM3\_0204  &  & 6  & 5  & 45  & 100  & hypothetical protein \tabularnewline
SpyM3\_0215  &  & 6  & 5  & 5, 26 & 100  & oligopeptide permease \tabularnewline
SpyM3\_0238  &  & 5  & 3  & 18  & 100  & hypothetical protein \tabularnewline
SpyM3\_0292  &  & 6  & 5  & 27  & 100  & putative tetrapyrrole methylase family 
protein \tabularnewline
SpyM3\_0429  &  & 2  & 1  & 0,22  & 100  & putative oligoendopeptidase 
F\tabularnewline
SpyM3\_0465  & 2 & 6  & 5  & 95  & 99  & putative dipeptidase \tabularnewline
SpyM3\_0466  & 2 & 6  & 5  & 16  & 100  & putative adhesion protein 
\tabularnewline
SpyM3\_0506  &  & 5  & 3  & 52  & 100  & phenylalanyl-tRNA synthetase subunit 
alpha \tabularnewline
SpyM3\_0580  & 3 & 3  & 0  & 3  & 100  & PTS system, fructose-specific IIABC 
component\tabularnewline
SpyM3\_0581  & 3 & 3  & 0  & 18  & 100  & putative peptidoglycan hydrolase 
\tabularnewline
SpyM3\_0581  &  & 6  & 5  & 11  & 100  & putative peptidoglycan hydrolase 
\tabularnewline
SpyM3\_0745  & 4 & 6  & 1  & 14  & 100  & putative two-component sensor 
histidine kinase \tabularnewline
SpyM3\_0746  & 4 & 6  & 1  & 55  & 100  & putative two-component response 
regulator \tabularnewline
SpyM3\_0785  & 5 & 4  & 1  & 78  & 92  & inorganic polyphosphate/ATP-NAD 
kinase\tabularnewline
SpyM3\_0786  & 5 & 4  & 1  & 100  & 100  & putative ribosomal large subunit 
pseudouridine synthase \tabularnewline
SpyM3\_0787  & 5 & 4  & 1  & 12  & 100  & phosphotransacetylase \tabularnewline
SpyM3\_1078  & 6 & 4  & 1  & 21  & 100  & putative cation/potassium uptake 
protein \tabularnewline
SpyM3\_1079  & 6 & 4  & 1  & 20  & 100  & putative ATP-dependent RNA helicase 
\tabularnewline
SpyM3\_1093  &  & 4  & 0  & 30  & 100  & putative heavy 
metal/cadmium-transporting ATPase \tabularnewline
SpyM3\_1151  &  & 3  & 1  & 34  & 100  & putative DNA repair and genetic 
recombination protein \tabularnewline
SpyM3\_1153  & 7 & 4  & 0  & 57  & 100  & putative hemolysin\tabularnewline
SpyM3\_1154  & 7 & 4  & 0  & 4  & 100  & putative geranyltranstransferase 
\tabularnewline
SpyM3\_1201  &  & 6  & 5  & 21  & 100  & putative two-component sensor 
histidine kinase \tabularnewline
SpyM3\_1267  & 8 & 6  & 5  & 7  & 100  & bifunctional methionine sulfoxide 
reductase A/B protein \tabularnewline
SpyM3\_1268  & 8 & 6  & 5  & 8  & 100  & hypothetical protein \tabularnewline
SpyM3\_1365  &  & 3  & 0  & 28  & 100  & putative Cof family 
protein/peptidyl-prolyl cis-trans isomerase, cyclophilin
type\tabularnewline
SpyM3\_1380  & 9 & 6  & 1  & 98  & 100  & putative Acetyl-CoA:acetoacetyl-CoA 
transferase b subunit \tabularnewline
SpyM3\_1381  & 9 & 6  & 1  & 100  & 100  & putative oxidoreductase 
\tabularnewline
SpyM3\_1382  & 9 & 6  & 1  & 23  & 38  & 3-hydroxybutyrate dehydrogenase 
\tabularnewline
SpyM3\_1512  &  & 4  & 0  & 59  & 100  & PTS system, mannose-specific IIC 
component \tabularnewline
SpyM3\_1518  & 10 & 1  & 2  & 63  & 100  & acetyl-CoA carboxylase subunit 
beta\tabularnewline
SpyM3\_1519  & 10 & 1  & 2  & 100  & 100  & acetyl-CoA carboxylase biotin 
carboxylase subunit \tabularnewline
SpyM3\_1520  & 10 & 1  & 2  & 100  & 100  & (3R)-hydroxymyristoyl-ACP 
dehydratase \tabularnewline
SpyM3\_1521  & 10 & 1  & 2  & 100  & 100  & acetyl-CoA carboxylase biotin 
carboxyl carrier protein subunit\tabularnewline
SpyM3\_1522  & 10 & 1  & 2  & 100  & 100  & 3-oxoacyl-(acyl carrier protein) 
synthase II \tabularnewline
SpyM3\_1523  & 10 & 1  & 2  & 100  & 100  & 3-ketoacyl-(acyl-carrier-protein) 
reductase \tabularnewline
SpyM3\_1524  & 10 & 1  & 2  & 100  & 100  & acyl-carrier-protein 
S-malonyltransferase\tabularnewline
SpyM3\_1525  & 10 & 1  & 2  & 100  & 100  & putative trans-2-enoyl-ACP 
reductase II \tabularnewline
SpyM3\_1561  &  & 6  & 5  & 11  & 97  & hypothetical protein \tabularnewline
SpyM3\_1577  &  & 4  & 1  & 11  & 100  & excinuclease ABC subunit A 
\tabularnewline
SpyM3\_1579  &  & 4  & 1  & 1  & 100  & hypothetical protein \tabularnewline
SpyM3\_1591  &  & 6  & 5  & 5  & 100  & ribonuclease HIII \tabularnewline
SpyM3\_1753  &  & 6  & 5  & 98  & 100  & putative transcriptional regulator 
\tabularnewline
SpyM3\_1753  & 11 & 3  & 1  & 53  & 100  & putative transcriptional regulator 
\tabularnewline
SpyM3\_1754  & 11 & 3  & 1  & 54  & 100  & putative transcriptional regulator 
\tabularnewline
SpyM3\_1754  &  & 6  & 5  & 95  & 100  & putative transcriptional regulator 
\tabularnewline
SpyM3\_1778  &  & 6  & 0  & 34  & 100  & putative cationic amino acid 
transporter protein \tabularnewline
SpyM3\_1809  &  & 4  & 0  & 5  & 100  & arginyl-tRNA synthetase \tabularnewline
SpyM3\_1813  & 12 & 3  & 0  & 19  & 100  & hypothetical protein \tabularnewline
SpyM3\_1814  & 12 & 3  & 0  & 29  & 100  & aspartyl-tRNA 
synthetase\tabularnewline
\hline
\end{tabular}

$^a$ Numbers represent a group of loci that are consecutive in gene order. Empty
cells denote stand alone loci,

$^b$ Donor branches of homologous recombinant edges, 0 (SDE1), 1 (SDE2), 2
(SDD), 3 (SPY1), 4 (SPY2), 5 (SDE), 6 (SPY), 7(SD), and 8(ROOT),

$^c$ Recipient branches of recombinant edges,

$^d$ Proportion of sites that the recombinant edge covers out of the sites in
the core alignment block, and

$^e$ Percentage of the sites in the core alignment block out of total sites of
the locus.
}

\end{table}
\clearpage{}
%=============================================================================

%=============================================================================
% Table S. ClonalOrigin's 1st stage result.
\begin{table}
\caption{\label{tab:three}Three population parameter estimates from
ClonalOrigin. 
The estimates of the three parameters were based on one of two independent
MCMC chains.  We considered the similar 
results of the two independent chains as confirmation of 
the convergence of the first stage of \texttt{ClonalOrigin} MCMC.
}
\noindent \begin{centering}
\begin{tabular}{ccc}
& Replicate 1 & Replicate 2\tabularnewline
Parameter & Estimate (IQR$^a$) & Estimate (IQR)\tabularnewline
\hline
$\theta^b$ & 0.081 (0.067, 0.094) & 0.081 (0.067, 0.099)\tabularnewline
$\rho^c$ & 0.012 (0.006, 0.019) & 0.012 (0.006, 0.020)\tabularnewline
$\delta^d$ & 744 (346, 2848) & 723 (348, 2870)\tabularnewline
\hline
\end{tabular}
\par\end{centering}

$^a$ IQR stands for interquartile range.

$^b$ The median of mutation rate per site weighted over the lengths of blocks.

$^c$ The median of recombination rate per site weighted over the lengths of blocks.

$^d$ The median value for the recombinant tract length weighted over the lengths of blocks.

\end{table}
\clearpage{}
%=============================================================================

%=============================================================================
% Table SXXX. Simulation study for the three population parameters.
\begin{table}
\caption{\label{tab:sim-three-population-parameters}
Simulation study of the three population parameters.}
\noindent \begin{centering}
\begin{tabular}{ccc}
 & True & Estimates (Standard Deviation)\tabularnewline
\hline
Mutation rate per site & 0.081 & 0.081 (0.0011)\tabularnewline
Recombination rate per site & 0.012 & 0.011 (0.0003)\tabularnewline
Recombinant tract length & 744 bp & 919 bp (28)\tabularnewline
\hline
\end{tabular}
\par\end{centering}
\end{table}
\clearpage{}
%=============================================================================

\begin{table}
\caption{\label{tab:functional-all}Functional categories asscoiated with 
recombination intensity based on all of the recombinant edge types.}
\begin{tabular}{cccc}
\hline 
P-Value & Count & GO term & Description\tabularnewline
\hline 
4.596e-06 & 26 & GO:0019843 & rRNA binding\tabularnewline
5.395e-06 & 40 & GO:0030529 & ribonucleoprotein complex\tabularnewline
7.935e-06 & 39 & GO:0003735 & structural constituent of ribosome\tabularnewline
1.546e-05 & 81 & GO:0006412 & translation\tabularnewline
3.518e-05 & 66 & GO:0005840 & ribosome\tabularnewline
4.288e-05 & 22 & GO:0006633 & fatty acid biosynthetic process\tabularnewline
9.098e-05 & 20 & GO:0008610 & lipid biosynthetic process\tabularnewline
0.0003186 & 33 & GO:0008982 & protein-N(PI)-phosphohistidine-sugar phosphotransferase activity\tabularnewline
0.000357 & 45 & GO:0008643 & carbohydrate transport\tabularnewline
0.0006318 & 38 & GO:0009401 & phosphoenolpyruvate-dependent sugar phosphotransferase system\tabularnewline
0.0009886 & 105 & GO:0003723 & RNA binding\tabularnewline
\hline 
\end{tabular}
\end{table}
\clearpage{}

%
\begin{table}
\caption{\label{tab:functional-notopology}Functional categories asscoiated with 
recombination intensity based on the topology-not-changing recombinant edge types.}
\begin{tabular}{cccc}
\hline 
P-Value & Count & GO term & Description\tabularnewline
\hline 
6.515e-05 & 39 & GO:0003735 & structural constituent of ribosome\tabularnewline
0.00020986 & 40 & GO:0030529 & ribonucleoprotein complex\tabularnewline
0.00045049 & 66 & GO:0005840 & ribosome\tabularnewline
0.00052925 & 26 & GO:0019843 & rRNA binding\tabularnewline
\hline 
\end{tabular}
\end{table}
\clearpage{}

\begin{table}
\caption{\label{tab:functional-sde-spy}Functional categories asscoiated with 
recombination intensity based on SDE-to-SPY recombinant edge types.}
\begin{tabular}{cccc}
\hline 
P-Value & Count & GO term & Description\tabularnewline
\hline 
1.854e-05 & 38 & GO:0009401 & phosphoenolpyruvate-dependent sugar phosphotransferase system\tabularnewline
2.703e-05 & 81 & GO:0006412 & translation\tabularnewline
6.482e-05 & 33 & GO:0008982 & protein-N(PI)-phosphohistidine-sugar phosphotransferase activity\tabularnewline
9.777e-05 & 105 & GO:0003723 & RNA binding\tabularnewline
0.00012793 & 40 & GO:0030529 & ribonucleoprotein complex\tabularnewline
0.00025176 & 26 & GO:0019843 & rRNA binding\tabularnewline
0.00031236 & 72 & GO:0005975 & carbohydrate metabolic process\tabularnewline
0.00044756 & 317 & GO:0005737 & cytoplasm\tabularnewline
0.00059675 & 45 & GO:0008643 & carbohydrate transport\tabularnewline
0.00066874 & 39 & GO:0003735 & structural constituent of ribosome\tabularnewline
0.00135304 & 10 & GO:0006098 & pentose-phosphate shunt\tabularnewline
\hline 
\end{tabular}
\end{table}
\clearpage{}

\begin{table}
\caption{\label{tab:functional-spy-sde}Functional categories asscoiated with 
recombination intensity based on SPY-to-SDE recombinant edge types.}
\begin{tabular}{cccc}
\hline 
P-Value & Count & GO term & Description\tabularnewline
\hline 
1.106e-06 & 45 & GO:0008643 & carbohydrate transport\tabularnewline
1.511e-06 & 33 & GO:0008982 & protein-N(PI)-phosphohistidine-sugar phosphotransferase activity\tabularnewline
1.595e-06 & 38 & GO:0009401 & phosphoenolpyruvate-dependent sugar phosphotransferase system\tabularnewline
1.443e-05 & 187 & GO:0006810 & transport\tabularnewline
0.0001106 & 371 & GO:0016740 & transferase activity\tabularnewline
0.0003596 & 72 & GO:0005975 & carbohydrate metabolic process\tabularnewline
0.0004023 & 13 & GO:0006814 & sodium ion transport\tabularnewline
0.000823 & 317 & GO:0005737 & cytoplasm\tabularnewline
0.001142 & 12 & GO:0031402 & sodium ion binding\tabularnewline
\hline 
\end{tabular}
\end{table}
\clearpage{}

%
\begin{table}
\caption{\label{tab:go-events-duplication}GO-term enrichment for different
evolutionary events - Duplication.}
\begin{tabular}{cccc}
\hline 
P-value & Count & GO-term & Description \\
\hline 
4e-81 &  67 & GO:0006313 & transposition, DNA-mediated\\
5e-80 &  59 & GO:0004803 & transposase activity\\
6e-70 &  23 & GO:0032196 & transposition\\
2e-58 &  52 & GO:0015074 & DNA integration\\
8e-25 &  86 & GO:0006310 & DNA recombination\\
1e-20 &  36 & GO:0016987 & sigma factor activity\\
1e-20 &  36 & GO:0006352 & transcription initiation\\
2e-19 & 197 & GO:0003676 & nucleic acid binding\\
2e-10 & 527 & GO:0003677 & DNA binding\\
4e-06 & 142 & GO:0043565 & sequence-specific DNA binding\\
\hline 
\end{tabular}
\end{table}
\clearpage{}

%
\begin{table}
\caption{\label{tab:go-events-loss}GO-term enrichment for different
evolutionary events - Loss.}
\begin{tabular}{cccc}
\hline 
P-value & Count & GO-term & Description \\
\hline 
1e-06 &  67 & GO:0006313 & transposition, DNA-mediated\\
2e-05 &  59 & GO:0004803 & transposase activity\\
5e-05 &  52 & GO:0015074 & DNA integration\\
5e-04 &  23 & GO:0032196 & transposition\\
5e-04 &  10 & GO:0046873 & metal ion transmembrane transporter activity\\
7e-04 &  36 & GO:0016987 & sigma factor activity\\
7e-04 &  36 & GO:0006352 & transcription initiation\\
1e-03 &  31 & GO:0006814 & sodium ion transport\\
3e-03 &  86 & GO:0006310 & DNA recombination\\
6e-03 & 197 & GO:0003676 & nucleic acid binding\\
\hline 
\end{tabular}
\end{table}
\clearpage{}

%
\begin{table}
\caption{\label{tab:go-events-recombining}GO-term enrichment for different
evolutionary events - Recombining transfers.}
\begin{tabular}{cccc}
\hline 
P-value & Count & GO-term & Description \\
\hline 
7e-09 & 153 & GO:0006412 & translation\\
1e-08 &  67 & GO:0030529 & ribonucleoprotein complex\\
1e-07 &  45 & GO:0019843 & rRNA binding\\
2e-06 &  73 & GO:0003735 & structural constituent of ribosome\\
3e-05 & 157 & GO:0005840 & ribosome\\
7e-05 & 221 & GO:0003723 & RNA binding\\
2e-04 &  11 & GO:0004826 & phenylalanine-tRNA ligase activity\\
3e-04 &  14 & GO:0015413 & nickel-transporting ATPase activity\\
4e-04 &  93 & GO:0008982 & protein-N(PI)-phosphohistidine-sugar phosphotransferase activity\\
5e-04 & 104 & GO:0009401 & phosphoenolpyruvate-dependent sugar phosphotransferase system\\
\hline 
\end{tabular}
\end{table}
\clearpage{}

\section*{Supplementary Figures}

\clearpage{}

%=============================================================================
% Figure S1: Mauve alignment of the five genomes.
\begin{figure}
\begin{center}
\includegraphics[scale=0.25]{figures/mauve}
\end{center}
\caption{\label{fig:mauve}The genome alignment of the five genomes. The top
first and second rows are SDE1 and SDE2 genomes, respectively. The third is SDD
genome. The fourth and fifth rows are SPY1 and SPY2 genomes, respectively.}
\end{figure}
\clearpage{}
%=============================================================================

%=============================================================================
% Figure S2: Species tree
\begin{figure}
\includegraphics[scale=0.5]{figures/cornellf-3-tree}
\caption{\label{fig:tree5}The species tree of the five genomes inferred using
\texttt{ClonalFrame}.  SPY is the branch of a clade of \textit{Streptococcus
pyogenes}, SDE is that for \textit{S.\ dysgalactiae} ssp.\textit{\ equisimilis},
and SD is the branch for the three individual genomes of \textit{S.\
dysgalactiae}. The scale is proportional to the expected number of mutations
given the Watterson's estimate of mutation rate $\theta$ and the tree.}
\end{figure}
\clearpage{}%
%=============================================================================

%=============================================================================
% Figure S3: Histogram of block sizes
\begin{figure}
\begin{center}
\includegraphics[width=5in]{figures/blocksize}
\end{center}
\caption{\label{fig:blocksize}The distribution of core alignment block sizes.}
\end{figure}
\clearpage{}%
%=============================================================================

%=============================================================================
% Figure S4. Three population parameters.
\begin{figure}
\begin{center}
\subfloat[]{\includegraphics[scale=0.5]{figures/scatter-plot-parameter-1\lyxdot out\lyxdot theta}}

\subfloat[]{\includegraphics[scale=0.5]{figures/scatter-plot-parameter-1\lyxdot out\lyxdot rho}}

\subfloat[]{\includegraphics[scale=0.5]{figures/scatter-plot-parameter-1\lyxdot out\lyxdot delta}}
\end{center}
\caption{\label{fig:scatter3}Three scatter plots of mutation rate, recombination
rate, and average tract length for the blocks. The plus signs are mean values
for each block. The density of each value for each block is depicted by black
coulds.  The red dashed lines are the global block-length weighted median of the
blocks.}
\end{figure}
\clearpage{}%
%=============================================================================

%=============================================================================
% Figure S5: Scatter plot of rho and log(delta)
\begin{figure}
\begin{center}
\includegraphics[width=5in]{figures/rho-vs-delta}
\end{center}
\caption{Plot of recombination rate and logarithm of average tract length.}
\label{fig:rhologdelta}
\end{figure}
\clearpage{}%
%=============================================================================

%=============================================================================
% Figure S6. A Recombinant Tree of ClonalOrigin Model
\begin{figure}
\includegraphics[scale=0.42]{figures/mauve-analysis-001}
\caption{\label{fig:clonalorigin}Recombinant tree in the \texttt{ClonalOrigin} model.
(A) A recombinant tree relating the five genomes with four recombinant edges.
(B) A genomic region of size 100 base pairs with subregions affected by the
four recombinant edges. Genomic positions are evenly marked by ticks every 10-th
base pair.  For each edge type and each site, we recorded the sampling frequency
of a recombinant edge of the type at that site.  The posterior probability that
a site experiences influx from a donor branch using a single recombinant tree.
Summing these posteriors over a certain set of edge types, we obtain a measure
of recombination intensity along the genome.  Each segment from b to g are
scored by the number of recombinant edges per site.  (C-G) Local gene trees
induced by the recombinant tree are illustrated for each region covered by
combinations of recombinant edges along the genomic region. For example, the
local gene tree (E) is created by three recombinant edges in the subregion
between 51 base pairs and 70 base pairs. The local gene tree (C) is the same as
the species tree (A) in their topology.  Because there are 105 possible rooted
topologies relating five taxa, we could count topologies.}
\end{figure}
\clearpage{}%
%=============================================================================

%=============================================================================
% Figure S7: recombination heatmap from mowgli
\begin{figure}
\begin{center}
% \includegraphics[width=5in]{figures/trans-recomb-heatmap}
\includegraphics[scale=0.45]{figures/heatmap-recedge-mowgli}
\end{center}
\caption{\label{fig:mowgli-recomb-heatmap}Distribution of recombining horizontal
gene transfers.}
\end{figure}
\clearpage{}
%=============================================================================

%=============================================================================
% Figure S8: gene family sizes
% FIGURE - gene family sizes
\begin{figure}
\begin{center}
\includegraphics[width=5in]{figures/famsizes}
\end{center}
\caption{\label{fig:famsizes}Distribution of gene family sizes.}
\end{figure}
\clearpage{}
%=============================================================================

%=============================================================================
% Figure S9: Comparison of ClonalOrigin and Mowgli analysis
\begin{figure}
\begin{center}
\includegraphics[width=5in]{figures/compareMowgliCoRecombining}
\end{center}
\caption{\label{fig:cmpcomowgli}Plot of recombination intensity against number
of recombining gene transfer.}
\end{figure}
\clearpage{}
%=============================================================================

%=============================================================================
% Figure S10: HGT Comparison of ClonalOrigin and Mowgli analysis
\begin{figure}
\begin{center}
\includegraphics[width=5in]{figures/compareMowgliCoHgt}
\end{center}
\caption{\label{fig:cmpcomowglihgt}Plot of recombination intensity against number
of horizontal gene transfer.}
\end{figure}
\clearpage{}
%=============================================================================

%=============================================================================
% Figure S: View of UCSC genome browser for bacterial gene transfer.
\begin{figure}
\includegraphics[scale=0.8]{figures/ucsc}
\caption{\label{fig:muts}Posterior probability of recombinant edges on 
aspartyl-tRNA
synthetase.  The
top line represents a genomic location of the current view in SPY1 genome. Locus
tags are represented below. The start of a locus tag is red colored, and the end
blue colored. Each of nine horizontal bars represents a branch of the five
species tree. The species tree branch is a recipient species.
The red area on SDE1 line represents gene fragment transfer from
SPY1 due to recombination in locus SpyM3\_1814 of aspartyl-tRNA synthetase.}
\end{figure}
\clearpage{}%
%=============================================================================

%=============================================================================
% Figure S: 
\begin{figure}
\subfloat[]{\includegraphics[scale=0.5]{figures/h1-R}}
\subfloat[]{\includegraphics[scale=0.5]{figures/h1-R-inset}}
\caption{\label{fig:h1}Numer of recombination events. The diagonal dashed
lines are with slope of 1 and intercept of 0. For each order
pairs of branches of a reference species tree the true number of recombination
events is computed as the mean of the 10 replicates of recombinant
trees from the simulation. With ten estimates of each replicate total
number of 100 estimats are shown as intervals of 90\% of quantile
range, or the lower value of the interval is the 5\% quantile, and
the upper one is the 95\% quantile. }
\end{figure}
\clearpage{}%
%=============================================================================

%=============================================================================
% Figure S:
\begin{figure}
\includegraphics{figures/ri1}
\caption{\label{fig:ri1}The surrogate measure of recombination intensity
of genes. Each point represents a pair of a true value of average number of
recombinant edges and its estimated value for a gene. The dashed line is with
slope of 1 and intercept of 0.}
\end{figure}
\clearpage{}%
%=============================================================================

%=============================================================================
% Figure S:
\begin{figure}
\subfloat[]{\includegraphics[scale=0.5]{figures/h2-R}}
\subfloat[]{\includegraphics[scale=0.5]{figures/h2-R-inset}}

\subfloat[]{\includegraphics[scale=0.5]{figures/h2-R-inset2}}

\caption{\label{fig:h2}Comparison of number of recombinations between real
and simulated data sets. The diagonal dashed lines are with
slope of 1 and intercept of 0. For each order pairs of branches of
a reference species tree the number of recombination events is estimated
from the real data set. Total number of 100 estimats from a simulated
data set are shown as intervals of 90\% of quantile range, or the
lower value of the interval is the 5\% quantile, and the upper one
is the 95\% quantile. }
\end{figure}
\clearpage{}%
%=============================================================================

\clearpage

\end{document}

\begin{figure}
\includegraphics[scale=0.7]{figures/plot-number-recombination-within-blocks-1\lyxdot out\lyxdot recomb}
\caption{\label{fig:recombination-boundaries}Distribution of number of 
recombination
events boundaries. The number of recombination event boundaries per
site for each block along the \textit{S.\ dysgalactiae} ssp.\textit{\ equisimilis}
ATCC 12394.}
\end{figure}
\clearpage{}%

The mechanisms of horizontal gene transfer
include conjugation by plasmid, transduction by bacteriophages, or
transformation of naked DNA uptake \citep{Thomas2005}. Additionally,
marine bacteria was reported to acquire foreign genes via gene transfer
agents, i.e., viral-like particles produced by $\alpha$-Proteobacteria
\citep{McDaniel2010}.

Although
recombinant edges between two species branches do not explicitly denote
gene flow between the two species in their original model, \citet{Didelot2010}
exploited their method to infer flows of genetic segment between \textit{Bacillus
cereus} species. This model-based approach to detecting homologous
recombining gene transfer was used to the core genomes of closely
related \textit{Streptococcus} species.

{[}WHERE SHOULD I PUT THESE UNIT OF GENE TRANSFER?{]}

\textbf{Units of gene transfer:} \citet{Chan2009a} showed that protein
domains were not gene transfer units. In another paper \citet{Chan2009}
studied gene with and without breakpoints, and found that more often
entire genes transfer in pathogens than in non-pathogens. In pathogenic
bacteria they found more genes with breakpoints than genes without
breakpoints.

% Table and SDE and SPY
Consistent
with the number of recombinant edges were that gene transfer fragments
with high posterior probability shown in Table S\ref{tab:genes-transfer}
often appeared to be from SPY to SDE. While four genes were inferred
to be transferred to SPY1, there were more genes transferred to SDE1
or SDE2.

\textbf{Evolutionary hypotheses about the ancestor of SPY and SDE:}
\citet{Kalia2001} proposed two scenarios of the evolution of the
two species: The first one was that the most recent common ancestor
of SPY and SDE was a GAS-like pathogen, SDE was losing virulence factors
since the divergence of the ancestor, and SDE regained virulence factors
from SPY because of the same niche. The second scenario was that the
most recent common ancestor of SPY and SDE was a GGS/GCS-like commensal,
SPY evolved virulence factors since the divergence of the ancestor,
and SDE gained virulence factors from SPY. Commensal bacteria could
evolve by aquiring virulence genes from its closely related pathogenic
bacteria. \citet{Kalia2001} suggested the divergence time of SDD
to be around the time when human started to domesticate wildlife animals
around 10,000 years ago. A demographic model could be made with two
species to study the divergence time quantitatively. This would require
ancestral recombination graphs embedded in a species tree.

% From discussion
This distinction of the two processes of
gene transfer was envisioned a decade ago \citep[e.g.,][]{Ochman2001},
and the methods of gene transfer detection by distinguishing the processes
were not yet employed even if some of the methods were already available.
We by no means provided any solutions to the problem of detecting
gene transfer here. Yet, we hoped that we should be able to discriminate
the two processes because each of the processes of gene transfer could
affect biological functions that were different in nature.

% From discussion
For example, genes transferred via non-recombining
gene transfer could be parts of functional units such
as protein domains, and secondary structures of ribosomal RNAs. Genes
transferred via homologous recombining gene transfer could be small
fragments of genes as long as the flanking sequences of the fragments
were more or less homologous.

\textbf{Overlap of genes under positive selection and higher recombination:}
As the uncertainty in gene tree topologies becomes larger, false positive
rates of detecting sites under positive selection would be high (CITE).
Genes could be divided into parts where we hoped no multiple gene
tree topologies exist. Dividing genes into those fragments might be
a way to avoid potentially high false positive rates of positively
selected sites. Considering the fact that nucleotide substitutions
seemed to be affected more by recombinations than by mutations we
suspect that fragmenting genes might not be helpful.

% Virulence gene and non-recombining gene transfer
We investigated bacterial gene
transfer between SPY and SDE using the genome sequences by comparing
gene tree and species tree. We considered both core and pan genomes
for detecting homologous recombining gene transfer and non-recombining
gene transfer. 
%
The extent to which genes were transferred would vary on this
mode of gene transfer. 
While homologous
recombination gene transfer would happen in the genes whose donor
and recipient species were rather closely related, non-homologous
gene transfer could happen virtually in any genes of donor
and recipient species as long as the species shared their habitats.
%
We did not find an excess of virulence
genes associated homologous recombining gene transfer. Actually, many important
virulence genes were missing in the core genomes. We found association
of virulence genes with non-recombining gene transfer.

% Number of recombinant edges (Methods)
This would indicate the degree to which the clonal
species tree was disrupted by bacterial recombination. 
%
We used the method to infer which types of bacterial recombination
were more prevalent than which other types in the lineages of the
tree relating five streptococcus genomes.

% Watterson's estimate
in that the rate of mutation
on the branches of the genealogy and branch lengths of the species
tree $\mathcal{T}$ are reciprocally proportional.

% simulation for ClonalOrigin stage 1
We hoped that we could understand the power
and limitations of \texttt{ClonalOrigin} under the parameter setup that we
inferred from our data set.

% Heat map for recombination from ClonalOrigin
Because the
lengths of SPY1 and SPY2 were larger than those of SDE1 and SDE2,
recombinant edges could more often depart from one of SPY1 and SPY2
lineages and arrive at one of SDE1 and SDE2 lineages than those for
the reverse direction \textit{a priori} (Figure S\ref{fig:tree5}). 
We took into account of the prior expected number of recombinant edges.

\textbf{Gene transfer and functional categories:} We showed much clearer
examples of lateral transfer of gene fragments.
Modular structures of aaRS genes made them
easy targets of gene transfer \citep{Woese2000}. The two transporter
genes might have been under selective pressure from environment.

% rRNA binding
Of the genes in Table S\ref{tab:genes-transfer} we did not find ribosomal
proteins that were heavily enriched in the gene ontology term, {}``rRNA
binding.'' 
Ribosomal proteins tended to cluster together in the genome,
and they appeared to be affected by the recombinant edges whose departures
were on the root branch. 
Recombinant edges that connect the root branch
are hard to be considered as gene flow. We ruled out those recombinant
edges in Table S\ref{tab:genes-transfer} as we mentioned.

% From Introduction
It has been suspected that opportunistic pathogenicity
of SDE may have been facilitated by the very genetic interaction between
the two species. 
Although virulence genes were suspected to be acquired via non-recombining gene
transfer rather than recombining transfer \citep{Groisman1994}, systematic
approaches to elucidating relative effects of the two types of gene transfer 
rarely have been applied.

This unfortunate situation motivated us to study the
closely related bacteria as efforts of better understanding relative 
association of gene transfer with virulence genes.
We hoped that understanding their genetic exchanges
would enhance our knowledge of mechanisms of their pathogenicity on
humans.

While compositional approaches were effective in analyzing
genome-scale data, phylognetic approaches have allowed to understand
evolutionary forces of gene exchange between species
\citep[e.g.,][]{Bapteste2004,Simonson2005,Beiko2009} by comparing differences
between gene trees and a reference species tree.  However, many of the methods
did not have explicit models of homologous recombining gene transfer that was
one of important evolutionary processes in horizontal gene transfer.  
% Overall,
% the development of methods for downstream analyses using bacterial genomes
% appears to lag behind that of high-throughput sequencing technology
% \citep[e.g.,][]{Green2011}.

% Heat map
% Another useful
% summary suggested by \citet{Didelot2010} was to classify recombinant edges by
% types defined according to donor/recipient branch pairs, and record the number
% of edges sampled for each recombinant edge type. For instance, our species tree
% has nine branches including rooting branch and 81 branch pairs, out of which 60
% are possible donor/recipient pairs by prohibiting recombinant edges from going
% backward in time. The posterior expected number of edges of each type can be
% compared to the prior expected number determined by the three population
% parameters fixed in the second stage of \texttt{ClonalOrigin}. This comparison
% was displayed in a heat map of Figure \ref{fig:Heatmap-of-recombination}, which
% indicated branch pairs that might experience an excess or depletion of
% recombination events across all blocks.

% \textbf{Functional categories associated with homologous recombination gene transfer:} 
% Having established relative correlation of the surrogate measure of recombination
% intensity in the simulation (Figure S\ref{fig:ri1}) we attempted to find
% functional categories that were associated with recombination.
% The nucleotide substitutions
% due to recombination events might not be selectively neutral because
% of large population of the bacteria, and accordingly natural selection
% could be more stronger than random genetic drift. Reflecting on the
% report of \citet{Lefebure2007} that genes that were likely to be
% under positive selection tended to be under higher recombination,
% we would expect that functional categories would be associated with higher
% recombination, and consequently homologous recombining gene transfer
% between species.

% block 76: SDEG\_0641
% ORTHOMCL43-1    SEQ\_1497        SpyM3\_0440      MGAS10750\_Spy0535 SDD27957\_03210  SDE12394\_03445  SDEG\_0641
% Not found in Davies2007
% putative calcium transporter
% MgtA

% block 97: SDEG\_0673
% ORTHOMCL1244-1  SEQ\_0860        SDD27957\_03385  MGAS10750\_Spy0623 SpyM3\_0465      SDE12394\_03650  SDEG\_0673
% Putative (put.) extracellular proteins: put. dipeptidase (SpyM3\_0465)
% Dipeptiase for the same human host might be important for SDE to survive human
% host environment. Scarvenge dipeptides.

% block 101: SDEG\_0739
% ORTHOMCL1207-1  SDD27957\_04880  SEQ\_0932        SDD27957\_03660  SDEG\_0739 SDE12394\_03975  MGAS10750\_Spy0677       SpyM3\_0511

% block 151: SDEG\_1447
% ORTHOMCL99-1	SEQ\_0680	SDD27957\_06980	SDEG\_1447	SDE12394\_07540	SpyM3\_1071	MGAS10750\_Spy1252
% Not found in Davies2007
% superoxide dismutase
% sodM
% Poyart2001
% Contribution of Mn-Cofactored Superoxide Dismutase (SodA) to the Virulence of Streptococcus agalactiae.

% block 244: SDEG\_2040
% ORTHOMCL37-1	SEQ\_0209	SpyM3\_1765	MGAS10750\_Spy1855	SDD27957\_10625	SDE12394\_10325	SDEG\_2040
% Not found in Davies2007
% chaperonin GroEL, 
% putative heat shock protein
% What is the role of this GroEL?
% Increase of GroEL in acid shock, which might be important for survival of the
% bacteria in human hosts.

% block 242: SDD27957\_00955
% ORTHOMCL730-1	SDE12394\_10000	SEQ\_2112	SDD27957\_00955	SDE12394\_00680	SDEG\_0169	SpyM3\_1755	MGAS10750\_Spy1843
% Not found in Davies2007
% putative pyruvate formate-lyase activating enzyme
% There are some upstream parts were replaced.

% block 240: SDD27957\_00985
% ORTHOMCL731-1	SDE12394\_10005	SEQ\_2106	SDD27957\_00985	SpyM3\_1749	MGAS10750\_Spy1837	SDE12394\_00710	SDEG\_0175
% Not found in Davies2007
% putative pyruvate formate-lyase
% pflD
% Yesilkaya2009
% Pyruvate Formate Lyase Is Required for Pneumococcal Fermentative Metabolism and Virulence
% SDE2 received from SPY1 via recombining gene transfer.

% block 167: SDD27957\_06790
% ORTHOMCL1139-1	SEQ\_0705	SpyM3\_1035	MGAS10750\_Spy1208	SDD27957\_06790	SDE12394\_07350	SDEG\_1372
% Not found in Davies2007
% putative internalin A precursor
% Adhesins

% block 255: MGAS10750\_Spy1922
% ORTHOMCL158-1	SEQ\_2161	SDD27957\_10925	SDE12394\_10560	SDEG\_2091	SpyM3\_1806	MGAS10750\_Spy1922
% Not found in Davies2007
% DNA mismatch repair protein MutS
% What makes MutS as a virulence genes?

% block 67: SDD27957\_03085
% ORTHOMCL121-1	SEQ\_1525	SDD27957\_03085	SpyM3\_0420	SDE12394\_03265	SDEG\_0610	MGAS10750\_Spy0514
% Not found in Davies2007
% lysyl-tRNA synthetase

% block 133: SDD27957\_04505
% ORTHOMCL58-1	SEQ\_1129	SDD27957\_04505	SDE12394\_04895	SDEG\_0866	SpyM3\_0764	MGAS10750\_Spy0977
% Not found in Davies2007
% putative spermidine/putrescine ABC transporter ATP-binding protein

% List of virulence genes.
% Genes for the Majority Of Group A Streptococcal Virulence Factors and
% Extracellular Surface Proteins Do Not Confer an Increased Propensity to Cause
% Invasive Disease

% block 170: SDD27957\_07725
% ORTHOMCL96-1	SEQ\_0613	SDD27957\_07725	SDEG\_1515	SDE12394\_07910	SpyM3\_1180	MGAS10750\_Spy1365
% Not found in Davies2007
% putative glucose kinase

% Evidence for Reductive Genome Evolution and Lateral Acquisition of Virulence
% Functions in Two Corynebacterium pseudotuberculosis Strains

% block 186: MGAS10750\_Spy1444
% ORTHOMCL93-1	SEQ\_1774	SDD27957\_08550	SDE12394\_08460	SDEG\_1691	SpyM3\_1369	MGAS10750\_Spy1444
% Not found in Davies2007
% putative serine/threonine kinase

% Blocks with virulence genes.
% Check virulence genes (they were combined) at 
% SDD27957\_02890, SDD27957\_02895
% MGAS10750\_Spy0455, MGAS10750\_Spy0456 (block 75)

% HGT and virulene genes in detail
% ORTHOMCL4-1	0	1	6	0	TRUE
% ORTHOMCL4-1	SDD27957_02925	SEQ_0563	MGAS10750_Spy0339	SpyM3_0298	MGAS10750_Spy0340	SDEG_0469	SDD27957_02245	SDE12394_02260	SDD27957_05495	SDEG_0933	SpyM3_1726	MGAS10750_Spy1807
% streptococcal C5a peptidase (scpB) (SDEG_0933) 
% 
% ORTHOMCL1826-1	0	2	6	0	TRUE
% ORTHOMCL1826-1	SDEG_1261	SDE12394_05305	SDE12394_03405	SDEG_0634	SDD27957_00125	SDD27957_02625	SDD27957_04790	SDD27957_05475	SDE12394_02555
% Nothing in Shimomura.
% 
% ORTHOMCL912-1	3	2	5	0	TRUE
% ORTHOMCL912-1	SDD27957_04815	SDD27957_05095	SDD27957_02300	SDD27957_11185	SDD27957_05765	SDD27957_06935	SpyM3_1053	MGAS10750_Spy1189	MGAS10750_Spy1234	SpyM3_1016	SDD27957_02260	MGAS10750_Spy0401	SpyM3_1054	MGAS10750_Spy1235
% Nothing in Shimomura.
% SpyM3_1053 /product="putative transposase"
% SpyM3_1054 /product="putative transposase"
% SpyM3_1016 /product="IS1076 transposase"
% 
% # ORTHOMCL4-2	1	1	4	0	TRUE
% 
% # ORTHOMCL29-2	0	0	4	0	TRUE
% ORTHOMCL29-1	1	1	3	1	TRUE
% ORTHOMCL29-1	SEQ_1712	SEQ_0303	SpyM3_0174	MGAS10750_Spy0201	SDD27957_10325	SDE12394_09770	SDEG_1964	SEQ_1095	SDEG_0515	MGAS10750_Spy0403
% Nothing in Shimomura.
% SpyM3_0174 /product="putative response regulator"
% 
% # ORTHOMCL26-2	0	0	3	0	TRUE
% ORTHOMCL26-1	0	0	3	0	TRUE
% ORTHOMCL26-1	SpyM3_0982	MGAS10750_Spy1149	SEQ_1437	SDD27957_06480	SDEG_1294	SDE12394_06795	SEQ_2022	SDE12394_00945	SDEG_0221
% Nothing in Shimomura.
% SpyM3_0982 /product="putative maltose operon transcriptional repressor"
% 
% # ORTHOMCL1564-2	0	2	3	0	TRUE
% 
% #ORTHOMCL126-2	0	1	3	0	TRUE
% ORTHOMCL126-1	0	1	3	0	TRUE
% ORTHOMCL126-1	SEQ_0301	SDD27957_10335	SDEG_1966	SpyM3_0171	MGAS10750_Spy0199	SEQ_1711	MGAS10750_Spy0404	SDE12394_02570	SDEG_0516
% Nothing in Shimomura.
% SpyM3_0171 /product="putative histidine kinase"
% 
% ORTHOMCL1071-2	1	2	3	2	TRUE
% % ORTHOMCL1071-2	SEQ_0845	SEQ_0190	SEQ_1729	MGAS10750_Spy0620	SpyM3_0731	SpyM3_0922	SpyM3_1208	SDEG_1106	SDD27957_07165
% Nothing in Shimomura.
% SpyM3_0731 /product="putative holin - phage associated"
% SpyM3_0922 /product="putative cell wall hydrolase - phage associated"
% SpyM3_1208 /product="putative cell wall hydrolase, lysin - phage associated"
% 
% ORTHOMCL1037-1	0	2	3	0	TRUE
% ORTHOMCL1037-1	MGAS10750_Spy1697	SDE12394_04580	MGAS10750_Spy0948	SDD27957_05395	SDD27957_04295	SDE12394_07035
% Nothing in Shimomura.
% 
% ORTHOMCL2721	0	4	2	0	TRUE
% ORTHOMCL2721	SDE12394_10035	SDE12394_09945	SDE12394_09940
% Nothing in Shimomura.
% 
% # ORTHOMCL1826-2	4	0	2	0	TRUE
% # ORTHOMCL1826-2	SDEG_1261	SDE12394_05305	SDE12394_03405	SDEG_0634	SDD27957_00125	SDD27957_02625	SDD27957_04790	SDD27957_05475	SDE12394_02555
% 
% ORTHOMCL1564-1	0	2	2	0	TRUE
% ORTHOMCL1564-1	SDEG_1103	SpyM3_1409	SDD27957_09010	SEQ_0696	SDD27957_02745	SDEG_0541
% GGS_124 possesses 5 genes that encode putative nucleases with a secretion signal peptide: genome-encoded streptodornase (SDEG_0541), others,  prophage-associated deoxyribonuclease (SDEG_1103), all of which are predicted to code for a secretion signal peptide.
% SpyM3_1409 /product="streptodornase (Sdn) - phage associated"
% 
% 
% ORTHOMCL1439-1	0	1	2	0	TRUE
% ORTHOMCL1439-1	SEQ_1606	SEQ_1999	SpyM3_0104	SDE12394_09890	SDD27957_10420	SDE12394_09885	SDEG_1984
% Nothing in Shimomura.
% SpyM3_0104 /product="protein F2 like fibronectin-binding protein"
% 
% ORTHOMCL1369-2	2	1	2	0	TRUE
% ORTHOMCL1369-2	MGAS10750_Spy1683	SDEG_1338	SDEG_1346	SDE12394_07225	SDE12394_04610	SDD27957_04325	SDD27957_05365	SDD27957_08360
% Nothing in Shimomura.
% 
% ORTHOMCL1283-1	3	0	2	0	TRUE
% ORTHOMCL1283-1	SDD27957_00840	SDEG_1543	SDEG_1531	SDE12394_03300	SDEG_0983	SDE12394_05495	SDE12394_08020	SDE12394_07960
% Nothing in Shimomura.
% 
% ORTHOMCL1232-2	0	1	2	0	TRUE
% ORTHOMCL1232-2	SDE12394_10675	SDE12394_10850	SDEG_2142	MGAS10750_Spy1897	SDD27957_00535	SDD27957_02435
% Nothing in Shimomura.
% 
% ORTHOMCL1229-2	0	2	2	0	TRUE
% ORTHOMCL1229-2	SDD27957_03510	SDD27957_05295	SDD27957_08295	SEQ_0753	MGAS10750_Spy1709
% Nothing in Shimomura.
% 
% --------------------------------
% 
% 
% ORTHOMCL9-2	1	1	1	0	TRUE
% ORTHOMCL9-2	SpyM3_0160	MGAS10750_Spy0188	SEQ_0271a	SEQ_0289a	SDD27957_10400	SDE12394_09855	SDEG_1980	SpyM3_1853
% 
% ORTHOMCL912-2	5	1	1	0	TRUE
% ORTHOMCL9-1	1	0	1	0	TRUE
% ORTHOMCL848-2	0	1	1	0	TRUE
%

%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Functional category association
Lateral transfer of transporter related genes are well known.  Functional
categories associated with higher recombining gene transfer were found to
include phosphotransferase systems and ATP-binding cassette (ABC) transporters,
i.e., carbohydrate transport (GO:0008643) (Table \ref{tab:functional}).

We also
found several aminoacyl tRNA synthetases in gene ontology of translation
(GO:0006412) to be significantly associated with higher recombining gene
transfer.

Functional categories concerning ribosomes were also
associated with higher recombining gene transfer. These categories were not
found significantly associated with higher recombining gene transfer when we
considered only gene transfer between SPY and SDE genomes, which implied that
many recombining gene transfer events should be attributed to those of internal
species branches as donors.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% From the discussion about virulence genes and two gene transfer
We suspected that virulence genes were exchanged between SPY and
SDE, but did not find any directionality of the gene flow.
This does not necessarily mean that
virulence genes were not exchanged between SPY and SDE via non-recombining
transfer.  Because the \texttt{Mowgli} analysis just did not reveal unbalanced
non-recombining or recombinding gene transfer, we suspected that \texttt{Mowgli}'s
parsimony might not be powerful in detecting unbalanced gene flow.  We
conjecture the following hypothesis of the evolution of SPY and SDE to support
the non-recombining virulence gene exchange between SPY and SDE.  Pathogenicity
of SDE might have existed before the common ancestor of SPY and SDE.  We
speculated that SPY might have been split off from the common ancestor that was
SDE-like because the population size of SDE was larger than that of SPY (CITE).
After the split, SPY might have been specialized in inhabting human hosts, and
SDE in colonizing veterinary hosts. Recent surge in cases of human diseases
attributed to SDE might be because of secondary contact of SPY and SDE since
their split from their common ancestor.  When SPY thrives in its environment of
human hosts and another relatively close species, SDE, starts to cohabit with
the indigenous species SPY, genetic elements in one of the two species could
benefit another species. If the two species do not compete with each other, then
the increased genetic variation from multiple species would eventually bring
advantage to both of the species.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Functional category association 
\textbf{Two functional categories and recombining gene transfer:}
Among the genes shown in Table S\ref{tab:functional} two classes of proteins
attracted our attentions: aminoacyl tRNA synthetases (aaRS) and
phosphotransferase system (PTS).  Aminoacyl tRNA synthetases charge tRNA with an
amino acid so that amino acids are concatenated from charged tRNA onto a growing
peptides through translation. The aaRS's are characterized by an amino acid that
they charge tRNA with.  This gene family had been a target for horizontal gene
transfer \citep[e.g.,][]{Lamour1994,Woese2000,Koonin2001}, which lent itself to
an example of gene transfer in bacteria.  There were also recent reports of gene
transfer in tRNA aminoacylation \citep{Andam2010,Wang2011}.  In our analysis,
some of them appeared to be transferred from SPY to SDE: leucyl-tRNA synthetase,
aspartyl-tRNA synthetase, and arginyl-tRNA synthetase. Half of phenylalanyl-tRNA
synthetase gene appeared to be transferred in the reverse direction.

Bacterial phosphotransferase system (PTS), also known as phosphoenolpyruvate
(PEP) group translocation, transports sugars such as glucose into the cell where
phosphorylation of the sugar happens. There have been studies of gene transfer
of PTS in bacteria \citep[e.g.,][]{Barabote2005,Zuniga2005}. As the main source
of energy for bacteria, sugars in the bacterial niches must have been acting as
important selective pressure. It is not surprising that PTS sugars differ among
bacterial groups because bacterial species are correlated with their environment
and energy sources theirof.  Table S\ref{tab:functional} shows that two PTS
genes appeared to be transferred from SPY to SDE: {}``PTS system,
mannose-specific IIC component'' and {}``PTS system, fructose-specific IIABC
component.'' Whereas almost half of the former gene was inferred to be affected
by recombinant edges, only 3\% of the latter gene appeared to be affected.

% This agreement
% between the two different methods, applied on somewhat different parts of the
% data, serves as validation for the signal picked up by these separate
% association studies.

% These topology-not-changing recombinant edges tend change only branch lengths.