### 2011-12-08 doc/manuscript.tex
Figures are changed. The manuscript is in the tex file.

% Potential reviewers and e-mail addresses 
% DANIEL FALUSH: d.falush@ucc.ie
% MATTHEW HAHN: mwh@indiana.edu 
% Debra Bessen: debra_bessen@nymc.edu
% Xavier Didelot: xavier.didelot@gmail.com
% Brian Golding: Golding@McMaster.CA

% TODO:
% Figure S\label{fig:calling-transfers} was not referred in the text.(Delete this?) 
% Figure S\label{fig:mowgli-sim} was not referred in the text. (Write comments)
% Table S\label{tab:sim-three} was not referred in the text. (Delete this?)
% 
%
% * Finalize section: Distribution of transfer events along the genome
% * Finalize section: Supp. Materials.
% Supp. Methods of Matt's modifications of Mowgli's method
% 
% Backup of the work
% swiftgen:/usr/projects/strep/gainloss/recombination

\documentclass[12pt]{article}

\usepackage{times}

% amsmath package, useful for mathematical formulas
\usepackage{amsmath}
% amssymb package, useful for mathematical symbols
\usepackage{amssymb}

% graphicx package, useful for including eps and pdf graphics
% include graphics with the command \includegraphics
\usepackage{graphicx}

% cite package, to clean up citations in the main text. Do not remove.
%\usepackage{cite}

\usepackage{color}

% Use doublespacing - comment out for single spacing
\usepackage{setspace}

% To rotate tables using sidewaystable instead of just table.
\usepackage{rotating}

% FIXME: Remove this package when submitting the manuscript.
\usepackage{url}

\usepackage{natbib}
\usepackage{../../latex/sty/genres}

% Text layout
\oddsidemargin 0in
\evensidemargin 0in
\topmargin -.5in
\textwidth 6.5in
\textheight 9in
%\topmargin 0.0cm
%\oddsidemargin 0.5cm
%\evensidemargin 0.5cm
%\textwidth 16cm 
%\textheight 21cm

% Bold the 'Figure #' in the caption and separate it with a period
% Captions will be left justified
\usepackage[labelfont=bf,labelsep=period,justification=raggedright]{caption}

% Remove brackets from numbering in List of References
\makeatletter
\renewcommand{\@biblabel}[1]{\quad#1.}

% FIXME: Remove the followings if you have to when submitting the manuscript.
\providecommand{\tabularnewline}{\\}
\newcommand{\lyxdot}{.}
\newcommand{\comment}[1]{}

% \@ifundefined{showcaptionsetup}{}{%
% \PassOptionsToPackage{caption=false}{subfig}}
% \usepackage{subfig}

\makeatother

% Leave date blank
\date{}

\pagestyle{myheadings}
%% ** EDIT HERE **
\markright{Horizontal gene transfer in Streptococcus}

\begin{document}

\begin{titlepage}

\title{Replacing and additive horizontal \\
gene transfer in {\em Streptococcus} }

\author{Sang Chul Choi$^{1}$, Matthew D.\ Rasmussen$^{1}$, 
Melissa J.\ Hubisz$^{1}$, \\
Ilan Gronau$^{1}$,
Michael J.\ Stanhope$^{2}$, and Adam Siepel$^{1}$}

\date{ }
\maketitle

\begin{footnotesize}
\begin{center}
$^1$Department of Biological Statistics and Computational Biology,\\
Cornell University, Ithaca, NY 14853, USA
\\[1ex]
$^2$Department of Population Medicine and Diagnostic Sciences,\\
College of Veterinary Medicine, Cornell University, Ithaca, NY 14853, USA
\\
\end{center}
\end{footnotesize}

\vspace{1in}

\begin{tabular}{lp{4.5in}}
{\bf Submission type:}& Research Article
\vspace{1ex}\\
{\bf Keywords:}&Bacterial evolutionary
genomics, recombination, {\em Streptococcus pyogenes}, {\em Streptococcus
  dysgalactiae} 
\vspace{1ex}\\
{\bf Running Head:}&Horizontal gene transfer in {\em Streptococcus}
\vspace{1ex}\\ 
{\bf Corresponding Author:}&
\begin{minipage}[t]{4in}
 Adam Siepel\\
 102E Weill Hall, Cornell University\\
 Ithaca, NY 14853\\
 Phone: +1-607-254-1157\\
 Fax: +1-607-255-4698\\
 Email: acs4@cornell.edu
\end{minipage}
\end{tabular}

\thispagestyle{empty}
\end{titlepage}

\doublespacing

% MBE limit is 350 words
\section*{Abstract}

The prominent role of horizontal gene transfer (HGT) in the evolution of
bacteria is now well documented, yet few studies have differentiated
between the mechanisms of homologous recombination and additive integration.
%as sources of HGT.  
These processes leave distinct phylogenetic signatures, because homologous
recombination causes a gene in one lineage to be replaced by a homolog from
another lineage (``replacing HGT''), while additive integration causes
addition without replacement (``additive HGT'').  Here we make use of these
signatures in a genome-wide investigation of HGT in the important human
pathogen {\em Streptococcus pyogenes} (SPY) and its close relatives {\em
  S.\ dysgalactiae} subspecies {\em equisimilis} (SDE) and {\em S.\
  dysgalactiae} subspecies {\em dysgalactiae} (SDD).  Using recently
developed statistical models and computational methods, we find evidence
for abundant gene flow of both kinds within the SPY and SDE clades, and of
generally reduced levels of exchange between SPY/SDE and SDD.  In addition,
our analysis strongly supports a previously reported, but unconfirmed,
finding of a pronounced asymmetry in SPY-SDE gene flow, favoring the
SPY-to-SDE direction, which may be associated with increasing virulence of
pathogenic SDE.  We find that this enrichment is driven by replacing
transfers, and likely reflects frequent homologous recombination between
co-occurring SPY and SDE cells in human hosts.  We find that virulence
genes are correlated with transfer events, but this correlation is driven
by additive, not replacement, HGTs.  The genes affected by additive HGTs
are enriched for transposition-related functions, while replacement HGTs
appear to influence a more diverse set of genes.  These findings shed new
light on the manner in which HGT has shaped pathogenic bacterial genomes
during relatively recent evolutionary time.

\thispagestyle{empty}
\clearpage

\section*{Introduction}

During the last twenty years, it has become increasingly apparent that
horizontal gene transfer (HGT) has played a major role in genomic
evolution, particularly among bacteria and other microbes
\citep{Smith1992,Smith1993,Ochman2000,Koonin2001}.  Indeed, it is likely
that HGT has been sufficiently prevalent throughout evolutionary history
that no present-day gene can trace an unbroken history of vertical descent
to a common ancestor for all species \citep{Zhaxybayeva2011}.  This would
mean that no single phylogenetic marker is a reliable indicator of the tree
of life.

For HGT to occur in bacteria, DNA must first be transferred from a donor to
a recipient cell by transformation, transduction, or conjugation, then must
be integrated into the recipient cell's genome (setting aside the case of
replication-proficient plasmids).  This integration typically occurs in one
of two major ways: either the new sequence replaces a homologous sequence
through the process of homologous recombination (similar to ``gene
conversion'' in sexually reproducing organisms), or it is acquired through
an additive (non-replacing) integration process \citep{Thomas2005}.  These
two types of transfers leave distinct molecular evolutionary signatures and
will be referred to in this article as ``replacing'' and ``additive'' HGTs,
respectively (fig.\ \ref{fig:hgt}).  The rates at which these processes
occur depend on many factors, including the co-occurrence of bacterial
species in the environment,
%the propensity of donor cells to release intact DNA,
the degree of competence of recipient cells, sequence similarity between
the donor and recipient, and barriers to conjugation such as surface
exclusion.  In addition, the persistence of horizontally transferred DNA
segments in present-day species depends strongly on the subsequent effects
of natural selection.  Not surprisingly, rates of HGT appear to vary
considerably across groups of bacteria \citep[e.g.,][]{Feil2001}.

HGT in the {\em Streptococcus} genus is of particular interest.  This group
of Gram-positive bacteria contains the human pathogens {\em S.\ pyogenes}
(the cause of Group A streptococcal infections, including pharyngitis,
impetigo, cellulitis, necrotizing fasciitis, and rheumatic fever), {\em S.\
  pneumoniae} (bacterial pneumonia), {\em S.\ agalactiae} (neonatal sepsis),
{\em S.\ mutans} (dental caries), and {\em S.\ dysgalactiae} ssp.\ {\em
  equisimilis} (cellulitis, peritonitis, pneumonia, and other infections),
as well as the agricultural pathogens {\em S.\ dysgalactiae} ssp.\ {\em
  dysgalactiae} (mastitis in cows, ewes, and goats), {\em S.\ equi} ssp.\
{\em equi} (strangles in horses), and {\em S.\ canis} (various infections
in dogs and other animals).  Many of these species are concentrated in the
predominantly beta-hemolytic {\em Pyogenes} group.  Despite their
relatively recent evolutionary divergence, these {\em Pyogenes}-group
species have adapted to a remarkable variety of distinct ecological niches.
Accordingly, they display high rates of recombination and HGT, as well as
substantial evidence of positive selection
\citep{Feil2001,Marri2006,Anisimova2007,Lefebure2007,Suzuki2011}.  Complete
genome sequences are now available for several species in this group
\citep[e.g.,][]{Ferretti2001,Holden2009,Shimomura2011,Suzuki2011},
providing a rich resource for the study of the genomic evolution of
pathogenic bacteria over relatively recent time scales.

Within the {\em Pyogenes} group, several studies have focused on the issue
of HGT between the human-colonizing species {\em S.\ pyogenes} (here
denoted SPY) and {\em S.\ dysgalactiae} ssp.\ {\em equisimilis} (SDE).
Until recently, SDE was considered to be primarily a commensal organism
\citep{Vandamme1996}, unlike SPY, but it has become increasingly apparent
that it also has an important pathogenetic role, with a disease spectrum
similar to that of SPY \citep{Brandt2009}.  Shared virulence genes are well
documented between the two species \citep[e.g.,][]{Davies2007a}, genetic
exchange presumably having been enabled both by their close evolutionary
relationship and shared ecological niches, but
% Towers et al 2004
% Maxted and Potter, 1967
% Simpson et al., 1992
% Sriprakash et al., 1996
in a widely noted phylogenetic study of seven loci and more than 200
isolates, \cite{Kalia2001} also found extensive evidence of SPY-SDE gene
flow among housekeeping genes.  Interestingly, they observed a pronounced
bias in the direction of gene flow, favoring the SPY-to-SDE direction.
Gene flow has been confirmed in subsequent studies
\citep{Kalia2004,Davies2005,Davies2007a,Davies2007}, but the evidence for
directional asymmetry is mixed, with some indications of gene flow in the
opposite (SDE-to-SPY) direction \citep{Sachse2002}.  This issue has been
further complicated by the eventual retraction of Kalia et al.'s
\citeyearpar{Kalia2001} paper, due to the authors' inability to replicate
their findings and concerns about sample contamination.  The authors argued
that whole-genome sequencing may be required to definitively resolve open
questions about gene flow between these species.  These issues are of
particular interest because of the possibility of increasing virulence of
SDE due, at least in part, to gene flow from SPY.

Most recent studies of gene flow in bacteria have been based on relatively
simple phylogenetic methods that make use of incongruity of inferred gene
trees across loci \citep[e.g.,][]{Kalia2001,Kalia2004,Ahmad2009}.  These
methods are sensitive to errors in phylogenetic reconstruction, and ignore
information from branch lengths and from phylogenetic correlation at
adjacent loci.  Recently alternative, model-based approaches have been
proposed for the case of homologous recombination (``replacing HGTs'')
\citep{Didelot2007,Didelot2010}.  Using Bayesian principles and Markov
chain Monte Carlo (MCMC) techniques, these methods allow for uncertainty in
the phylogeny, make use of branch lengths and correlations across loci, and
allow for inference not only of individual recombination events but also of
parameters describing global rates and patterns of recombination.  In
addition, several recent methods have been introduced to detect HGT in a
phylogenetic framework, by parsimoniously reconciling reconstructed gene
trees with a given species tree, allowing for duplication, transfer, and
loss (DTL) events \citep{Merkle2010,David2011,Doyon2011,Doyon2011b}.  These
new model- and parsimony-based methods are complementary in many respects,
and could be particularly useful in combination.

In this article, we take a fresh look at the issue of HGT in the {\em
  Pyogenes} group, making use of newly available complete genome sequences
for SPY, SDE, and the related species {\em S.\ dysgalactiae} ssp.\ {\em
  dysgalactiae} (SDD), and new statistical and parsimony-based methods for
analysis.  The combined use of these methods allows us to examine both
replacing and additive HGTs, and compare their genome-wide effects.  We
find evidence for abundant gene flow within SPY, within SDE, and between
SPY and SDE, and for greatly reduced exchange between SPY/SDE and SDD.  In
addition, our genome-wide analysis supports Kalia et al.'s
\citeyearpar{Kalia2001} previously unconfirmed finding of a pronounced
preference for the SPY-to-SDE direction in SPY-SDE gene flow.
Interestingly, this property appears to hold for replacing but not for
additive transfers.  More generally, we find evidence for an association
between virulence genes and additive, but not replacing, HGTs.  Our results
have been made available as tracks in our recently released {\em
  Streptococcus} genome browser.

\section*{Methods}

\subsection*{Genome sequences and alignments}

Our primary analysis was based on five complete genome sequences, including
two representatives of {\em S.\ pyogenes} (accessions NC\_004070
[\citealp{Beres2002}; SPY1] and NC\_008024 [\citealp{Beres2006}, SPY2]),
two of {\em S.\ dysgalactiae} ssp.\ {\em equisimilis} (CP002215
[\citealp{Suzuki2011}; SDE1] and NC\_012891 [\citealp{Shimomura2011};
SDE2]), and one of {\em S.\ dysgalactiae} ssp.\ {\em dysgalactiae}
(AEGO00000000 [\citealp{Suzuki2011}; SDD]).  In addition, we used 
\textit{S.\ equi} ssp.\textit{\ equi} strain 4047 (NC\_012471
[\citealp{Holden2009}; SEE]) as an outgroup in our parsimony-based
analysis (supplementary table S\ref{tab:genome}).

For the model-based analysis of replacing transfers, we obtained
high-quality alignments of the five genomes (SPY1 \& 2, SDE1 \& 2, and
SDD) using the 
pipeline detailed by Didelot and colleagues
\citep{Didelot2007,Didelot2010}.  Briefly, we first aligned the genomes
using progressiveMauve v2.3.1 \citep{Darling2004,Darling2010} with default
options (supplementary fig.\ S\ref{fig:mauve}), and then used stripSubsetLCBs (from the
ClonalOrigin package; \citealp{Didelot2010}) to identify 276 blocks of
sequence alignments that exceed a length threshold of 1500 bp.  This
conservative threshold was selected to avoid biases from boundary effects
in short alignment blocks, based on a preliminary analysis.  Two blocks
containing long gaps were excluded from the analysis.
The final set of 274
alignment blocks was 1147632 nucleotide long.
Of 1951 genes of SPY1 genome 1125 genes were covered by the blocks.
Note that this alignment
procedure effectively focused on the ``core genome,'' by discarding regions
that were not conserved among the five species in question.  The 1125 genes
contained 0.17\% of gaps on average.

For the parsimony-based analysis, we began with 2314 gene family
clusters representing the six genomes (including the SEE outgroup)
from \cite{Suzuki2011}.  We aligned protein sequences corresponding to
each family using MUSCLE \citep{Edgar2004a}, and then obtained
nucleotide alignments by reverse translation, using the genomic
sequences as a guide.  Next we identified likely intragenic
recombinations using the Single Breakpoint Recombination (SBP) method
from the HyPhy package \citep{KosakovskyPond2006}.  Based on the
Akaike Information Criterion (AIC), SBP identified at least one
topology-altering recombination in 47.5\% (1094) of gene
families. These alignments were split into putatively non-recombining
gene fragments.  The final set of alignments contained 3408 gene
fragment families, 1794 of which contained one gene in each of the six
species.  Families ranged as small as two genes (417 fragments) to as
large as 59 (2 fragments) (supplementary fig.\ S\ref{fig:famsizes}).  Notably, in
contrast to the genome-wide alignments produced by progressiveMauve,
these alignments were representative of the ``pan genome,'' including
regions that frequently turn over, rather than just the ``core
genome'' \citep[e.g.,][]{Tettelin2005,Lefebure2007}.

\subsection*{Model-based analysis of replacing HGTs}

{\bf ClonalFrame and ClonalOrigin.}  To study replacing HGTs, we made use
of the coalescent-based model of homologous recombination and the Markov
chain Monte Carlo (MCMC) algorithm recently developed by
\cite{Didelot2010}.  The method was implemented in a 
program, called ClonalOrigin, that could produce samples from the posterior
distribution of {\em recombinant trees} given a set of alignment blocks. A
recombinant tree consists of a rooted tree with branch length, called a clonal
frame, for the five bacterial genomes,
which is estimated in a pre-processing step, augmented by a set of
{\em recombinant edges}. Each recombinant edge is defined by the two points
along the branches of clonal frame corresponding to the {\em donor} and
{\em recipient} in the hypothesized recombination event, and by the start
and end of the genomic interval that is affected by the event.  The donor
must pre-date the recipient, and a delayed coalescence is possible
(see \textbf{Supplementary Material} for details of summarizing recombinant
trees and edges). 
Another program, called ClonalFrame
\citep{Didelot2007}, can be used to infer the phylogeny that serves as the
basis of the recombinant trees.  ClonalFrame makes use of a similar, but
somewhat simpler probabilistic model, which permits inference
of the clonal frame. % and a set of model parameters.

{\bf Fitting the model.}  We used
ClonalFrame (v1.1) to estimate 
a clonal frame for the five-way
alignments, running the algorithm for $10^4$ burn-in iterations followed by
another $10^4$ sampling iterations.  A series of seven additional sampling
runs with 
random initializations converged to an essentially identical clonal frame, % model parameters,
indicating stable convergence.  We then used an initial run of
ClonalOrigin (subversion r19) to estimate global parameters including
the mutation rate, recombination rate, and recombinant tract length. We
applied the method separately to each of the 274 blocks, in all cases using
$10^6$ burn-in iterations, $10^7$ sampling iterations, and sub-sampling
every $10^5$ iterations. A global estimate for each parameter was then
obtained by taking the block-length weighted median of the block-specific
posterior means, which down-weighed less reliable estimates obtained from
shorter alignment blocks.  We generally followed \cite{Didelot2010} in 
fitting the model. % , with some minor differences.

{\bf Sampling recombinant trees}.  Next we ran a second stage of ClonalOrigin
MCMC
on all of the alignment blocks by fixing the clonal frame and global model
parameters at their pre-estimated values, and sampled
recombinant trees.  For each block, we ran the sampler for $10^7$ burn-in
iterations and $10^8$ sampling iterations, sub-sampling every $10^5$
iterations, resulting in 1001 recombinant trees per alignment block.  We
performed a replicate of the two stages of the ClonalOrigin inference to
ensure adequate convergence of the sampler.

% A few alignment blocks were not successfully finished even after a month of
% computation. Because those were not expected to affect the global estimates of
% population parameters significantly, we used only finished blocks to estimate
% mutation rate, recombination rate, and recombinant tract length.

{\bf Summary statistics.}  We used various summary statistics to describe
the sampled recombinant trees (see \textbf{Supplementary Material} for details). 
First, we considered the distribution of
tree topologies at each site in the alignments, by computing the site-wise
relative frequencies of all 105 possible rooted tree topologies relating five
taxa. Second, we measured frequencies in the posterior samples
of all possible types of recombinant edge, where an edge
type is defined by the branches of the clonal frame associated with its
donor and recipient.
%also searched the posterior sample for transfer scenarios that were
%significantly enriched, compared to the prior expectation defined by the
%clonal frame and the global estimates of the three population
%parameters. We followed the approach of \cite{Didelot2010} and defined
%transfer types by the branches of the clonal frame associated with each
%recombinant edge. 
%For each transfer type, we recorded the average number of
%recombinant edges of that type sampled in the posterior sample of
%recombinant trees, 
We compared these posterior frequencies with 
expected values under the prior distribution assumed by the model.
%as well as the expected number of such edges implied by
%the prior distribution assumed by the model. 
%These counts were then used to
%find types of transfer that showed a high deviation in our inference from
%their expected counts. 
To assess significance of the ratio of posterior to prior frequencies of
recombinant edges we performed a simulation; 
this empirical approach also 
had the advantage of correcting for possible biases in the sampling
process. 
\textbf{CHECK:}
We simulated the model to generate 100 data sets, for which we proceeded with
the inference of the ratio of posterior to prior frequencies of recombinant
edges. Comparing the obtained null distribution of the ratios from the
simulation and the ratio from the real data analysis, we assessed the
significance of the ratio.
% (GIVE VERY BRIEF SUMMARY OF SIMULATION APPROACH)
%Ilan: point to supplementary methods for the simulation tests?
Third,
we traced the posterior
probabilities of transfer types in a position-specific manner. 
%For each
%distinct type of transfer (determined by a pair of donor and recipient
%edges), and each site in the core genome alignment, we computed the
%posterior probability that that site had been affected by a transfer of
%that type. 
Finally, we defined a general, branch-independent
 {\em recombination intensity} at each site by 
summing the posterior probabilities of recombination.
In some cases, we also computed intensities for particular types of
replacing transfers, such as 
those between the SPY and SDE clades, or those producing topologies
different from the clonal frame, or \textit{topology-altering} transfers.  
% CHECK: Justify why we used only topology-altering edges for computing
% recombination intensity.
Note that \cite{Didelot2010} described the second and third summarization
procedures.

\subsection*{Parsimony-based analysis}

% move stuff down about gene tree estimation

{\bf Mowgli.} To study additive HGTs, we used a parsimony-based method
for reconciling gene and species trees called Mowgli \citep{Doyon2011}.
Given a gene tree and a species tree, Mowgli finds a reconciliation scenario 
that minimizes the number of gene duplication, loss, and
transfer (DLT) events by considering tree topologies with branching orders, also
known as labeled histories \citep{Edwards1970}.  We estimated
unrooted gene trees for the nucleotide alignments for each of our
putatively non-recombining gene fragment using RAxML \citep{Stamatakis2006},
then applied Mowgli to each of these trees together with the species
tree estimated by ClonalFrame (augmented with the SEE outgroup).
Since Mowgli required rooted gene trees, we ran Mowgli for all possible
rootings of each gene tree, and chose the scenario that produced the
minimum DLT score. In the case of multiple most parsimonious
reconciliations, one was chosen uniformly at random.

{\bf Distinguishing between replacing and additive HGTs.}  The transfer
events inferred by Mowgli had to be classified as replacing or
additive in a post-processing step because the program did not
distinguish between these types of events.  A naive approach 
would be to infer a replacing HGT whenever an inferred transfer
coincided exactly with an inferred loss in the recipient lineage.  However,
loss events were difficult to place accurately because they were frequently
pushed toward the root of the tree due to reconstruction error or parallel
losses in multiple lineages.  Therefore, we classified a transfer event as
additive whenever a descendant of the recipient lineage in the species tree
contained both a gene descended from the transferred copy and a gene not
descended from that copy.  All other transfers were considered replacing
HGTs.  This was conservative about calling additive HGTs, and would
label any HGT that could plausibly be explained by replacing transfers
as such (see fig.\ \ref{fig:hgt}).

{\bf Statistical enrichment}. As with replacing
HGTs, we classified transfer events by donor and recipient branch, 
%sought to identify over-represented scenarios of additive HGT. 
%and identified distinct types of additive
%transfers by their donor and recipient branches in the species
%tree, 
and recorded the number of inferred additive HGTs of each type.
%as well as the prior expected number of transfers
%of that type. 
We compared these numbers to expectations based on simulations that
assumed 
constant
rates of duplication, loss, and transfer across the species tree, given 
the total numbers of events inferred across all 2314 gene families
divided by the total branch lengths of the trees
(see \textbf{Supplementary Material}).

%Ilan: add some detail about simulations, or point to supp

\subsection*{Gene category associations and virulence genes}

We performed a series of analyses to search for certain classes of genes
enriched for replacing or additive of gene transfers. In this analysis,
each gene, or gene family, was associated with a certain score. For
replacing gene transfers we used the recombination intensity along
a given gene, and for additive gene transfers we used the number of
additive transfers inferred for a given family. Mann-Whitney tests were
used to identify sets of genes that showed significantly elevated values in
each score.  A 5\% False Discovery Rate (FDR) correction
\citep{Benjamini1995} was used to determine an appropriate significance
threshold for our p-values. We applied the association test for gene
classes defined by Gene Ontology (GO) terms assigned  
%using the method described in Suzuki et al.\ \cite{Suzuki2011}. In short, 
by comparing the Streptococcus genes to bacterial proteins from the
Uniref90 database using blastP, and then assigning the same GO
classification as the target gene of the uniProt GOA database, if the
match had an E-values less than $1.0\times10^{-5}$. We also classified a
category of \textit{virulence genes} using the virulence profile obtained
by \cite{Suzuki2011}. In both cases, a gene family was
associsated with a certain classification if any of its genes were
associated with that classification.

\section*{Results}

\subsection*{Clonal frame and global parameter estimates}
% The 5-species tree was
% (((1:0.045557,2:0.045557)6:0.195330,3:0.240887)8:0.089874,(4:0.074544,5:0.074544)7:0.256218)9:0.000000
By applying ClonalFrame to our five-way genomic alignments 
(supplementary fig.\ S\ref{fig:mauve}), we obtained a
phylogeny in which the pairs of SPY and SDE samples were grouped together,
as expected, and in which SDD and SDE formed a clade, with SPY as an
outgroup (fig.\ \ref{fig:tree5}).  This tree was consistent with previous
results from 16S rRNA sequences \citep{Facklam2002} and with genome-wide
trends \citep{Suzuki2011}.  Despite the grouping of SDD and SDE, their
estimated divergence was substantial, at $\sim$0.48 substitutions per site.  
The estimated divergence between SPY and SDD/SDE was $\sim$0.66
substitutions per site.  
%These estimates correspond to average
%genomic divergence times of roughly 660 and 900 thousand years,
%respectively, assuming a mutation rate of $5 \times 10^{-10}$ substitutions
%per site per generation \citep{Ochman2003} and two cell divisions per day
%\citep{XX}.  
An initial analysis with ClonalOrigin (see \textbf{Methods}) produced an
estimated
per-site population-scaled mutation rate of $\theta = $ 0.081
[interquartile range across blocks: (0.067, 0.094)], a per-site
population-scaled recombination rate of $\rho = $ 0.012 (0.006, 0.019), and
an average recombinant tract length of $\delta = $ 744 (346, 2848) bp.  The
full distributions of these quantities across alignment blocks are shown in
supplementary
figure S\ref{fig:scatter3}.  Our estimates implied a ratio of $\rho/\theta
= 0.15$.  For comparison, \cite{Didelot2010} obtained estimates of $\theta
= 0.044$, $\rho = 0.017$, $\delta = 236$, and $\rho/\theta = 0.405$ for the
{\em Bacillus cereus} group, but differences of this magnitude were not
surprising, given likely differences in effective population sizes,
recombination rates, and mutation rates \citep[e.g.,][]{Feil2001}.  We
found that estimates of the recombination tract length, $\delta$, were
quite sensitive to our threshold for minimum block length, with the
inclusion of short alignments producing much larger estimates due to edge
effects.  We experimented with a range of thresholds and selected one (1500
bp) at which the estimates stabilized.

\subsection*{Model-based analysis of replacing HGTs}

With a second stage of ClonalOrigin MCMC, we were able to obtain samples
from the posterior distribution of recombinant trees (i.e., the clonal
frame augmented by recombinant edges) along the genome, conditional on our
five-way alignments and the previously estimated parameters (see \textbf{Methods}).
This distribution indicated that most sites (66\%) were
consistent with the topology of the clonal frame.
% what does this mean?  most likely tree topology per site?  or expected value?
Interestingly, the second most frequent tree topology (9.3\%) had
SPY and SDE as sister clades and SDD as an outgroup (supplementary table
S\ref{tab:Gene-tree-topologies}), providing an initial indication of gene
transfer between SPY and SDE.  No other tree topology appeared with
appreciable frequency.
% This indicated that SPY and SDE appeared to be more closely related with each
% other than each to SDD's genome in about tenth of the core genome. The other
% topologies had less than 4\% of proportions.

% eliminate some redundancy with methods
To gain further insight into rates and patterns of HGT, we estimated the
rates of occurrence of recombinant edges, grouping them by the associated
donor and recipient edges in the clonal frame (see \textbf{Methods}).  Because donor
edges must pre-date recipient edges under the model, 21 of the 81 edge
pairs for the nine-branch rooted topology are prohibited, leaving 60
possible recombinant edge types, each corresponding to
a class of replacing HGTs.  
% Because of the possibility of delayed
% coalescence, these classes include not only transfers from the donor to the
% recipient edge in the clonal frame, but also transfers from any descendant
% of the donor to the recipient (with the implicit constraint of
% contemporaneous donors and recipients).  Thus, recombinant edges are
% informative about replacing HGTs, but with some ambiguity.  
Because recombinant edges are not all equally likely \textit{a priori}, 
we summarized the estimated
rates by taking ratios with respect to their prior expectations under the
model used by ClonalOrigin, which consider the branch lengths of the clonal
frame.  Experiments conducted on simulated data demonstrated that our
approach could accurately detect recombination scenarios that deviate from
the prior, with some underestimation of enrichment levels 
(see \textbf{Supplementary Material}).

These normalized rates for our alignments can be summarized in a $9 \times
9$ heatmap, consisting of 60 informative and 21 undefined cells (fig.\
\ref{fig:Heatmap-of-transfers}A; see also supplementary tables
S\ref{tab:heatmap} and S\ref{tab:obsheatmap}).  Most cells in this heatmap
fell in the blue-to-white range, corresponding to rates of occurrence less
than or equal to what was expected under the prior.  Some of these
correspond to edges that were fairly strongly under-represented (blue),
such as those between the SDD branch and the branches of the SPY clade
(P-value of $<0.01$, based on an empirical test; see \textbf{Methods}),
possibly reflecting ecological isolation of human and strictly veterinary
pathogens.  By contrast, the recombination edges between the two SDE
genomes, and between the two SPY genomes, were significantly
over-represented (P-value of $<0.01$), consistent with previous evidence
for intraspecies recombination in {\em Streptococcus} \citep{Feil2001}
(CHECK).  In addition, a pronounced enrichment was observed in
recombination edges between the SPY and SDE clades, as had been reported
previously \citep{Kalia2001,Sachse2002,Kalia2004,Davies2005,Davies2007a}.
While these edges occurred at elevated rates in both directions, we found a
clear asymmetry, with edges in the SPY-to-SDE direction occuring at up to
four times the expected rate, while those in the reverse direction occur at
only about twice the expected rate.  \texttt{CHECK:} To directly compare
the gene flow of SPY-to-SDE with that of SDE-to-SPY for each of 1001
recombinant trees from the posterior distribution we summed numbers of
recombinant edges from each of the three branches SPY1, SPY2, and SPY to
each of the three SDE1, SDE2, and SDE, and did so for the reverse
direction.  Taking the ratio of the first sum to the second resulted in a
distribution of the relative strength of gene flow between the two clades.
None of the ratio values was less than the corresponding ratio from the
prior (95\% quantile range was $1.79 <$ posterior ratio $< 2.24$, and prior
ratio was 1.41; P-value of $< 0.001$).  This provided genome-wide support
for Kalia et al.'s \citeyearpar{Kalia2001} unconfirmed observations of
directional asymmetry based on MLST data for seven loci.  We also observed
a slight enrichment for SDE-to-SDD edges.

% WORK IN DISCUSSION OF TABLE supplementary S\ref{tab:gene-counts-replacing}, SPECIFIC
% NUMBERS OF SPY$\rightarrow$SDE and SDE$\rightarrow$SDD TRANSFERS.  
\textbf{CHECK:} We counted genes with noticeable recombination probability 
($>0.6$) for types of recombinant edges
(supplementary table S\ref{tab:gene-counts-replacing}). There were more genes
that underwent SPY-to-SDE replacing gene transfer than those of SDE-to-SPY
transfer. This recapitulated the argument of asymetry of recplacing gene flow.
In the subset of virulence genes, there were three genes that underwent
SPY-to-SDE replacing transfer whereas there was one gene with the reverse
direction. 
Virulence genes and non-virulence genes did not seem to be different in the
relative frequency of replacing gene transfer
although the proportion of virulence genes with SPY-to-SDE replacing gene
transfer decreased compared with that of the total genes including the
virulence genes.

We generated new tracks for our recently released {\em Streptococcus}
Genome Browser (http://strep-genome.bscb.cornell.edu) that summarize the
results of the ClonalOrigin analysis alongside known genes, alignments, and
other annotations (fig.\ \ref{fig:ucsc}).  These tracks can be used to
inspect loci of interest and to compare the results of our model- and
parsimony-based analyses.  They can also be queried and intersected with
other tracks using the UCSC Table Browser.

% follow-up
% first, give bayesian posterior prob of asymmetry
% then something about effect of prior
% then try Ilan's time-interval test

\subsection*{Parsimony-based analysis}

% bring out comparison with replacing (will have to explore more)
% especially asymmetry, SPY-SDE

To gain further insight into HGT in {\em Streptococcus},
we estimated gene trees for 2314 gene families representing the six 
genomes (including an SEE outgroup), used the Mowgli
program \citep{Doyon2011} to find parsimonious 
reconciliations of these gene trees with the estimated clonal frame
(fig.\ \ref{fig:tree5}), and then, in a post-processing step, partitioned
the inferred transfers into replacing and additive HGTs (see \textbf{Methods}).
This analysis produced estimated numbers of five types of events on each
branch of the phylogeny (gene duplications, losses, replacing and additive
transfers, and ``appearances,'' i.e., transfers from phylogenetically
distant species), as well as an estimated number of genes at each ancestral
node of the tree (fig.\ \ref{fig:Gene-duplication-loss}).
% total numbers of events:
% appearance: 1957
% duplication: 257
% loss: 1293
% additive: 290
% replacement 928
We observed large numbers of appearances on all branches of the tree,
suggesting a steady influx of genes into the clade by HGT.  Indeed,
appearance events accounted for 1957 of the 3432 (57\%) gene additions (by
duplication or transfer).  Interestingly, the somewhat reduced numbers of
genes in SPY (particularly in SPY1) were primarily explained by reduced rates
of gene appearance, rather than by increased loss or other factors.  The
estimated numbers of ancestral genes tended to decrease toward the root of
the tree, but this likely reflects under-estimation due to parallel losses
of some genes (especially on the branches beneath the root) rather than a
true increase in gene number over evolutionary time.  An exception was the
branch leading to the ancestor of the two SPY individuals, which was
enriched for gene loss events, perhaps associated with niche adaptation.
High rates of loss also occurred on the branches to SDE1, SDE2, and SDD.
Duplications were relatively infrequent overall, but, as had been noted
previously \citep{Marri2006}, they were substantially enriched on external
branches of the phylogeny.

%When normalized for branch length (Figure \ref{fig:XX})

%The largest family consisted of 59 genes while nearly half of gene families
%(1066) had exactly one gene per species, or six genes total (Figure
%S\ref{fig:famsizes}).

We conservatively identified 290 of the 1218 (23.8\%) transfers (excluding
appearances) as additive HGTs. 
As with the replacing HGTs evaluated in the model-based analysis, these
predicted additive transfers were enriched within the SPY and SDE clades 
(fig.\ \ref{fig:Heatmap-of-transfers}B).  
They were also significantly enriched between
SDE to SDD (with a preference for the SDE-to-SDD direction), and
significantly depleted for SDD-to-SPY events.
However, unlike the replacing HGTs above, these additive transfers were only
slightly enriched between SPY and SDE, and did not exhibit a pronounced
directional asymmetry.  This suggests that the apparent
gene flow between SPY and SDE, particularly in the SPY-to-SDE
direction, has been driven by homologous recombination rather than additive
transfer.  However, our analyses of replacing and
additive transfers were not directly comparable for various reasons---for
example, the first considered the core genome only, while the second
considered the pan genome; and the two types of analysis might have quite
different power for different types of events.
To examine this issue further, we.... (REPORT BRIEF FOLLOW-UP ANALYSIS)
% follow-up analysis
We also compared the replacing HGTs identified in the parsimony-based
analysis with those from the model-based analysis, and found reasonable
concordance, despite substantial
differences between the data sets and methods 
(supplementary fig.\ S\ref{fig:cmpcomowgli}).
%Ilan: need to refer to appropriate supp figure

\subsection*{Functional Categories of Transferred Genes}

To gain insight into the functional impact of HGT, we assigned genes to
functional classes and tested each class to see whether it was enriched for
genes or gene families predicted to have experienced gene transfer events
(see \textbf{Methods}).  First, we identified a subset of genes whose homologs
in other
genera of bacteria were known to exhibit virulence phenotypes (based on
VFDB; see \citealp{Suzuki2011}) and considered all other genes to be
non-virulence genes.  These virulence genes were indeed significantly
enriched among gene families for which we inferred additive HGTs 
(P-value of $1.33 \times 10^{-11}$), 
possibly reflecting an increased rate of fixation of
these transfers due to positive selection for virulence-related functions.
% too much?
Interestingly, SPY-to-SDE events were strongly over-represented (by
$>$2x) among virulence genes that were associated with additive HGTs 
(supplementary table S\ref{tab:gene-counts-additive}).  
Transfers in the reverse direction
(SDE-to-SPY) are also over-represented, but slightly less so
(1.7x).  By contrast, virulence genes as a group showed no significant
enrichment for replacing gene transfers, and virulence genes that exhibit
strong evidence of replacing transfers were not enriched for SDE/SPY % SDE/SDD
transfers (rather, they were enriched for SDE-to-SDD transfers;
supplementary table S\ref{tab:gene-counts-replacing}).
We also found a weak association between virulence genes and gene
duplications (P-value of $7.13 \times 10^{-3}$).

% (POSSIBLY HIGHLIGHT AN EXAMPLE OR TWO WITH A BROWSER SHOT)
% Genome browser parts is moved to page 13.

% reduce this to a couple of sentences
We also 
searched for gene ontology (GO) terms significantly associated with the
various types of 
transfer. Among the categories found to be associated with additive gene
transfers (table \ref{tab:go-events}), we found several transposase-related 
transposase.  
%This can be explained by the role transposition plays in
%integrating transfered genes into a new host.
%Ilan: this explanation requires some work
Replacing transfers appear to show quite different associations from the
additive ones (table \ref{tab:functional}). (NEED TO EXPAND THIS A BIT)
%We
%compared associations found with replacing gene transfers inferred in our
%model-based analysis with associations found with the number of inferred
%replacing transfer events in the parsimony analysis, and overall found
%consistent associations (Table S\ref{tab:go-events-recombining}).

% can we still do something with examples?

%Ilan: this part still under construction

%THE FOLLOWING PARAGRAPH NEEDS SOME WORK

%We used the genome browser tracks for posterior probability of
%recombination to locate genes which showed an elevated signal of replacing
%gene transfer. We constructed a list of genes that contain sites showing a
%posterior probability greater than 0.8 for a given type of transfer
%(indicated by donor and recipient branches). Each such gene is associated
%with a transfer type (or several types). We find overall ?? genes fitting
%these criteria, with ?? of them associated with replacing transfers from
%SPY to SDE, and ?? associated with replacing transfers from SPY to
%SDE... Focusing on virulence genes we see a similar distribution of transfer
%types...

%Ilan: do we want a table with the number of genes associated with each transfer type?

% IS THERE SOMETHING WORTH KEEPING HERE?
%\texttt{INFORMATIONAL / OPERATIONAL ANALYSIS:}
%We grouped informational genes based on gene ontology to test if there was
%differences in replacing gene transfer between informational genes and
%operational genes. We found differences in replacing gene transfer from three
%SDE branches to three SPY branches (p-value=0.04409 with Wilcoxon rank sum
%test).
%\\
%\\

\subsection*{Positive Selection and Gene Transfer}
% selection analysis 
We also tested for evidence of positive selection within the gene
families and their association with various events.  We performed a
likelihood ratio test using PAML's ``sites models'' (M1a and M2b) on
each of the 2991 gene-fragment families with three or more genes.  Of
those families tested, 31 showed evidence of positive selection
(FDR$<$5\%).  For gene trees with high bootstrap ($>$90\% on every
branch), we find that duplication events are slightly enriched with
positive selection (P-value of $<0.046$; Mann-Whitney U) as well as
families with additive transfers (P-value of $<0.002$; Mann-Whitney
U).  Of the 31 families with significant positive selection, 13
families were part of the core genome analyzed by ClonalOrigin.  We
found that replacing transfers from SDE to SPY and from SPY to SDE to
be slightly enriched within the positively selected families (P-value
of $<0.005$ and $<0.01$; Mann-Whitney U).  For 11 of the 13
(85\%) families the SPY-to-SDE direction had greater intensity than
SDE-to-SPY, consistent with overall directional bias (83.3\%).  
The same directional bias was not seen for additive
transfers between the two clades.

% we listed several virulence genes with high ranks (Table S\ref{tab:virhgt}).
% The whole gene of  putative dipeptidase (SpyM3\_0465) was transferred, which was
% a exemplary case for the finding \citep{Chan2009} that virulence genes were transferred as a whole
% unit.  Another example of whole gene transfer was superoxide
% dismutase (SpyM3\_1071).  We investigated the unit of recombining gene transfer
% by finding 58 genes for which a recombinant edge of certain types were sampled
% with probability at least 0.9 by focusing on specific types of recombinant edges
% (Table S\ref{tab:genes-transfer}).  In some cases, parts of a gene were inferred
% to have been transferred, for instance, transfer from SPY1 to SDE1 was inferred
% for 29\% of the aspartyl-tRNA synthetase gene. Recombinant tracts often were
% shown to span two genes.  In other cases, the inferred transfer units spaned a
% set of as much as eight adjacent genes: loci from SpyM3\_1518 to SpyM3\_1525.

\section*{Discussion}

% accurate to say "recent"?  has been going on for some time
There has been a great deal of recent interest in the process of horizontal
gene transfer (HGT) and the manner in which it has influenced phylogenetic
relationships, particularly among bacteria \citep{}.  Most previous studies,
however, have either simply examined patterns of phylogenetic discordance,
without differentiating between additive and replacing HGTs \citep{}, or
have focused specifically on the process of homologous recombination,
corresponding to replacing HGTs
\citep{Didelot2007,Didelot2010}.
% too strong?  neither are we
% cite Feil?   Marri.  LEfebre.  recent review (Doolittle?)
% another homologous recombination review -- see Awadalla
To our knowledge, this is the first study to examine both
processes on a genome-wide scale, using both model- and parsimony-based
methods.

We have focused our analysis on {\em Streptococcus pyogenes}
(SPY) and two of its close relatives, {\em S.\ dysgalactiae} ssp.\ {\em
  equisimilis} (SDE) and {\em S.\ dysgalactiae} ssp.\ {\em dysgalactiae}
(SDD), a human commensal-like organism and a strict veterinary pathogen,
respectively.  Our combined analysis reveals strong evidence of gene flow
both within and between the SPY and SDE groups.  Interestingly, the SPY-SDE
transfers are far more prominent among replacing than among additive
transfers, suggesting that they are driven by homologous recombination.
Moreover, we find strong support for an asymetry in replacing gene
transfers between SPY and SDE, with a preference for the SPY-to-SDE
direction.  The situation with SDD is more complex, with somewhat elevated
rates of SDE-to-SDD transfer, more pronounced for additive than for
replacing HGTs, mixed evidence for SDD-to-SDE transfer, and reduced rates
of transfer between SDD and SPY, perhaps owing to a combination of genetic
divergence and different ecological niches.
% maybe revise the below after tuning up virulence section
We also find that virulence genes are significantly enriched among
additive, but not replacing, gene transfers, as are genes under positive
selection.  Genes that have undergone replacing transfers between SPY and
SDE are also enriched for positive selection.
A central component of our study is the careful use of simulations to
demonstrate that our methods have good power and accuracy in the
detection of both types of transfer events.

%We also confirm a species phylogeny (clonal frame) consistent with previous
%results 
%\citep{Facklam2002,Suzuki2011}

%paragraph on Kalia et al. and the evidence for SDE/SPY gene flow
%reasons for asymmetry -- barriers, etc.

The finding of abundant, asymmetric gene flow between
SPY and SDE raises a number of questions.
% this is perhaps not so pressing; discuss briefly
First, can Kalia et al.'s \citeyearpar{Kalia2001} observations be
confirmed using our data, despite the authors' inability to reproduce them?
\texttt{CHECK:} 
Interestingly, we did find 
that the six of their seven loci represented in our five-way alignments
(\textit{gki}, \textit{gtr}, 
\textit{murI}, \textit{mutS}, \textit{recP}, and \textit{xpt}) showed
consistent results with those of \citet{Kalia2001}; high recombination in 
\textit{recP}, and \textit{xpt} but not the others. 
Recombination in \textit{gki} and \textit{mutS}, albeit weak, were also
observed (despite that Kalia et al.\ found no evidence for {\em mutS}).
Our visual inspection of the recombination probability on 
\textit{recP} and \textit{xpt} indicated that 
the replacing gene transfer appeared to be biased towards
the directions from SPY to SDE.
% In S. pyogenes MGAS315 genome track, we can check
% gki -> not in the region but its neighboring
% chr1:1207587-1208084
% /locus_tag="SpyM3_1180"
% 
% gtr -> no recombination 
% chr1: 1186522-1186073
% /locus_tag="SpyM3_1160"
%
% murI -> no recombination
% chr1:303215-303652
% /locus_tag="SpyM3_0262"
%
% mutS -> a little bit
% chr1:1836841-1837245
% /locus_tag="SpyM3_1806"
%
% recP -> somewhat strong (consistent with Kalia2001's conclusion)
% chr1:1454877-1455335
% /locus_tag="SpyM3_1462"
%
% xpt -> somewhat strong
% chr1:846784-847233
% /locus_tag="SpyM3_0794"
%
% yqil -> no data available
% chr1:131266-131699
% /locus_tag="SpyM3_0108"
%
%\texttt{SC: I am not sure, yet, because I wrote,
%Except for \textit{mutS} where
%they did not find recombination, our results were
%consistent with those of \citet{Kalia2001}. Let me make sure this.}
Of course, our study differs from theirs in numerous respects, but this
suggests the two loci are prone to recombination.
%Except for \textit{mutS} where
%\cite{Kalia2001} did not find recombination, our results were
%consistent with those of \cite{Kalia2001}.  
%\cite{Pinho2010} reported recombination in the gene \textit{parC}, in which we
%also observed moderate recombination.
Second, are the observed patterns of gene flow truly driven by replacing
transfers, or could additive transfers also be contributing?  Here, our
evidence is weaker, because our parsimony-based method has limited ability
to resolve the direction of transfer.  More genomes will be needed to
evaluate the directionality of additive transfers between these groups.
Third, what could explain the observed asymmetry in gene flow?  Possible
mechanisms include differences in the sizes or demographic structure of
co-mingling populations of cells, or differences in barriers to gene
transfer \citep{Thomas2005}.  \cite{Kalia2001} speculated that 
SDE might be inherently less capable than SPY to acquire DNA by homologous
recombination, for example, due to differences in the 
restriction-modification system (RMS), hetero-duplex formation, and/or mismatch
repair.  
%Restriction-modification
%system is a biological self-defense mechanism of bacteria, which is diverse
%in prokaryote world.  
Interestingly,
while most of \textit{S.\ pyogenes} genome contain both type I and type II
RMSs, the two SDE genomes studied here contain
only type II RMSs.  (CHECK: THIS NEEDS SOME WORK)

% can we integrate some of this?  reread Kalia
%More recently, 
%As biological roles of RMS might be to maintain and
%control species identity \citep{Jeltsch2003}, \cite{Budroni2011a} recently
%studied that \textit{Neisseria meningitidis} 
%phylogenetic clades were associated with RMS, claiming that RMS modulated
%homologous recombination in the bacteria of \textit{N.\ meningitidis}.

%Most of \textit{S.\ pyogenes} genomes contained restriction-modification
%type I system as well as type II system. The two genomes of SDE contained
%only type II system. A more robust self-defense system in SPY could have
%made the species less capable of acquiring foreign genetic material from
%SDE.
% can we follow-up on this in some way?

%More SDE genomes might be beneficial to future studies.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% be careful not to undercut our findings here.

% maybe more than a paragraph of limitations, make it one about insight --
% idea of the two events being identified with different parts of the genome

In comparing replacing and additive HGTs, it is worth bearing in mind that
these events are likely to be strongly correlated with the core and
``dispensable'' portions of the genome (we take the dispensible genome to
be the difference between the pan and core genomes).  Methods for detecting
recombination (replacing transfer) tend to rely on large-scale alignments
spanning multiple syntenic loci in order to gain power.  As a result, these
methods are effectively limited to application in the core genome, even
though homologous recombination probably also occurs (perhaps at a reduced
rate) between genes that are less widely shared across species.  By
contrast, phylogenetic reconciliation methods can be applied to the entire
pan genome, as in this study.  Moreover, the additive transfer events that
they detect will necessarily be enriched in the dispensible genome.  Thus,
the replacing HGTs identified by recombination-based methods will tend to
reflect the core genome and the additive HGTs identified by
reconcilation-based methods will tend to reflect the dispensible genome.
In our view, this is not necessarily a limitation of the comparison,
because (with the exception of some replacing transfers in the dispensible
genome) we expect the inferred transfer events to be fairly representative
of true HGTs, but it is useful to keep in mind that some apparent
differences between the two 
types of transfer events may simply reflect general differences between the
core and dispensible genome.

%In addition, these methods are likely to have quite different power for various
%kinds of events, because of factors such as
%branch-length dependency
%use of multiple loci or single loci
%larger trees

%paragraph on future methods
As sequence data becomes available for many more species, including both
closely and distantly related organisms, it will become increasingly
important to devise improved methods that consider both additive and
replacing HGT.  As noted above, the current statistical-models of
recombination, such as ClonalOrigin, are effectively limited to considering
orthologs, with one copy per species, in the core genome.  The phylogenetic
reconciliation methods, on the other hand, fail to allow for multi-locus
gene transfer events (which are known to occur), and also ignore
information from branch lengths.  We have shown here that it is possible,
to a degree, to extend them to distinguish between additive and replacement
transfers but our methods are limited to a degree by their dependency on
heuristic rules and parsimony assumptions (reference supplementary section
on simulation experiments).  It may be possible to develop integrated
statistical models that consider both types of gene transfers, and to use
them to more directly compare the rates with which they occur.  Another
extension worth considering is direct modeling of incomplete lineage
sorting (ILS), which is likely to be increasingly important as larger
phylogenies are considered, especially given the large effective population
sizes of many bacteria.  ILS has recently been integrated into models for
duplication and loss (Rasmussen and Kellis) but has yet to be considered in
reconciliation-based methods that also allow for gene transfer.  It is
permitted by ClonalOrigin, but only in an indirect and approximate manner,
through the introduction of recombination edges to ancestral branches
(which can be misinterpreted as evidence of recombination).  The
combination of improved models and richer data sets could allow for a much
more detailed understanding of horizontal gene transfer.

%Molecular evolution of bacterial genomes is multifaceted. Researches of
%bacterial evolution often focused on one side of many evolutionary processes.
%While understanding one aspect of bacterial evolutionary processes we could be
%ignorant of the other features.  Horizontal gene transfer process was often
%considered in macro-evolutionary scale using a great deal of bacterial species.
%As closely related species or groups of individual bacteria are becoming a focus
%of evolutionary study, recombining gene transfer process would surface as
%another facet of evolutionary forces.  Because the two gene transfer processes
%would act on closely related bacterial genomes simultaneously, ignoring one of
%the two processes in studying evolution of bacteria would lead to a narrowed
%point of view to the underlying evolutionary forces. Our study of SPY and SDE
%genomes undertook the work of discriminating recombining and horizontal gene
%transfer processes because the two species were very closely related.  In this
%study, we showed that the two gene transfer processes acted on different
%biological functions.  Recombining gene transfer process in the analysis of SPY
%and SDE genomes was related with genes encoding membrane-bound molecules and
%aminoacyl tRNA synthetases. On the contrary, horizontal gene transfer process
%was found to be associated with transposases, which would allow horizontal gene
%transfer of their neighboring genes.  Because SPY was a human-specific pathogen,
%and SDE was an opportunistic human pathogen, we were interested in association
%of pathogenicity with the two gene transfer processes. We also showed that
%virulence genes were more associated with horizontal gene transfer process than
%with recombining gene transfer process.  With the current data of only SPY and
%SDE genomes we could not conclude whether the association of virulence genes
%with horizontal gene transfer not recombining gene transfer was applicable
%generally to cases of other species.  However, without the discrimination of the
%two processes it would have been difficult to find the disparate association of
%biological functions or virulence genes with the different gene transfer
%processes. 
%
%We showed differences in the profile of functional category association of the
%two gene transfer processes thanks to the approach of dicriminating the two gene
%transfer processes.  Sugar phosphotransferases could be important in signal
%transduction to recognize environmental changes.  Transporters, and fatty acid
%and lipid biosyntheses would be involved in membrane-bound molecules.  Because
%bacteria as a unicellular organism should be able to promptly respond to changes
%to environment, some evolutionary pressures must have been acting on genes
%encoding membrane-bound molecules.  We also found that the recombination
%intensity and requency of recombining gene transfer events would be useful in
%associating several aminoacyl tRNA synthetases genes belonging to gene ontology
%term of translation (GO:0006412) with recombining gene transfer. It had been a
%focus of gene transfer in bacterial species \citep{Woese2000}.  The biological
%role of transposases was concordant with horizontal gene transfer process
%because the process involved in transposases would allow bacteria to acquire
%foreign genes without the requirement of homologous regions in the bacteria.  
%
%Another contribution of our study to bacterial gene transfer research is to
%elucidate the net directionality of gene flow from SPY to SDE in the core
%genomes. Although the issue of net directionality has been tackled, our study
%was the first attempt of resolving the issue in genome-wide scale. Yet, it still
%remains to be seen whether this net directionality of gene flow from SPY to SDE
%happened in the pan genome. A model-based method with more refined models would
%be necessary because the parsimony-based method we employed simply could not
%show any net directionality between SPY and SDE in the pan genome. Although we
%are unsure whether much larger number of genomes could be useful in finding net
%directionality, more SDE genomes might be beneficial to future studies. 
%
%Were there unbalanced recombining gene transfer between the two species, what
%were the barriers \citep{Thomas2005} to cause the unbalanced gene transfer in
%the species?  Kalia et al.\ \cite{Kalia2001} discussed restriction-modification
%system (RMS) as a possible mechanism of the unbalanced gene flow.
%Restriction-modification system is a biological self-defense mechanism of
%bacteria, which is diverse in prokaryote world.  As biological roles of RMS
%might be to maintain and control species identity \citep{Jeltsch2003}, Budroni
%et al.\ \cite{Budroni2011a} recently studied that \textit{Neisseria
%meningitidis} phylogenetic clades were associated with RMS, claiming that RMS
%modulated homologous recombination in the bacteria of \textit{N.\ meningitidis}.
%Most of \textit{S.\ pyogenes} genomes contained restriction-modification type I
%system as well as type II system. The two genomes of SDE contained only type II
%system. More tight self-defense system of SPY may have made the species less
%capable of aquiring foreign genetic material from SDE \citep{Kalia2001}.
%
%We reexamined the seven genes that Kalia et al.\ \cite{Kalia2001} had studied
%for net directionality of gene transfer between SPY and SDE.  The internal
%fragments of the six housekeeping genes (\textit{gki}, \textit{gtr},
%\textit{murI}, \textit{mutS}, \textit{recP}, \textit{xpt}) were located in the
%core genome that we investigated in this study.  Except for \textit{mutS} where
%\cite{Kalia2001} did not find recombination, our results were
%consistent with those of \cite{Kalia2001}.  
%\cite{Pinho2010} reported recombination in the gene \textit{parC}, in which we
%also observed moderate recombination.
%
%It remains to be seen whether the inability of detecting net directionality
%between SPY and SDE in the pan genomes was because of lack of power of the
%parsimony-based approach that we employed. Considering gene duplication, loss,
%and horizontal gene transfer in a model-based approach would be desirable in the
%coming ages of abundant bacterial genomes. The model might consider genome
%alignments where each taxon does not necessarily have a unique gene; some taxa
%might have no gene due to gene loss, and other taxa multiple genes due to gene
%gain, duplication, or horizontal gene transfer. With these types of models will
%we be able to address more interesting evolutionary questions in bacteria. This
%need of more refined models would be also true even for the study of net
%directionality between species in the core genome.  In other words, one could
%envision a model where a population two-taxa (i.e., SPY and SD groups) tree is
%superimposed on the species tree shown in Figure S\ref{fig:clonalorigin}A. A
%splitting event of the population tree could happen somewhere between the root
%of the species tree and the internal node of SPY. Recombinant edges that connect
%branches within SPY or SD clades would be more frequent than those that connect
%branches between SPY and SDE clades.  With the reference to the recombination
%within a clade one parameter can dictate the degree of gene from from SPY to
%SDE, and another for the reverse direction.  If the two parameters could be
%incorporated in the model, then they might be used to compare the degrees of
%gene flow between the two species in a more statistically sound way.  We hope
%that we will see this refinement of the model to better illustrate bacterial
%evolution in forthcoming bacterial genomic era.

% ACS: Let's let the journal do this 
%\section*{Supplementary Material}
%Supplementary figures S1-S10 and tables S1-S9 are available at Molecular Biology
%and Evolution online (http://mbe.oxfordjournals.org/).

% Do NOT remove this, even if you are not including acknowledgments
\section*{Acknowledgments}

We thank Xavier Didelot for assistance with the ClonalOrigin analysis, and
Haruo Suzuki for assistance with the data for SDD. This work was primarily
supported by the National Institute of Allergy and Infectious Disease, US
National Institutes of Health, under grant number AI073368-01A2 (to
M.J.S. and A.S).  Additional support was provided by National Science
Foundation CAREER Award DBI-0644111 and a David and Lucile Packard
Fellowship for Science and Engineering (to A.S.).

%\section*{Author Contributions}

%Conceived and designed the experiments: SCC MDR IG MJS AS.
%Performed the experiments: SCC MDR MJH.
%Analyzed the data: SCC MDR MJH IG.
%Contributed reagents/materials/analysis tools: MJS.
%Wrote the paper: SCC IG AS (with review and contributions from all authors).
%\clearpage

% The bibtex filename
\renewcommand*{\refname}{Literature Cited}
\bibliographystyle{../../latex/bst/mbe}
\bibliography{siepel-strep}
\clearpage{}

\section*{Tables}
%\begin{table}[!ht]
%\caption{
%\bf{Table title}}
%\begin{tabular}{|c|c|c|}
%table information
%\end{tabular}
%\begin{flushleft}Table caption
%\end{flushleft}
%\label{tab:label}
% \end{table}

%=============================================================================
% Table 2. 
\begin{table}[!ht]

\caption{{\bf Functional category enrichments for additive gene
  transfers}}
\vspace{1ex}

\noindent \begin{centering}
\begin{tabular}{ccccl}
\hline 
$p^a$ & $q^b$ & Count$^c$ & GO term & Description \\
\hline 
7e-61 & 3e-58 & 67 & GO:0006313 & transposition, DNA-mediated\\
3e-58 & 7e-56 & 59 & GO:0004803 & transposase activity\\
1e-30 & 2e-28 & 52 & GO:0015074 & DNA integration\\
1e-28 & 1e-26 & 23 & GO:0032196 & transposition\\
1e-13 & 1e-11 & 527 & GO:0003677 & DNA binding\\
2e-12 & 2e-10 & 86 & GO:0006310 & DNA recombination\\
1e-11 & 7e-10 & 197 & GO:0003676 & nucleic acid binding\\
3e-09 & 1e-07 & 36 & GO:0016987 & sigma factor activity\\
3e-09 & 1e-07 & 36 & GO:0006352 & transcription initiation\\
3e-08 & 1e-06 & 142 & GO:0043565 & sequence-specific DNA binding\\
4e-07 & 1e-05 & 12 & GO:0008170 & N-methyltransferase activity\\
1e-06 & 4e-05 & 12 & GO:0019867 & outer membrane\\
2e-06 & 5e-05 & 25 & GO:0007059 & chromosome segregation\\
1e-05 & 4e-04 & 15 & GO:0006306 & DNA methylation\\
2e-05 & 6e-04 & 15 & GO:0006470 & protein amino acid dephosphorylation\\
2e-04 & 4e-03 & 76 & GO:0005618 & cell wall\\
3e-04 & 7e-03 & 25 & GO:0008236 & serine-type peptidase activity\\
3e-04 & 7e-03 & 65 & GO:0009986 & cell surface\\
5e-04 & 1e-02 & 10 & GO:0003872 & 6-phosphofructokinase activity\\
1e-03 & 2e-02 & 11 & GO:0009307 & DNA restriction-modification system\\
2e-03 & 5e-02 & 84 & GO:0006974 & response to DNA damage stimulus\\
\hline 
\end{tabular}
\par\end{centering}
\begin{flushleft}
$^a$$P$-values based a Mann-Whitney $U$ test.  Shown are values that
correspond to FDR$^b$  $<0.05$ based on the 
Benjamini-Hochberg method (Methods).\\
$^c$Number of genes in category.\\
\end{flushleft}
\label{tab:go-events}
\end{table}
%\clearpage{}
%=============================================================================

%=============================================================================
% /Users/goshng/Documents/Projects/Mauve/output/cornellf/3/run-analysis/significant-1.txt
% Table 1. 
\begin{table}[!ht]
\caption{
{\bf Functional category enrichments for replacing gene transfers$^a$}} 
\vspace{1ex}
\noindent \begin{centering}
\begin{tabular}{ccccl}
\hline 
$p^b$ & $q^c$ & Count$^d$ & GO term & Description\\
\hline 
2e-06 & 2e-04 &  38 & GO:0009401 & phosphoenolpyruvate-dependent sugar phosphotransferase system\\
3e-06 & 2e-04 &  33 & GO:0008982 & protein-N(PI)-phosphohistidine-sugar phosphotransferase activity\\
5e-05 & 3e-03 &  22 & GO:0006633 & fatty acid biosynthetic process\\
9e-05 & 4e-03 &  21 & GO:0008610 & lipid biosynthetic process\\
3e-04 & 9e-03 &  45 & GO:0008643 & carbohydrate transport\\
5e-04 & 1e-02 &  82 & GO:0006412 & translation\\
6e-04 & 1e-02 &  13 & GO:0006814 & sodium ion transport\\
1e-03 & 2e-02 &  40 & GO:0030529 & ribonucleoprotein complex\\
1e-03 & 2e-02 &  12 & GO:0031402 & sodium ion binding\\
1e-03 & 2e-02 & 244 & GO:0008152 & metabolic process\\
2e-03 & 3e-02 & 378 & GO:0016740 & transferase activity\\
2e-03 & 3e-02 &  39 & GO:0003735 & structural constituent of ribosome\\
2e-03 & 3e-02 &  26 & GO:0019843 & rRNA binding\\
4e-03 & 4e-02 &  66 & GO:0005840 & ribosome\\
\hline 
\end{tabular}
\par\end{centering}
\begin{flushleft}
$^a$Based on recombination intensities restricted to topology-altering
recombinant edges (Methods).\\
$^b$$P$-values based a Mann-Whitney $U$ test.  Shown are values that
correspond to FDR$^c$  $<0.05$ based on the 
Benjamini-Hochberg method (Methods).\\
$^d$Number of genes in category.\\
\end{flushleft}
\label{tab:functional}
\end{table}
\clearpage{}
%=============================================================================

\section*{Figures}

%=============================================================================
\begin{figure}[!ht]
\begin{center}
\includegraphics[width=3in]{figures/cells.pdf}
\end{center}
\caption{ {\bf Replacing and additive horizontal gene transfer.}  (A)
  Foreign DNA can be integrated into a recipient genome by homologous
  recombination (a {\em replacing transfer}, left) or additive integration
  (an {\em additive transfer}, right). (B) These types of transfers produce
  distinct phylogenetic signatures.  In homologous recombination a segment
  of DNA is effectively overwritten by a homologous segment from another
  species, which causes a lineage in the phylogeny to be replaced by a
  transferred lineage.  This type of transfer can be identified in
  gene-tree/species-tree reconciliation by the appearance of coinciding
  transfer and loss events (left tree).  Additive integration, on the other
  hand, leads to a transfer event that is not paired with a loss event
  (right tree).  Of course, parallel losses can cause an additive transfer
  to appear similar to a replacing transfer.  In our analysis, we
  conservatively require that at least one descendant species (here, {\bf
    B}) contains genes that both descend from ($b_1$) and do not descend
  from ($b_2$) a lineage predicted to have been transferred when
  inferring an additive transfer event.}
\label{fig:hgt}
\end{figure}
\clearpage{}%
%=============================================================================

%=============================================================================
\begin{figure}[!ht]
\noindent \begin{centering}
\includegraphics[scale=0.5]{figures/cornellf-3-tree}
\par\end{centering}
\caption{
{\bf Clonal frame inferred for the five genomes.}  The branch lengths are
drawn to scale in the horizontal dimension.  Units are 
expected substitutions per site given Watterson's estimate of the
mutation rate. The branch length of the outgroup SEE is not scaled. 
The labels for the ancestral nodes of the tree and the
internal branches (SPY, SDE, and SD) are used throughout the paper.}
\label{fig:tree5}
\end{figure}
\clearpage{}%
%=============================================================================

% A transfer is a additive if there exists an extant descendant species
% (e.g. C and D) that contains both descendants (bold) and
% non-descendants (normal weight) of the transferred gene (black
% circle).

%=============================================================================
\begin{figure}
% \includegraphics[scale=0.45]{figures/heatmap-recedge}
\includegraphics[scale=0.5]{figures/heatmap}
\caption{\label{fig:Heatmap-of-transfers}
{\bf Heatmaps showing rates of replacing and additive transfers.}
Each cell of the heat map represents the base-2 logarithm 
of the ratio of the estimated number recombination events to its expectation
under the prior, for the corresponding donor ($y$ axis) and recipient ($x$
axis) branches.  
The black colored cells indicate prohibited transfer events.
%
(A) Replacing transfers, as inferred by the model-based approach. The
plotted values represent 
average values across sampled recombinant
trees. The prior considers the clonal frame with branch lengths and the
global estimates of the three population parameters. Prohibited transfer events
are ones for which the recipient branch is strictly older than the donor
branch. Significance levels are denoted by asterisks: one for $<0.01$.
%
(B) Additive transfers, as inferred by the parsimony-based
approach. The plotted values represent
numbers of inferred additive transfers across all gene families.
The prior considers the clonal  frame (with branch lengths) and the total number
of events of each type inferred by the analysis. Prohibited transfer events are
ones for which the recipient and donor branch do not share a common time
interval.
branch. Significance levels are denoted by asterisks: one for $<0.01$, two for
$<0.005$, and three for $<0.001$.}
\end{figure}
\clearpage{}
%=============================================================================

%=============================================================================
\begin{figure}[!ht]
\includegraphics[scale=0.5]{figures/ucsc}
% used genomic interval chr1:511,000-513,500
\caption{ {\bf Genome Browser tracks.}  Our inferences of replacing and
  additive horizontal gene transfers are summarized in new tracks in the
  {\em Streptococcus} Genome Browser \citep{Suzuki2011}.  (A) The main
  browser display, with gene annotations for the selected reference genome
  ({\em S.\ pyogenes} strain MGAS315) shown in red and putative virulence
  genes highlighted in blue. The third track from top (in black) shows the
  putatively non-recombining gene fragments analyzed by Mowgli.  The next series
  of tracks shows the results of the ClonalOrigin analysis of
  replacing gene transfers.  Each track in this series corresponds to a
  recipient lineage in the phylogeny (fig.\ \ref{fig:tree5}) and describes
  the posterior probabilities along the genome of recombination edges from
  all possible donor lineages (shown in different colors; see key at
  right).  Here the putative virulence gene SpyM3\_0465, a dipeptidase,
  shows strong evidence of a recombination edge from the SPY to the 
  SDE lineage, as well as some evidence of a SPY$\rightarrow$SPY2 edge.
  The genome-wide multiple alignment obtained with Mauve is shown at
  bottom.  (B) The gene tree that is displayed after clicking on the
  highlighted gene fragment.  The topology inferred by RAxML is consistent
  with the combined influence of SPY$\rightarrow$SDE and
  SPY$\rightarrow$SPY2 replacing transfers, as inferred by ClonalOrigin.}
\label{fig:ucsc}
\end{figure}
\clearpage{}%
%=============================================================================

%=============================================================================
\begin{figure}
\includegraphics[width=7in]{figures/strep-events}
\caption{\label{fig:Gene-duplication-loss} 
{\bf Distribution of inferred gene duplication, loss, and transfer events
  for the six-species phylogeny.}  Gene counts for
each extant and ancestral species are given at each node of the tree.  For
each branch numbers of gene appearances (A*), duplications (D*), 
losses (L*), additive transfers (T*), and replacing transfers (R*) are listed.
Transfer events are recorded on the recipient branch (donors
are not indicated).}
\end{figure}
\clearpage{}%
%=============================================================================

\clearpage{}\setcounter{figure}{0}
\setcounter{table}{0}
\renewcommand{\figurename}{Supplementary Figure}
\renewcommand{\tablename}{Supplementary Table}

\section*{Supplementary Tables}

%=============================================================================
\begin{table}[!ht]
\caption{
{\bf The six bacterial genomes.}
Each bacterial genome is referred to as the name at the first column.}
\noindent \begin{centering}
\begin{tabular}{cccc}
\hline 
Name & NCBI accession & Size (base pairs) & Reference\tabularnewline
\hline
SDE1 & CP002215 & 2159491 & Suzuki et al. (2011)\tabularnewline
SDE2 & NC\_012891 & 2106340 & Shimomura et al. (2011)\tabularnewline
SDD & N/A$^a$ & 2141837 & Suzuki et al. (2011)\tabularnewline
SPY1 & NC\_004070 & 1900521 & Beres et al. (2002)\tabularnewline
SPY2 & NC\_008024 & 1937111 & Beres et al. (2006)\tabularnewline
SEE & NC\_012471 & 2253793 & Holden et al. (2009)\tabularnewline
\hline
\end{tabular}
\par\end{centering}
\begin{flushleft}
$^a$ SDD genome was not yet available from NCBI.
\end{flushleft}
\label{tab:genome}
\end{table}
%\clearpage{}
%=============================================================================

%=============================================================================
\begin{table}[!ht]
\caption{
{\bf Posterior probability of gene tree topologies.}
See Table S\ref{tab:genome} for the reference of the bacteria
individual names.}
\noindent \begin{centering}
\begin{tabular}{cc}
\hline 
Gene tree topologies & Posterior probability\tabularnewline
\hline
(((SDE1,SDE2),SDD),(SPY1,SPY2)) & 66.7\%\tabularnewline
(((SDE1,SDE2),(SPY1,SPY2)),SDD) & 9.3\%\tabularnewline
((SDE1,SDE2),(SDD,(SPY1,SPY2))) & 3.5\%\tabularnewline
(((SDE1,(SDE2,SDD)),(SPY1,SPY2)) & 1.5\%\tabularnewline
(((SDE1,SDD),SDE2),(SPY1,SPY2)) & 1.3\%\tabularnewline
((SDE1,SDD),(SDE2,(SPY1,SPY2))) & 1.2\%\tabularnewline
((((SDE1,SDE2),SPY1),SPY2),SDD) & 1.0\%\tabularnewline
((SDE2,SDD),(SDE1,(SPY1,SPY2))) & 1.0\%\tabularnewline
\hline
\end{tabular}
\par\end{centering}
\label{tab:Gene-tree-topologies}
\end{table}
%\clearpage{}
%=============================================================================

%=============================================================================
\begin{table}[!ht]
\caption{
{\bf Ratios of logarithm base 2 of the observed
recombinant edges to the expected number of recombinant edges.}  Figure
\ref{fig:Heatmap-of-transfers}A shows the values graphically using heat map.
NA's are for the branch pairs that are impossible under the model-based
approached. Rows are for donor, and columns for recipient.}
\noindent \centering{}\begin{tabular}{cccccccccc}
\hline
& SDE1 & SDE & SDE2 & SD & SDD & ROOT & SPY1 & SPY & SPY2 \tabularnewline
\hline
SDE1& 0.176  &   NA  &1.170  &    NA & 0.7564  & NA & 0.67 &    NA & 0.663\tabularnewline
SDE & -0.118 & 0.081 &-0.202 &     NA&  0.3395 &  NA& -0.15&  0.045& -0.323\tabularnewline
SDE2& 1.201  &   NA  &0.092  &    NA & 0.6978  & NA & 0.60 &    NA & 0.387\tabularnewline
SD  & -0.629 & 0.022 &-0.661 &0.00670&  0.0078 &  NA& -0.94& -0.299& -1.030\tabularnewline
SDD & -0.317 & 0.371 &-0.197 &     NA&  0.0752 &  NA& -1.06& -0.525& -0.915\tabularnewline
ROOT& -1.745 &-0.698 &-1.786 &0.23039& -0.4854 &  NA& -1.37&  0.497& -1.306\tabularnewline
SPY1& 1.245  &1.883  &1.251  &    NA &-2.1997  & NA & 0.06 &    NA & 1.062\tabularnewline
SPY & 0.089  &0.200  &0.298 &0.00098 &-0.7903  & NA &-0.19 &-0.058 &-0.159\tabularnewline
SPY2& 1.174  &1.938  &1.287  &    NA &-1.9740  & NA & 1.07 &    NA & 0.063\tabularnewline
\hline
\end{tabular}
\label{tab:heatmap}
\end{table}
%\clearpage{}
%=============================================================================

%=============================================================================
\begin{table}[!ht]
\caption{
{\bf The observed number of recombinant edges.}
The zeros are for the branch pairs that are impossible under the model-based
approached. Rows are for donor, and columns for recipient.}
\noindent \centering{}\begin{tabular}{cccccccccc}
\hline
& SDE1 & SDE & SDE2 & SD & SDD & ROOT & SPY1 & SPY & SPY2 \tabularnewline
\hline
SDE1&8.896 &  0.00 &17.712  & 0.00 & 12.872   & 0 &12.59  &  0.00  &12.53\tabularnewline
SDE &44.626& 136.49& 42.088 &  0.00& 227.061  &  0& 73.39 &  99.95 & 65.25\tabularnewline
SDE2&18.099&   0.00&  8.393 &  0.00&  12.360  &  0& 11.97 &   0.00 & 10.36\tabularnewline
SD  &9.671 &101.74 & 9.464  &31.26 &115.796   & 0 &13.68  & 98.38  &12.86\tabularnewline
SDD &45.539& 168.58& 49.513 &  0.00& 194.706  &  0& 43.09 &  67.14 & 47.65\tabularnewline
ROOT&43.553& 602.67& 42.337 &765.53& 802.813  &  0& 98.86 &2146.01 &103.60\tabularnewline
SPY1&40.800&  11.46& 40.973 &  0.00&   4.371  &  0& 21.15 &   0.00 & 42.35\tabularnewline
SPY &57.618& 262.08& 66.594 & 30.92& 163.518  &  0& 83.56 & 207.00 & 85.10\tabularnewline
SPY2&38.837&  11.90& 42.000 &  0.00&   5.111  &  0& 42.45 &   0.00 & 21.19\tabularnewline
\hline
\end{tabular}
\label{tab:obsheatmap}
\end{table}
%\clearpage{}
%=============================================================================

%=============================================================================
\begin{table}[!ht]
\caption{
{\bf Number of genes showing significant signal of replacing transfer.}  For
each type of recombinant edge (replacing gene transfer event in
\texttt{ClonalOrigin} analysis), we record the number of genes with an average
posterior probability greater than 0.6 for a recombinant edge of that type
across their sites. Recombinant edges from the root branch are not considered
here, and SPYclade and SDEclade each signify three branches in the clonal frame
corresponding to the respective clades. Edge types not shown here have a zero
count. The same is done for the subset of virulence genes. Absolute numbers are
shown together with ratios relative to the total number of genes, as well as the
total number of significant replacing transfers found in this analysis.}
\noindent \centering{}\begin{tabular}{lll}
\hline
Recombinant edge type & Gene count & Virulence gene count \tabularnewline
\hline
SDEclade $\rightarrow$ SDD & 43 (33.1\%) & 5 (50\%) \tabularnewline
SDEclade $\rightarrow$ SPYclade & 11 (8.5\%) & 1 (10\%) \tabularnewline
SDD $\rightarrow$ SDEclade & 1 (0.8\%)  & 0 (0\%) \tabularnewline
SPYclade $\rightarrow$ SDEclade & 59 (45.4\%) & 3 (30\%) \tabularnewline
SPYclade $\rightarrow$ SPYclade & 16 (12.3\%) & 1 (10\%) \tabularnewline
All types & 130  (100\%) & 10   ( 100\%)\tabularnewline
% &  (11.8\% of total number of genes) & (13\% of virulence genes)\tabularnewline
\hline
\end{tabular}
\label{tab:gene-counts-replacing}
\end{table}
%\clearpage{}
%=============================================================================

%=============================================================================
\begin{table}[!ht]
\caption{
{\bf Number of inferred transfer events of different types.}  We record the number of inferred additive transfer events of each type. SPYclade and SDEclade each signify three branches in the clonal frame corresponding to the respective clades. Edge types not shown here have a zero count. The same is done for the subset of virulence genes. Absolute numbers are shown together with ratios relative to the total number of genes, as well as the total number of significant replacing transfers found in this analysis.}
\noindent \centering{}\begin{tabular}{lll}
\hline
Recombinant edge type & Gene count & Virulence gene count \tabularnewline
\hline
SDEclade $\rightarrow$ SDEclade & 28 (10.6\%) &  5 (12.5\%)\tabularnewline
SDEclade $\rightarrow$ SDD & 64 (24.2\%) & 2 (5.0\%)\tabularnewline
SDEclade $\rightarrow$ SPYclade & 28 (10.6\%) &  7 (17.5\%)\tabularnewline
SDD $\rightarrow$ SDEclade & 29 (10.9\%) &  5 (12.5\%)\tabularnewline
SDD $\rightarrow$ SPYclade & 17 (6.4\%) &  2 (5.0\%)\tabularnewline
SD $\rightarrow$  SPYclade &  4 (1.5\%) &  0 (0.0\%)\tabularnewline
SPYclade $\rightarrow$ SDEclade & 32 (12.1\%) & 10 (25.0\%)\tabularnewline
SPYclade $\rightarrow$ SDD & 33 (12.5\%) &  3 (7.5\%)\tabularnewline
SPYclade $\rightarrow$ SD &   3 (1.1\%) &  0 (0.0\%)\tabularnewline
SPYclade $\rightarrow$ SPYclade & 27 (10.2\%) &  6 (15.0\%)\tabularnewline
All types  &  265 (100\%) & 40 (100\%)  \tabularnewline
\hline
\end{tabular}
\label{tab:gene-counts-additive}
\end{table}
%\clearpage{}
%=============================================================================

%=============================================================================
\begin{table}[!ht]
\caption{
{\bf Functional categories enriched in 
recombining transfer events that are inferred using the parsimony-based
approach.}}
\noindent \begin{centering}
\begin{tabular}{ccccl}
\hline 
$p^a$ & $q^b$ & Count$^c$ & GO term & Description \\
\hline 
7e-09 & 3e-06 & 153 & GO:0006412 & translation\\
1e-08 & 3e-06 & 67 & GO:0030529 & ribonucleoprotein complex\\
1e-07 & 2e-05 & 45 & GO:0019843 & rRNA binding\\
2e-06 & 2e-04 & 73 & GO:0003735 & structural constituent of ribosome\\
3e-05 & 3e-03 & 157 & GO:0005840 & ribosome\\
7e-05 & 5e-03 & 221 & GO:0003723 & RNA binding\\
2e-04 & 1e-02 & 11 & GO:0004826 & phenylalanine-tRNA ligase activity\\
3e-04 & 2e-02 & 14 & GO:0015413 & nickel-transporting ATPase activity\\
4e-04 & 2e-02 & 93 & GO:0008982 & protein-N(PI)-phosphohistidine-sugar phosphotransferase activity\\
5e-04 & 2e-02 & 104 & GO:0009401 & phosphoenolpyruvate-dependent sugar phosphotransferase system\\
5e-04 & 2e-02 & 13 & GO:0006265 & DNA topological change\\
5e-04 & 2e-02 & 13 & GO:0003916 & DNA topoisomerase activity\\
5e-04 & 2e-02 & 733 & GO:0005737 & cytoplasm\\
6e-04 & 2e-02 & 10 & GO:0046873 & metal ion transmembrane transporter activity\\
7e-04 & 2e-02 & 12 & GO:0009435 & NAD biosynthetic process\\
8e-04 & 2e-02 & 19 & GO:0016820 & hydrolase activity, acting on acid anhydrides, \\
      &       &    &       &catalyzing transmembrane movement of substances \\
1e-03 & 3e-02 & 20 & GO:0005694 & chromosome\\
2e-03 & 4e-02 & 31 & GO:0006814 & sodium ion transport\\
\hline 
\end{tabular}
\par\end{centering}
\begin{flushleft}
$^a$$P$-values based a Mann-Whitney $U$ test.  Shown are values that
correspond to FDR$^b$  $<0.05$ based on the 
Benjamini-Hochberg method (Methods).\\
$^c$Number of genes in category.\\
\end{flushleft}
\label{tab:go-events-recombining}
\end{table}
\clearpage{}
%=============================================================================

%=============================================================================
\begin{table}[!ht]
\caption{
{\bf Three population parameter estimates from the model-based approach.}
The estimates of the three parameters were based on one of two independent
MCMC chains.  We considered the similar 
results of the two independent chains as confirmation of 
the convergence of the first stage of MCMC.}
\noindent \begin{centering}
\begin{tabular}{ccc}
\hline
& Replicate 1 & Replicate 2\tabularnewline
Parameter & Estimate (IQR$^a$) & Estimate (IQR)\tabularnewline
\hline
$\theta^b$ & 0.081 (0.067, 0.094) & 0.081 (0.067, 0.099)\tabularnewline
$\rho^c$ & 0.012 (0.006, 0.019) & 0.012 (0.006, 0.020)\tabularnewline
$\delta^d$ & 744 (346, 2848) & 723 (348, 2870)\tabularnewline
\hline
\end{tabular}
\par\end{centering}
\begin{flushleft}
$^a$ IQR stands for interquartile range.

$^b$ The median of mutation rate per site weighted over the lengths of blocks.

$^c$ The median of recombination rate per site weighted over the lengths of blocks.

$^d$ The median value for the recombinant tract length weighted over the lengths of blocks.
\end{flushleft}
\label{tab:three}
\end{table}
%\clearpage{}
%=============================================================================

%=============================================================================
\begin{table}[!ht]
\caption{
{\bf Simulation study of the three population parameters.}}
\noindent \begin{centering}
\begin{tabular}{ccc}
\hline
Parameter & True & Estimates (Standard Deviation)\tabularnewline
\hline
$\theta^a$ & 0.081 & 0.081 (0.0011)\tabularnewline
$\rho^b$   & 0.012 & 0.011 (0.0003)\tabularnewline
$\delta^c$ & 744 bp & 919 bp (28)\tabularnewline
\hline
\end{tabular}
\par\end{centering}
\begin{flushleft}
$^a$ Mutation rate per site.

$^b$ Recombination rate per site.

$^c$ Recombinant tract length.
\end{flushleft}
\label{tab:sim-three}
\end{table}
%=============================================================================

\section*{Supplementary Figures}

%=============================================================================
\begin{sidewaysfigure}[!ht]
\begin{center}
\includegraphics[scale=0.3]{figures/mauve}
\end{center}
\caption{\label{fig:mauve}
{\bf The genome alignment of the five genomes.} The top
first and second rows are SDE1 and SDE2 genomes, respectively. The third is SDD
genome. The fourth and fifth rows are SPY1 and SPY2 genomes, respectively.}
\end{sidewaysfigure}
%=============================================================================

%=============================================================================
\begin{figure}[!ht]
\begin{center}
\includegraphics[width=5in]{figures/famsizes}
\end{center}
\caption{{\bf Distribution of gene family sizes.}}
\label{fig:famsizes}
\end{figure}
\clearpage{}
%=============================================================================

%=============================================================================
\begin{figure}[!ht]
\begin{center}
\includegraphics[scale=0.5]{figures/scatter-plot-parameter-1-out-theta}

\includegraphics[scale=0.5]{figures/scatter-plot-parameter-1-out-rho}

\includegraphics[scale=0.5]{figures/scatter-plot-parameter-1-out-delta}
\end{center}
\caption{
{\bf Three scatter plots of mutation rate, recombination
rate, and recombinant tract length for the blocks.} The plus signs are mean values
for each block. The density of each value for each block is depicted by black
clouds.  The red dashed lines are the global block-length weighted median of the
three parameter estimates.}\label{fig:scatter3}
\end{figure}
\clearpage{}%
%=============================================================================

%=============================================================================
\begin{figure}[!ht]
\begin{center}
\includegraphics[width=5in]{figures/compareMowgliCoRecombining}
\end{center}
\caption{
{\bf Plot of recombination intensity against number
of replacing gene transfer.}}
\label{fig:cmpcomowgli}
\end{figure}
\clearpage{}
%=============================================================================

%=============================================================================
\begin{figure}[!ht]
\includegraphics[scale=0.42]{figures/clonalorigin}
\caption{
{\bf Recombinant tree in the ClonalOrigin model.}
(A) A recombinant tree relating the five genomes with four recombinant edges.
(B) A genomic region of size 100 base pairs with subregions affected by the
four recombinant edges. Genomic positions are evenly marked by ticks every 10-th
base pair.  For each edge type and each site, we recorded the sampling frequency
of a recombinant edge of the type at that site.  The posterior probability that
a site experiences influx from a donor branch using a single recombinant tree.
Summing these posteriors over a certain set of edge types, we obtain a measure
of recombination intensity along the genome.  Each segment from b to g are
scored by the number of recombinant edges per site.  (C-G) Local gene trees
induced by the recombinant tree are illustrated for each region covered by
combinations of recombinant edges along the genomic region. For example, the
local gene tree (E) is created by three recombinant edges in the subregion
between 51 base pairs and 70 base pairs. The local gene tree (C) is the same as
the species tree (A) in their topology.  Because there are 105 possible rooted
labeled topologies relating five taxa, we could count topologies.}
\label{fig:clonalorigin}
\end{figure}
\clearpage{}%
%=============================================================================

%=============================================================================
\begin{sidewaysfigure}

\includegraphics[scale=0.6]{figures/sim2-1}

\includegraphics[scale=0.6]{figures/sim2-2}

\caption{
{\bf Simulations of the model-based approach for assessing the power of
recovering recombinant edges using prior recombinant trees.} 
At X-axis labels are shown pairs of species tree branches, for each of which one
interval with an open circle and a closed circle next to it
are displayed.  Intervals with open circle are for 
the ratio of the number of recombinant edges inferred from the 100
simulated data sets under the prior with respect to the expected number of
recombinant edges under the prior.  Closed circles are for 
the ratio of the number of recombinant edges estimated
from the real data with respect to the expected number of recombinant edges
under the prior.  The horizontal lines are at the zero.  The intervals range
from 5\% quantile to 95\% of quantile.}
\label{fig:sim2}
\end{sidewaysfigure}
\clearpage{}%
%=============================================================================

%=============================================================================
\begin{sidewaysfigure}

\includegraphics[scale=0.6]{figures/sim3-1}

\includegraphics[scale=0.6]{figures/sim3-2}

\caption{
{\bf Simulations of the model-based approach for assessing the power of
recovering recombinant edges using posterior recombinant trees.} 
At X-axis labels are shown pairs of species tree branches, for each of which one
interval with an open circle and an closed circle next to it
are displayed.  Intervals with open circle are for 
the ratio of the number of recombinant edges inferred from the 100
simulated data sets with the posterior samples of recombinant trees 
with respect to the expected number of
recombinant edges under the prior.  Closed circles are for 
the ratio of the number of recombinant edges estimated
from the real data with respect to the expected number of recombinant edges
under the prior.  The horizontal lines are at the zero.  The intervals range
from 5\% quantile to 95\% of quantile.}
\label{fig:sim3}
\end{sidewaysfigure}
\clearpage{}%
%=============================================================================

%=============================================================================
\begin{sidewaysfigure}

% \includegraphics[width=6in]{figures/real_vs_null_additive.pdf}
\includegraphics[scale=0.6]{figures/sim4-1}

% \includegraphics[width=4in]{figures/real_vs_null_dup.pdf}
\includegraphics[scale=0.6]{figures/sim4-2}

\caption{
{\bf The power of recovering additive transfers and duplications using simulations.} (top) For each pairs of species tree branches and clades SDE and SPY,
we plot the number of reconstructed additive transfers in the real data 
(circle) normalized
by the true number of transfers in our simulations.  Additive transfers
reconstructed from simulations are also normalized by the true number
and are plotted with 90\% confidence intervals.
(bottom) For each branch in the species tree, we plot the number of 
reconstructed duplications in the real data (circles) normalized by the true
number of duplications in our simulations.  Duplications
reconstructed from simulations are also normalized by the true number
and are plotted with 90\% confidence intervals.
}
\label{fig:mowgli-sim}
\end{sidewaysfigure}
\clearpage{}%
%=============================================================================

%=============================================================================
\begin{figure}
\begin{center}
\includegraphics[width=5in]{figures/calling}
\end{center}
\caption{
{\bf Distinguishing between replacing and additive horizontal gene transfers.}
We consider a transfer ``additive'' if there exists an extant descendant species
(e.g. C and D) that contains both descendants (bold) and
non-descendants (non-bold) of the transferred gene (black circle).}
\label{fig:calling-transfers}
\end{figure}
\clearpage{}%
%=============================================================================

\clearpage{}
\section*{Supplementary Material}

% \subsection*{Preliminary ClonalFrame and ClonalOrigin analysis}
% 
% {\bf IG: we should write here details about the first part of the analysis: data
% preparation, estimating species tree, and three population parameters. TBD. For
% now we have a collection of details extracted from previous version of
% results.}
% \texttt{SC: Do we need this section because the method seems to describe everything enough
% in detail?}
% 
% A genome alignment of the five genomes showed two
% distinct genome structures of SPY and SDE (Figure S\ref{fig:mauve}).
% 
% ClonalFrame estimated a relative strength of
% recombination compared to mutations as $r/m=4.66$ meaning that four times more
% sites had been affected by recombination-driven substitutions than by
% mutation-driven substitutions (Table S\ref{tab:clonalframe}).  This indicated
% potentially significant roles of recombination in the evolutionary process of
% shaping these genomes.  Estimates of population mutation rate ($\theta$),
% recombination rate ($\rho$), and recombinant tract length ($\delta$) were obtained
% for the alignment blocks analyzed (Figure S\ref{fig:scatter3}), and a median of
% the estimates weighted over block lengths was obtained as a global estimate for
% each of the three parameters (Table S\ref{tab:three}).  The local estimates
% typically seemed to be moderately fluctuating around the global estimate, with
% the exception of what appeared to be a recombination hotspot at a region 200
% kilobase pairs from the origin of replication of SDE1.

\subsection*{Summarization of recombinant tree samples}
We describe the four summarizing procedures including tree topology along the
genome, numbers of recombinant edges, probability of recombination, and
recombination intensity using supplementary fig.\ S\ref{fig:clonalorigin}.  
Supplementary figures S\ref{fig:clonalorigin}A 
and S\ref{fig:clonalorigin}B fully specify a
recombinant tree.  First, the recombinant tree induces  five gene tree
topologies along the genome of size being 100 bp. The fraction of gene tree
topology of supplementary figure S\ref{fig:clonalorigin}C is 0.15, and 
that of supplementary figure S\ref{fig:clonalorigin}G is 0.05. 
The fractions of the other three gene tree
topologies of 
supplementary figures S\ref{fig:clonalorigin}D, S\ref{fig:clonalorigin}E, 
and S\ref{fig:clonalorigin}F are all 0.2.  Second, the heatmap values are computed
by counting edges for each of 60 possible types and dividing the number of edges
by prior expected number of recombinant edges given the species tree and the
three global population parameters.  One edge for each of the four types of
recombinant edges is observed in the recombinant tree of Figure
S\ref{fig:clonalorigin}A.  Third, the probability of recombination at a site for
a particular recombinant edge type is the fraction of recombinant trees, in
which the recombinant edge spans the site.  
Supplementary figure S\ref{fig:clonalorigin}B
shows the probability of recombination along the 100 bp genomic positions.
Branch SDE1 received the genomic segments ranging from 11 bp to 100 bp from
branch SDE.  Similarly, branch SDE2 received the segments ranging from 31 bp to
70 bp from branch SPY1.  Each of recombinant trees from a posterior sample would
induce a map of recombination such as 
supplementary figure S\ref{fig:clonalorigin}B.
Recombination maps from samples of recombinant trees can be averaged as shown in
figure \ref{fig:ucsc}.  
Fourth, the recombination intensity of a site is the
number of recombinant edge types at the site 
(supplementary fig.\ S\ref{fig:clonalorigin}B).
Note that it is not the number of recombinant edges.

The recombination intensity allowed us to find genomic regions enriched for
recombination events in general, or regions enriched for recombination events of
certain types (e.g., between SPY and SDE).  
Our simulation study indicated that edges of
certain types were better indicators of recombination events than others.  For
instance, recombinant edges which resulted in a recombinant tree with the same
topology as the species tree appeared to be more prone to spurious sampling.
Such edges were ones where the donor was the same branch, sister branch, or
parent branch of the recipient branch.  Therefore, in some cases we elected to
remove such edges from consideration when measuring recombination intensity.  
% The tree topologies would show more effectively regions where
% recombination occurs along the genome (see subsection of \textbf{UCSC Genome
% Browser}).

\subsection*{Comparison of model-based and parsimony-based inferences on
replacing gene transfer}

The similarity between the two lists of gene categories found to be associated
with replacing gene transfers in the separate analyses -- model-based and
parsimony-based --  indirectly indicated concordance between the two independent
approaches.  In order to obtain a more direct measure of correlation, we used
the set of genes present in both analyses to compute the correlation between
recombination intensity scores and the number of replacing gene transfers
inferred by the parsimony-based approach. We observed a positive correlation of
$0.39$ (P-value $2.2\times10^{-16}$; supplementary fig.\ S\ref{fig:cmpcomowgli})
between the
two measures.  For comparison, the correlation between recombination intensity
and the number of additive gene transfers was 0.03 (P-value 0.18).
%Figure S\ref{fig:cmpcomowglihgt}).

\subsection*{Simulation Studies for the model-based and parsimony-based approaches}
We performed four sets of simulations.  In the first simulation, we 
showed that ClonalOrigin could infer population mutation rate, tract
length, and recombination rate with reasonable accuracy from our data sets of
the five individual bacteria. In the second
simulation, we checked whether recombinant edges were recovered from
simulated data with recombinant trees generated from the prior under the
parameters estimated from our data set. The third
simulation was conducted using simulated data with recombinant trees sampled
from the
posterior distribution.  We used a simulation feature of ClonalOrigin
to simulate the process of mutation and recombination across the five species.
In the fourth simulation, we checked if the parsimony-based approach worked with
our data set.

\textbf{Simulation 1 - three population parameters.}
We describe a procedure of simulations for checking if the three
population parameters were accurately recovered using the model-based approach.
Ten recombinant trees were sampled given the three parameters (table
S\ref{tab:three}) and the species tree (fig.\ \ref{fig:tree5}) set to the
values inferred from the five bacterial genomes.  Each of the recombinant trees
was used to simulate the Jukes-Cantor model for DNA sequence evolution to create
274 blocks of alignments.  The block lengths configuration was the same as that
of the blocks used in the analysis of the five genomes.

% s15
% FIXME: Newick tree in text in a file.
% FIXME: Publish the result and the data set with the code.

\textbf{Simulation 2 - numbers of recombinant edges from prior.}
Because the number of statistics increases compared with Simulation 1, we
increased the sample size from 10 to 100.  The recombinant trees were created
from the prior distribution by adding Recombinant edges to the species tree with
no particular directions of replacing gene transfers between species tree
branches.  We computed the mean number of recombinant edges of each type from
the inference with each of the 100 data sets. Dividing the mean number by the
prior expected number of recombinant edges we obtained the heatmap value.  The
distribution of the 100 heatmap values in the scale of logarithm to base 2 is
shown at supplementary figure S\ref{fig:sim2}.  
We compared this distribution from simulation
with the heatmap values from the real data analysis.  We counted recombinant
edges for the recombinant tree from each iteration of the posterior sample, and
divided the number of edges by the prior expected number of edges. This allowed
us to measure variation among recombinant trees in the posterior sample.  The
distribution of the 1001 heatmap values in the scale of logarithm to base 2 is
also shown next the simulation result at supplementary figure S\ref{fig:sim2}.

% The following was used for checking recombination intensity of genes. We
% decided not to describe it.
% 
% We first sampled ten recombinant trees given the mutation rate, recombination
% rate, and tract length (table S\ref{tab:three}) and the species tree (Figure
% \ref{fig:tree5}), and then created ten independent simulated data sets with
% the Jukes-Cantor model for DNA sequence evolution using each of the recombinant
% trees. This consisted of 100 replicates of data sets.  The recombinant trees
% were created from the prior distribution and recombinant edges were added to the
% species tree with no particular directions of replacing gene transfers between
% species tree branches.  We computed the fraction of the mean number of recombinant
% edges for a pair of donor and recipient species tree branches over the simulated
% 10 recombinant trees against the median of recombinant edges estimated from the
% 10 simulated data sets, and then took the logarithm of the fraction to base 2. We
% also computed 5\% and 95\% quantiles to find the 90\% quantile range for the log
% base 2 fraction.  We expect that the log base 2 fraction would near zero. We also
% wished to see if how the recombinant edge count estimates from the real data set
% deviate from the prior. Instead of using the number of recombinant edges from
% the ten recombinant trees, we used the numbers of recombinant edges that were
% estimated from the real data. 
% 
% We tested if our measure of recombination intensity provided a reliable estimate
% of the true intensity. We artificially annotated our synthetic blocks to genes
% according to the real annotations from SPY1 genome.  For each gene in each of
% the simulated 100 data sets, we plotted the average estimated recombination
% intensity along that gene against the average number of different recombinant
% edge types from the ten simulated recombinant tree.

% s16

% s16
% FIXME 1: IG: RESULTS FROM THIS POINT ON REQUIRE SOME REFINEMENT. SC: Let's see.
% FIXME 2: ``TRUE" TO BE EXPECTED VALUE
% RATHER THAN AVERAGE OVER THE 10 REC.  TREES. SC: OKAY. LET'S FIND THE PRIOR
% EXPECTED NUMBER OF EVENTS. THIS MUST BE GIVEN SOMEWHERE}.  
% FIXME 3: CONSIDER SHOWING DETAILED COMPARISON OF ESTIMATED VALUES TO ONES 
% EXPECTED FROM THE ACTUAL REC. TREES. THESE DON'T LOOK AS NICE, BUT THEY MIGHT 
% BE ABLE TO DEMONSTRATE OUR ABILITY TO DETECT DEPARTURE FROM EXPECTATION. SC:
% THIS MAY BE ABOUT CHANGING AVERAGE TO PRIOR EXPECTED NUMBER OF EVENTS.
% FIXME 4: WE SHOULD PROBABLY CONSIDER ALTERNATIVE WAYS OF USING THESE RESULTS TO 
% ARGUE THAT OUR DIRECTIONAL SIGNAL IS REAL. SC: I MIGHT CONSIDER SIMULATIONS AND
% ARTIFICIALLY DOUBLE OR TRIPLE THE RECOMBINATION EVENTS OF ONE OF TYPES.

\textbf{Simulation 3 - number of recombinant edges from posterior.}
We used 100 recombinant trees sampled from the posterior distribution
of the recombiant tree instead of using simulated recombinant trees.
We created one simulated data set for each of the 100 recombinant trees with
the Jukes-Cantor model for DNA sequence evolution 
using the mutation rate estimate (supplementary table S\ref{tab:three}).
This consisted of 100 replicates of data sets.  
We follow the same analysis as in the simulation 2 
(supplementary fig.\ S\ref{fig:sim3}).

% s17

\textbf{Simulation 4 - Simulation of additive horizontal gene transfers.}
We also performed a simulation analysis to understand the power of 
the parsimony-based pipeline to recover events such as gene duplication 
and additive transfer.  The simulations were also used as a null model
for studying variations in the rate and biases in the direction of
additive transfers.

We implemented a new program that simulates all four events:
duplication, loss, additive transfer, and replacing transfer.  The
generative process begins with a single gene lineage located at the
root of the species tree.  The gene lineage then grows down a species
tree branch until one of the following events occur.

\begin{itemize}
\item If a species tree node is encountered (the root node is the
  first encountered), the gene lineage undergoes {\em speciation} and
  bifurcates into two lineages, each one evolving independently down a
  child branch of the species tree node.  
\item Using a Poisson process at constant rate $D$, a gene lineage
  will undergo {\em duplication} and bifurcate into two lineages, both
  growing independently within the same species branch.  
\item At a constant rate $L$, a gene lineage can be {\em lost} and
  terminate.  
\item At a constant rate $T$, a gene lineage can undergo {\em additive
  transfer}, bifurcating into two lineages, one lineage within
  the current species branch and the second lineage within one
  of the other contemporary species branches chosen uniformly.  Both
  gene lineages continue to evolve independently of one another.  
\item At a constant rate $R$, a gene lineage $a$ can undergo {\em replacing
  transfer}.  From the other contemporary species branches, a second
  gene lineage $b$ (if it exists) is chosen to be replaced.  The gene
  lineage $b$ is terminated and lineage $a$ is bifurcated, one lineage $a_1$
  remaining in the current species and the other $a_2$ transferring to the
  species branch of lineage $b$.  Both lineages $a_1$ and $a_2$ continue
  to evolve independently.
\item Lastly, if a gene lineage encounters a species tree leaf, the
  gene lineage terminats and is denoted an {\em extant gene copy}.
\end{itemize}

As a final step, any gene lineage that terminates before the species tree
leaves, is declared extinct and is pruned from the gene tree.

Given a gene tree generated by this process, we simulate sequence
evolution using a continuous-time Markov process (e.g. HKY)
parameterized to match the base frequency and transition/transversion
ratio observed in the real data.  This results in a gapless simulated
gene alignment.

In our simulation analysis, we used our simulation program within a
larger analysis pipeline that followed the same procedure we have used
to analyzed the real dataset, thus giving an accurate view of the
power of the entire pipeline.  In our simulation pipeline, we
simulated 10,000 gene alignments.  For each alignment we reconstructed 
gene trees using RAxML, reconciled them using Mowgli, and classified
additive and replacing gene transfers using the procedure described in the
main text.

To create a genome-wide sample of gene trees we randomly chose 2314
gene trees (the same number of real families) from the 10,000
reconstructed.  To study the distribution of this tree set, we
resampled with replacement 1000 times (This resampling strategy was
chosen for computational efficiency).  This gave us the following
samples: (1) a distribution of the ``true'' simulated gene trees along
with their event counts, (2) a distribution of the reconstructed
simulated gene trees and event counts, and (3) the original set of
gene trees reconstructed from real data.  Differences between datasets
(1) and (2) indicate the power of our pipeline to accurately infer
each of the event types, and differences between datasets (2) and (3)
indicate unusual deviations of the real data from the null model
(i.e. variations in event rates or biases in transfer direction).

\subsection*{Simulation-based support for biased directionality
in replacing gene transfer between SPY and SDE}

We conducted a series of experiments on the simulated data to check the reliability
of estimates obtained by the model-based approach. We found that the population
parameters obtained in the first phase were reliable, with the exception that
the recombinant tract length was over-estimated as this was observed
by \citealp{Didelot2010} (Simulation 1). Our experiments also showed that the recombination
intensity used in our functional category association tests had fairly good
correlation between the true value and the simulated ones (P-value of 0.79).  
% (Figure S\ref{fig:ri1}). 
One thing we specifically wished to test in these experiments
was whether the model-based approach reliably detected deviations of recombinant
edge counts from their expectation, as observed in the heatmap of Figure
\ref{fig:Heatmap-of-transfers}. Our major concern was that these deviations
were not directly accounted for by the probabilistic model of ClonalOrigin, but
rather indirectly observed by summary of the posterior sample of recombinant
trees.

We simulated the model to generate a hundred simulated data sets using a
collection of recombinant trees sampled from the prior distribution defined by
the species tree and three population parameters, as estimated from our genomic
data (Simulation 2).  We then repeated the inference procedure of the second stage of
ClonalOrigin on each data set and, as in the data analysis, and counted
for each donor-recipient branch pair the average number of recombinant edges.
By normalizing the counts by the prior expected count we obtained intervals of
base 2 logarithms of the ratio of the observed number of recombinant edges to
the prior expectionation (supplementary fig.\ S\ref{fig:sim2}). 
The comparison showed a good
concordance of the estimated counts and the prior expected counts for 
the topology-altering edge types.
For each edge type, we also plotted in supplementary figure S\ref{fig:sim3} the
log-ratio of the count estimated for it in our data analysis and the expected
count; these are the values making up
the heatmap of Figure \ref{fig:Heatmap-of-transfers}. Comparing the counts
estimated in the data analysis and the counts estimated from the simulated data
allowed to evaluate the significance of the divergence from expectation observed
in the heatmap.  Recall the heatmap indicated an excess of recombinant edges
between the SPY and SDE clades in both directions. Our simulations indicated
that the excess observed in the direction SDE to SPY was much less significant,
as the counts estimated from real data often fell very close to the
corresponding counts estimated from simulated data. The excess observed in the
direction SPY to SDE appeared not only to be larger, but also more significant.
This further supported our claim for the biased directionality in replacing gene
transfer between these two species.

We used simulated data (Simulation 3) to test not only the ability of the ClonalOrigin
sample to \textit{detect} deviations from expected counts of recombinant edges, but
also its ability to reliably \textit{quantify} this deviation.  For this purpose we
generated a hundred simulated data sets using a collection of recombinant trees
sampled from the posterior sample of ClonalOrigin on the genomic data.  The
difference between the two simulated batches was that in the first one, the
recombinant trees used represent the prior distribution, and in the second one,
they represented the posterior distribution, which we knew deviated from the
prior. Again, we recorded the recombinant edge counts obtained by ClonalOrigin
and normalized these counts by the prior expected counts (intervals with closed
circles in supplementary figure S\ref{fig:sim3}). We expected these counts to be
in better agreement with
the counts observed in the data analysis. For recombinant edges that showed a
defficiency in our data analysis (e.g., from the SPY clade to SDD), the counts
appeared to fit well with the count estimated from genomic data. However, for
recombinant edges that showed an excess in our data analysis (e.g., from the SPY
clade to the SDE clade), the counts appeared to be significantly influenced by the
prior. We noted that the influence of the prior appeared to be even stronger for
topology-preserving recombinant edges. All things considered, we found that defficiency of
a certain recombinant edge type was reliably estimated whereas excesses were 
typically underestimated. Therefore, we concluded that the actual extent to which
replacing transfers between SPY and SDE \textit{preferred} the direction from SPY to
SDE was possibly somewhat underestimated in our analysis.

\end{document}