### 2011-12-13 doc/supplementary.tex

\documentclass[12pt]{article}

\usepackage{times}

% amsmath package, useful for mathematical formulas
\usepackage{amsmath}
% amssymb package, useful for mathematical symbols
\usepackage{amssymb}

% graphicx package, useful for including eps and pdf graphics
% include graphics with the command \includegraphics
\usepackage{graphicx}

% cite package, to clean up citations in the main text. Do not remove.
%\usepackage{cite}

\usepackage{color}

% Use doublespacing - comment out for single spacing
\usepackage{setspace}

% To rotate tables using sidewaystable instead of just table.
\usepackage{rotating}

% FIXME: Remove this package when submitting the manuscript.
\usepackage{url}
\usepackage{xr}

\usepackage{natbib}
\usepackage{../../latex/sty/genres}

\externaldocument{manuscript}
% Text layout
\oddsidemargin 0in
\evensidemargin 0in
\topmargin -.5in
\textwidth 6.5in
\textheight 9in
%\topmargin 0.0cm
%\oddsidemargin 0.5cm
%\evensidemargin 0.5cm
%\textwidth 16cm 
%\textheight 21cm

% Bold the 'Figure #' in the caption and separate it with a period
% Captions will be left justified
\usepackage[labelfont=bf,labelsep=period,justification=raggedright]{caption}

% Remove brackets from numbering in List of References
\makeatletter
\renewcommand{\@biblabel}[1]{\quad#1.}

\providecommand{\\}{\\}
\newcommand{\lyxdot}{.}
\newcommand{\comment}[1]{}

\makeatother

\renewcommand{\thefootnote}{\textit{\alph{footnote}}}

% Leave date blank
\date{}

\pagestyle{myheadings}
%% ** EDIT HERE **
\markright{Horizontal gene transfer in Streptococcus}

\begin{document}

\setcounter{secnumdepth}{-1}

\noindent \begin{LARGE}{\bf Replacing and additive horizontal 
    \\[0.7ex] 
gene transfer in {\em Streptococcus}}\end{LARGE}\\[3ex]
{\bf Sang Chul Choi, Matthew D.\ Rasmussen, Melissa J.\ Hubisz, \\
Ilan Gronau, Michael J.\ Stanhope, and Adam Siepel}\\[0.5ex]

\tableofcontents

\listoftables

\listoffigures

\doublespacing

\section{Supplementary Text}

\subsection{Recombinant tree summaries}

We use a recombinant tree portrayed in supplementary figure
\ref{fig:clonalorigin} to demonstrate the various summaries we use in our
analysis of recombinant trees sampled by ClonalOrigin. This recombinant tree is
augmented by four recombinant edges denoted e1 to e4 of four different edge
types (supplementary fig.\ \ref{fig:clonalorigin}A): SDE-to-SDE1, SPY1-to-SDE2,
SPY-to-SDD, and ROOT-to-SPY2. Edges e2, e3 and e4 are topology-altering ones
whereas edge e1 is topology-preserving. Each recombinant edge is associated with
a contiguous genomic segment  (supplementary fig.\ \ref{fig:clonalorigin}B),
inducing a series of six different gene trees, T1 through T6, along the genome
(supplementary fig.\ \ref{fig:clonalorigin}C).  Gene trees T1 and T2 are of the
same tree topology, but the other trees are of distinct topologies with the
following genomic frequencies: 0.3 (T1 and T2), 0.2 (T3, T4, and T5), and 0.1
(T6).  The recombination
intensity at a given site is defined by the number of recombinant edge types
present at that site; it can be computed from the map of recombinations along
the genome (bottom of supplementary fig.\ \ref{fig:clonalorigin}B). In this
example the recombination intensity ends up being the number of recombinant
edges present at the site.
% , but that does not generally have to be the case
% (there might be several edges of the same type across a given site).

\subsection{Simulations performed for the model-based analysis}

We performed several simulation suites in order to validate the performance of
ClonalOrigin on our data, and found significant trends. We used the simulation
component of ClonalOrigin to produce these simulations. Each simulated data set
was constructed to be of the same length distribution as used in our data
analysis. In each simulation, sequence data was generated according to the
Jukes-Cantor substitution model 
\citep{Jukes1969}
using a designated recombinant tree defined over
a clonal frame identical to the one inferred for our data (fig.\
\ref{fig:tree5}).

We started by testing accuracy of the estimates of the three population
parameters: population mutation rate ($\theta$), recombination rate ($\rho$),
and average recombinant tract length ($\delta$). We generated ten simulated sets
using ten different recombinant trees sampled from the prior distribution
given the clonal frame and the values of the three population parameters
estimated in data analysis: $\theta=0.081$, $\rho=0.012$, and $\delta=744$ bp.
Estimates of the mutation and recombination rates appeared to be accurately
inferred with average global estimates of $0.081\pm 0.0011$, and $0.011\pm
0.0003$, respectively. The average recombinant tract length were over-estimated at
$919\pm28$ bp as previously observed by \citealp{Didelot2010}.

In order to evaluate significant trends in our data analysis, we generated a
more extensive suite with 100 data sets simulated using 100 recombinant trees
generated from the model prior as the previous simulation setup.  As in the real
data analysis, we compared the average number of recombinant edges of each type
to the corresponding prior expected number (supplementary fig.\
\ref{fig:sim2}). The ratios between inferred and expected edge counts ranged
from 0.64 to 1.6 for all edge types across all simulations. We observed no
notable bias in the estimates. The counts estimated from the 100 simulated sets
were also used to assign empirical $p$-values to the deviations from the prior
portrayed in the heatmap of figure \ref{fig:Heatmap-of-transfers}. We also used
simulations under the prior model to assess the accuracy of our estimates for
recombination intensity. We found a high correlation of 0.79 between inferred
intensities and true ones when restricting the intensity measure to
topology-altering recombinant edges, and a lower correlation of 0.63 for the
total intensity. Therefore, we used the topology-altering intensity for our GO
enrichment tests.

Additionally, we constructed 100 simulated data sets using recombinant trees
sampled from the posterior distribution (selecting every tenth tree in the
posterior sample generated by ClonalOrigin in the real data analysis). Using
these data we were able to assess the effect the prior had on the estimates
produced by ClonalOrigin.
% when the data does not fit the prior. 
Summarizing average recombinant edge counts across
the 100 data sets we observed that, as
expected, these counts were in better agreement with the counts inferred in our
data analysis than were counts inferred from data simulated under the
prior (supplementary fig.\ \ref{fig:sim3}). 
The agreement was notably better for edge types where the inferred
count was less than the prior expected count (e.g., edges from the SPY clade to
SDD). The prior distribution did appear to have an effect on estimated counts
associated with enriched edge types, leading to an under-estimation of the
enrichment level. We concluded, therefore, that the enrichment observed for
SPY/SDE replacing gene transfer was likely under-estimated in our analysis.

\subsection{Simulations performed for the parsimony-based analysis}

We performed separate simulations to understand the power of 
the parsimony-based pipeline to recover events such as gene duplication 
and additive transfer.  The simulations were also used as a null model
for assessing significance of the deviation from prior expectation.

We implemented a new program that simulated all of the four events including
duplication, loss, additive transfer, and replacing transfer.  The program
assumes that each of these events occurs at a (different) constant rate, 
generating gene trees given a species tree and specified rates (see description
below). Using this simulation program, we generated $10^4$ gene trees, setting
the rates of the four types of events to the average rates inferred in our data
analysis (i.e., number of events divided by total tree length). We then
simulated sequence evolution using the HKY substitution model 
\citep{Hasegawa1985}
with base
frequencies and transition/transversion ratio as observed in the real data,
resulting in gapless alignments.
For each of the generated alignments, we performed the entire analysis pipeline,
inferring the gene tree using RAxML, reconciling it using Mowgli, and classifying
additive transfers as described in the main text. This allowed
us to assess the power of the entire pipeline. To create a genome-wide sample of
gene trees we wanted to generate sets of 2,314 gene trees (the same number of
families in our data analysis). Hence, for computational efficiency, we
resampled with replacement 1,000 subsets of 2,314 trees from the original set of
10,000. Comparison of event counts in the true (simulated) gene trees with the
ones reconstructed from the simulated data 
allowed us to assess the accuracy of our pipeline of parsimony-based approach
(supplementary fig.\ \ref{fig:mowgli-sim}).
Comparison of the counts inferred in our data analysis with the
ones inferred from simulated data allowed us to assess significance of deviation
from the null model, as portrayed in figure \ref{fig:Heatmap-of-transfers}B.

\subsection{Program simulating duplication, loss, transfer, and recombination}

The program takes as input a species tree and four rate parameters $D,L,T,$ 
and $R.$
It generates gene trees by starting with a single gene lineage located at the
root of the species tree, and then traversing down branches of the species
tree, allowing for the following events: 
\begin{itemize}
\item If a species tree node is encountered (the root node is the
  first encountered), the gene lineage undergoes {\em speciation} and
  bifurcates into two lineages, each one evolving independently down a
  child branch of the species tree node.  
\item Using a Poisson process at constant rate $D$, a gene lineage
  undergoes {\em duplication} and bifurcates into two lineages, both
  growing independently within the same species branch.  
\item At a constant rate $L$, a gene lineage can be {\em lost} and
  terminate.  
\item At a constant rate $T$, a gene lineage can undergo {\em additive
  transfer}, bifurcating into two lineages, one lineage within
  the current species branch and the second lineage within one
  of the other contemporary species branches chosen uniformly.  Both
  gene lineages continue to evolve independently of one another.  
\item At a constant rate $R$, a gene lineage $a$ can undergo {\em replacing
  transfer}.  From the other contemporary species branches, a second
  gene lineage $b$ (if it exists) is chosen to be replaced.  The gene
  lineage $b$ is terminated and lineage $a$ is bifurcated, one lineage $a_1$
  remaining in the current species and the other $a_2$ transferring to the
  species branch of lineage $b$.  Both lineages $a_1$ and $a_2$ continue
  to evolve independently.
\end{itemize}

If a gene lineage encounters a species tree leaf, the gene lineage terminates
and is denoted an {\em extant gene copy}. Any gene lineage that terminates
before the species tree leaves, is declared extinct and is pruned from the gene
tree.

\subsection{Replacing gene transfers inferred by the parsimony-based method.}

We studied transfers inferred by Mowgli that were not identified as additive
transfers by our criterion. Majority of these inferred
transfer events were expected to be of replacing type. We set out to compare the
inference with our model-based analysis of replacing gene transfers. First, we
compared the set of Gene Ontology (GO) categories associated with the two sets
of inferred transfers. We found that 8 of the 14 categories associated with
replacing gene transfer in the model-based analysis were also associated with
the parsimony-based inferred replacing gene transfer (supplementary tables
\ref{tab:functional} and \ref{tab:go-events-recombining}).  Some disagreement
was expected because of inherent differences of the two inference methods.  We
also computed the correlation between recombination intensity scores and the
number of replacing gene transfers inferred by the parsimony-based approach over
the set of genes present in both analyses.  We found a significant positive
correlation of $0.39$ ($p=2.2\times10^{-16}$) between the two measures.

\renewcommand*{\refname}{Literature Cited}
\bibliographystyle{../../latex/bst/mbe}
\bibliography{siepel-strep}

\clearpage{}\setcounter{figure}{0}
\setcounter{table}{0}

\makeatletter 
\renewcommand{\thefigure}{S\@arabic\c@figure} 
\renewcommand{\thetable}{S\@arabic\c@table} 
\renewcommand{\figurename}{Supplementary Figure}
\renewcommand{\tablename}{Supplementary Table}
\makeatother

\section{Supplementary Tables}

%=============================================================================
\begin{table}[!ht]
\caption[The six bacterial genomes]{
{\bf The six bacterial genomes.}
Each bacterial genome is referred to as the name at the first column.}
\noindent \begin{centering}
\begin{tabular}{cccc}
\hline 
Name & NCBI accession & Size (base pairs) & Reference\\
\hline
SDE1 & CP002215 & 2159491 & Suzuki et al. (2011)\\
SDE2 & NC\_012891 & 2106340 & Shimomura et al. (2011)\\
SDD & CM001076 & 2141837 & Suzuki et al. (2011)\\
SPY1 & NC\_004070 & 1900521 & Beres et al. (2002)\\
SPY2 & NC\_008024 & 1937111 & Beres et al. (2006)\\
SEE & NC\_012471 & 2253793 & Holden et al. (2009)\\
\hline
\end{tabular}
\par\end{centering}
\label{tab:genome}
\end{table}
%=============================================================================

%=============================================================================
\begin{table}[!ht]
\caption[Posterior probability of gene tree topologies]{
{\bf Posterior probability of gene tree topologies.} Frequencies of the eight
most frequent topologies in Newick format are shown.}
\noindent \begin{centering}
\begin{tabular}{cc}
\hline 
Gene tree topologies & Posterior probability\\
\hline
(((SDE1,SDE2),SDD),(SPY1,SPY2)) & 66.7\%\\
(((SDE1,SDE2),(SPY1,SPY2)),SDD) & 9.3\%\\
((SDE1,SDE2),(SDD,(SPY1,SPY2))) & 3.5\%\\
(((SDE1,(SDE2,SDD)),(SPY1,SPY2)) & 1.5\%\\
(((SDE1,SDD),SDE2),(SPY1,SPY2)) & 1.3\%\\
((SDE1,SDD),(SDE2,(SPY1,SPY2))) & 1.2\%\\
((((SDE1,SDE2),SPY1),SPY2),SDD) & 1.0\%\\
((SDE2,SDD),(SDE1,(SPY1,SPY2))) & 1.0\%\\
\hline
\end{tabular}
\par\end{centering}
\label{tab:Gene-tree-topologies}
\end{table}
%=============================================================================

%=============================================================================
\begin{table}[!ht]
\caption[The observed values of recombinant edges]{
{\bf The observed values of recombinant edges.}}  
\noindent \begin{centering}
\begin{tabular}{cccccccccc}
\hline
Edge$^a$ & SDE1 & SDE & SDE2 & SD & SDD & ROOT & SPY1 & SPY & SPY2 \\
\hline
SDE1&8.896 &0.00$^b$ &17.712  & 0.00 & 12.872   & 0 &12.59  &  0.00  &12.53\\
SDE &44.626& 136.49& 42.088 &  0.00& 227.061  &  0& 73.39 &  99.95 & 65.25\\
SDE2&18.099&   0.00&  8.393 &  0.00&  12.360  &  0& 11.97 &   0.00 & 10.36\\
SD  &9.671 &101.74 & 9.464  &31.26 &115.796   & 0 &13.68  & 98.38  &12.86\\
SDD &45.539& 168.58& 49.513 &  0.00& 194.706  &  0& 43.09 &  67.14 & 47.65\\
ROOT&43.553& 602.67& 42.337 &765.53& 802.813  &  0& 98.86 &2146.01 &103.60\\
SPY1&40.800&  11.46& 40.973 &  0.00&   4.371  &  0& 21.15 &   0.00 & 42.35\\
SPY &57.618& 262.08& 66.594 & 30.92& 163.518  &  0& 83.56 & 207.00 & 85.10\\
SPY2&38.837&  11.90& 42.000 &  0.00&   5.111  &  0& 42.45 &   0.00 & 21.19\\
\hline
\end{tabular}
\par\end{centering}
\begin{flushleft}
$^a$ Rows and columns indicate the donor and the recipient branches,
respectively.\\
$^b$ Zeros are given for branch pairs that are impossible under the model-based
approach.\\
\end{flushleft}
\label{tab:obsheatmap}
\end{table}
%=============================================================================

%=============================================================================
\begin{table}[!ht]
\caption[Heatmap values for replacing gene transfer]{
{\bf Logarithm base 2 of the ratios of the inferred to the expected number of
recombinant edges.} This table lists the values graphically displayed in the
heatmap of figure \ref{fig:Heatmap-of-transfers}A.}
\noindent \begin{centering}
\begin{tabular}{cccccccccc}
\hline
Edge$^a$ & SDE1 & SDE & SDE2 & SD & SDD & ROOT & SPY1 & SPY & SPY2 \\
\hline
SDE1& 0.176  &   NA$^b$  &1.170  &    NA & 0.7564  & NA & 0.67 &    NA & 0.663\\
SDE & -0.118 & 0.081 &-0.202 &     NA&  0.3395 &  NA& -0.15&  0.045& -0.323\\
SDE2& 1.201  &   NA  &0.092  &    NA & 0.6978  & NA & 0.60 &    NA & 0.387\\
SD  & -0.629 & 0.022 &-0.661 &0.00670&  0.0078 &  NA& -0.94& -0.299& -1.030\\
SDD & -0.317 & 0.371 &-0.197 &     NA&  0.0752 &  NA& -1.06& -0.525& -0.915\\
ROOT& -1.745 &-0.698 &-1.786 &0.23039& -0.4854 &  NA& -1.37&  0.497& -1.306\\
SPY1& 1.245  &1.883  &1.251  &    NA &-2.1997  & NA & 0.06 &    NA & 1.062\\
SPY & 0.089  &0.200  &0.298 &0.00098 &-0.7903  & NA &-0.19 &-0.058 &-0.159\\
SPY2& 1.174  &1.938  &1.287  &    NA &-1.9740  & NA & 1.07 &    NA & 0.063\\
\hline
\end{tabular}
\par\end{centering}
\begin{flushleft}
$^a$ Rows and columns indicate the donor and the recipient branches,
respectively.\\
$^b$ NA's are given for the branch pairs that are impossible under the model.\\
\end{flushleft}
\label{tab:heatmap}
\end{table}
%=============================================================================

%=============================================================================
\begin{table}[!ht]
\caption[Number of genes experiencing replacing transfer]{
{\bf Number of genes experiencing replacing transfer inferred from ClonalOrigin.} 
Numbers of genes with an average posterior probability
greater than 0.6 for recombinant edge types across their sites are 
shown together with
ratios relative to the total number of genes across all edge types.
The same is done for the subset of virulence genes.}
\label{tab:gene-counts-replacing}
\noindent \begin{centering}
\begin{tabular}{lll}
\hline
Transfer type$^a$ & Gene count & Subset of virulence genes
\\
\hline
SDEclade$^b$ $\rightarrow$ SDD & 43 (33.1\%) & 5 (50\%) \\
SDEclade $\rightarrow$ SPYclade$^c$ & 11 (8.5\%) & 1 (10\%) \\
SDD $\rightarrow$ SDEclade & 1 (0.8\%)  & 0 (0\%) \\
SPYclade $\rightarrow$ SDEclade & 59 (45.4\%) & 3 (30\%) \\
SPYclade $\rightarrow$ SPYclade & 16 (12.3\%) & 1 (10\%) \\
\hline
All types & 130  (100\%) & 10   ( 100\%)\\
% &  (11.8\% of total number of genes) & (13\% of virulence genes)\\
\hline
\end{tabular}
\par\end{centering}
\begin{flushleft}
$^a$Edge types not shown have a zero count. 
Recombinant edges from the root branch are not considered.\\
$^b$SDEclade denotes three branches of SDE1, SDE2 and SDE in the clonal frame.\\
$^c$SPYclade denotes three branches of SPY1, SPY2 and SPY in the clonal frame.\\
\end{flushleft}
\end{table}
%\clearpage{}
%=============================================================================

%=============================================================================
\begin{table}[!ht]
\caption[Number of inferred additive transfer events]{
{\bf Number of inferred additive transfer events of different types.}  
Numbers of inferred additive transfer events of each type across
all genes and subset of virulence genes 
are
shown together with ratios relative to the total number of inferred
additive transfers.}
\label{tab:gene-counts-additive}
%
\noindent \begin{centering}
\begin{tabular}{lll}
\hline
Transfer type$^a$ & Number of events & Subset of virulence genes \\
\hline
SDEclade$^b$ $\rightarrow$ SDEclade & 28 (10.6\%) &  5 (12.5\%)\\
SDEclade $\rightarrow$ SDD & 64 (24.2\%) & 2 (5.0\%)\\
SDEclade $\rightarrow$ SPYclade$^c$ & 28 (10.6\%) &  7 (17.5\%)\\
SDD $\rightarrow$ SDEclade & 29 (10.9\%) &  5 (12.5\%)\\
SDD $\rightarrow$ SPYclade & 17 (6.4\%) &  2 (5.0\%)\\
SD $\rightarrow$  SPYclade &  4 (1.5\%) &  0 (0.0\%)\\
SPYclade $\rightarrow$ SDEclade & 32 (12.1\%) & 10 (25.0\%)\\
SPYclade $\rightarrow$ SDD & 33 (12.5\%) &  3 (7.5\%)\\
SPYclade $\rightarrow$ SD &   3 (1.1\%) &  0 (0.0\%)\\
SPYclade $\rightarrow$ SPYclade & 27 (10.2\%) &  6 (15.0\%)\\
\hline
All types  &  265 (100\%) & 40 (100\%)  \\
\hline
\end{tabular}
\par\end{centering}
\begin{flushleft}
$^a$Edge types not shown have a zero count.\\
$^b$SDEclade denotes three branches of SDE1, SDE2 and SDE in the clonal frame.\\
$^c$SPYclade denotes three branches of SPY1, SPY2 and SPY in the clonal frame.\\
\end{flushleft}
\end{table}
%=============================================================================

%=============================================================================
% /Users/goshng/Documents/Projects/Mauve/output/cornellf/3/run-analysis/significant-1.txt
\begin{table}[!ht]
\caption[Functional category enrichments for replacing transfers with ClonalOrigin]{
{\bf Functional category enrichments for replacing gene transfers
inferred by the model-based approach$^a$}}
\vspace{1ex}
\noindent \begin{centering}
\begin{tabular}{ccccp{3.5in}}
\hline 
$p^b$ & $q^c$ & Count$^d$ & GO term & Description\\
\hline 
2e-06 & 2e-04 &  38 & GO:0009401 & phosphoenolpyruvate-dependent sugar phosphotransferase system\\
3e-06 & 2e-04 &  33 & GO:0008982 & protein-N(PI)-phosphohistidine-sugar phosphotransferase activity\\
5e-05 & 3e-03 &  22 & GO:0006633 & fatty acid biosynthetic process\\
9e-05 & 4e-03 &  21 & GO:0008610 & lipid biosynthetic process\\
3e-04 & 9e-03 &  45 & GO:0008643 & carbohydrate transport\\
5e-04 & 1e-02 &  82 & GO:0006412 & translation\\
6e-04 & 1e-02 &  13 & GO:0006814 & sodium ion transport\\
1e-03 & 2e-02 &  40 & GO:0030529 & ribonucleoprotein complex\\
1e-03 & 2e-02 &  12 & GO:0031402 & sodium ion binding\\
1e-03 & 2e-02 & 244 & GO:0008152 & metabolic process\\
2e-03 & 3e-02 & 378 & GO:0016740 & transferase activity\\
2e-03 & 3e-02 &  39 & GO:0003735 & structural constituent of ribosome\\
2e-03 & 3e-02 &  26 & GO:0019843 & rRNA binding\\
4e-03 & 4e-02 &  66 & GO:0005840 & ribosome\\
\hline 
\end{tabular}
\par\end{centering}
\begin{flushleft}
$^a$Based on recombination intensities restricted to topology-altering
recombinant edges (Methods).\\
$^b$$P$-values based a Mann-Whitney $U$ test.\\
$^c$Values corresponding to FDR$<0.05$ based on the Benjamini-Hochberg method (Methods).\\
$^d$Number of genes in category.\\
\end{flushleft}
\label{tab:functional}
\end{table}
%=============================================================================

%=============================================================================
\begin{table}[!ht]
\caption[Functional categories enriched in replacing transfers with Mowgli]{
{\bf Functional categories enriched in 
replacing transfer events inferred by the parsimony-based approach$^a$}}
\noindent \begin{centering}
\begin{tabular}{ccccp{3.5in}}
\hline 
$p^b$ & $q^c$ & Count$^d$ & GO term & Description \\
\hline 
7e-09 & 3e-06 & 153 & GO:0006412 & translation\\
1e-08 & 3e-06 & 67 & GO:0030529 & ribonucleoprotein complex\\
1e-07 & 2e-05 & 45 & GO:0019843 & rRNA binding\\
2e-06 & 2e-04 & 73 & GO:0003735 & structural constituent of ribosome\\
3e-05 & 3e-03 & 157 & GO:0005840 & ribosome\\
7e-05 & 5e-03 & 221 & GO:0003723 & RNA binding\\
2e-04 & 1e-02 & 11 & GO:0004826 & phenylalanine-tRNA ligase activity\\
3e-04 & 2e-02 & 14 & GO:0015413 & nickel-transporting ATPase activity\\
4e-04 & 2e-02 & 93 & GO:0008982 & protein-N(PI)-phosphohistidine-sugar phosphotransferase activity\\
5e-04 & 2e-02 & 104 & GO:0009401 & phosphoenolpyruvate-dependent sugar phosphotransferase system\\
5e-04 & 2e-02 & 13 & GO:0006265 & DNA topological change\\
5e-04 & 2e-02 & 13 & GO:0003916 & DNA topoisomerase activity\\
5e-04 & 2e-02 & 733 & GO:0005737 & cytoplasm\\
6e-04 & 2e-02 & 10 & GO:0046873 & metal ion transmembrane transporter activity\\
7e-04 & 2e-02 & 12 & GO:0009435 & NAD biosynthetic process\\
8e-04 & 2e-02 & 19 & GO:0016820 & hydrolase activity, acting on acid anhydrides, \\
      &       &    &       &catalyzing transmembrane movement of substances \\
1e-03 & 3e-02 & 20 & GO:0005694 & chromosome\\
2e-03 & 4e-02 & 31 & GO:0006814 & sodium ion transport\\
\hline 
\end{tabular}
\par\end{centering}
\begin{flushleft}
$^a$Transfer events inferred by Mowgli are considered to be replacing if they do
not pass our criterion for additive transfers (see supplementary fig.\
\ref{fig:calling-transfers}). \\
$^b$$P$-values based a Mann-Whitney $U$ test. \\
$^c$Values corresponding to FDR$<0.05$ based on the Benjamini-Hochberg method (Methods).\\
$^d$Number of genes in category.\\
\end{flushleft}
\label{tab:go-events-recombining}
\end{table}
\clearpage{}
%=============================================================================

\clearpage{}
\section{Supplementary Figures}

%=============================================================================
\begin{sidewaysfigure}[!ht]
\begin{center}
\includegraphics[scale=0.3]{figures/mauve}
\end{center}
\caption[The genome alignment of the five genomes]{\label{fig:mauve}
{\bf The genome alignment of the five genomes.} The top
first and second rows are SDE1 and SDE2 genomes, respectively. The third is SDD
genome. The fourth and fifth rows are SPY1 and SPY2 genomes, respectively.}
\end{sidewaysfigure}
%=============================================================================

%=============================================================================
\begin{figure}[!ht]
\begin{center}
\includegraphics[width=5in]{figures/famsizes}
\end{center}
\caption[Distribution of gene family sizes]{
{\bf Distribution of gene family sizes.}}
\label{fig:famsizes}
\end{figure}
\clearpage{}
%=============================================================================

%=============================================================================
\begin{figure}[!ht]
\includegraphics[scale=0.42]{figures/clonalorigin}
\caption[Recombinant tree in the ClonalOrigin model]{
{\bf Recombinant tree in the ClonalOrigin model.}
(A) A recombinant tree with a clonal frame over our five samples, and four
recombinant edges.  (B) Each recombinant edge is mapped to a genomic region
(each dotted vertical line corresponds to 20\% of the genome). Recombination
intensities (number of edge types present) and gene tree labels are plotted at
the bottom of the map. (C) The local gene trees induced by the recombinant tree. 
Bold edges portray branches of the clonal frame, and dashed edges portray
recombinant edges. See section \textbf{Recombinant tree summaries} in
\textbf{Supplementary Text}.}
\label{fig:clonalorigin}
\end{figure}
\clearpage{}%
%=============================================================================

%=============================================================================
\begin{figure}
\begin{center}
\includegraphics[width=5in]{figures/calling}
\end{center}
\caption[Distinguishing between replacing and additive gene HGT]{
{\bf Distinguishing between replacing and additive horizontal gene transfers.}
We consider a transfer ``additive'' if there exists an extant descendant species
(e.g. C and D) that contains both descendants (bold) and
non-descendants (non-bold) of the transferred gene (black circle).}
\label{fig:calling-transfers}
\end{figure}
\clearpage{}%
%=============================================================================

%=============================================================================
\begin{figure}[!ht]
\begin{center}
\includegraphics[scale=0.5]{figures/scatter-plot-parameter-1-out-theta}

\includegraphics[scale=0.5]{figures/scatter-plot-parameter-1-out-rho}

\includegraphics[scale=0.5]{figures/scatter-plot-parameter-1-out-delta}
\end{center}
\caption[Three scatter plots of population parameters]{
{\bf Three scatter plots of mutation rate, recombination
rate, and average recombinant tract length for the blocks.} The plus signs are mean values
for each block. The density of each value for each block is depicted by black
clouds.  The red dashed lines are the global block-length weighted median of the
three parameter estimates.}\label{fig:scatter3}
\end{figure}
\clearpage{}%
%=============================================================================

%=============================================================================
\begin{sidewaysfigure}

\includegraphics[scale=0.6]{figures/sim2-1}

\includegraphics[scale=0.6]{figures/sim2-2}

\caption[ClonalOrigin simulation using prior of recombinant trees]{
{\bf Simulations of the model-based approach using recombinant trees sampled from the prior distribution.} 
For each recombinant edge type (X-axis), we record the ratio between the average
count of edges as inferred in our data analysis (closed circle; see also
supplementary tables \ref{tab:obsheatmap} \& \ref{tab:heatmap}), and as inferred
across 100 data sets simulated according to the prior distribution (intervals
with open circles). The open circle corresponds to the median across the
different sets, and the intervals correspond to the range between the  5\% and
95\% quantiles.}
\label{fig:sim2}
\end{sidewaysfigure}
\clearpage{}%
%=============================================================================

%=============================================================================
\begin{sidewaysfigure}

\includegraphics[scale=0.6]{figures/sim3-1}

\includegraphics[scale=0.6]{figures/sim3-2}

\caption[ClonalOrigin simulation using posterior of recombinant trees]{
{\bf Simulations of the model-based approach using recombinant trees sampled from the posterior distribution.} 
For each recombinant edge type (X-axis), we record the ratio between the average
count of edges as inferred in our data analysis (closed circle; see also
supplementary tables \ref{tab:obsheatmap} \& \ref{tab:heatmap}), and as inferred
across 100 data sets simulated using posterior trees sampled from the posterior
distribution (intervals with open circles). The open circle corresponds to the
median across the different sets, and the intervals correspond to the range
between the  5\% and 95\% quantiles.}
\label{fig:sim3}
\end{sidewaysfigure}
\clearpage{}%
%=============================================================================

%=============================================================================
\begin{sidewaysfigure}

\includegraphics[scale=0.6]{figures/sim4-1}

\includegraphics[scale=0.6]{figures/sim4-2}

\caption[Mowgli simulation for additive transfers and duplications]{
{\bf The power of recovering additive transfers and duplications using
simulations.} (top) For each pair of species tree branches, we plot the number
of inferred additive transfers in the real data (closed circle) normalized by
the expected number of transfers (computed assuming a constant rate of transfer
along the species tree). The same is done for each of the 1,000 simulated data
sets. The open circle corresponds to the median across the different sets, and
the intervals correspond to the range between the  5\% and 95\% quantiles.
(bottom) For each branch in the species tree, we plot the number of inferred
duplications in the real data (closed circles) normalized by the expected number
of duplications in our simulations (computed assuming a constant rate of
duplications along the species tree).  The same is done for each of the 1,000
simulated data sets. The open circle corresponds to the median across the
different sets, and the intervals correspond to the range between the  5\% and
95\% quantiles.}
\label{fig:mowgli-sim}
\end{sidewaysfigure}
\clearpage{}%
%=============================================================================

\end{document}